### 2018-11-29 inst/draft/draft.tex
Added draft

\documentclass[article,nojss,shortnames]{jss}

\usepackage{thumbpdf,lmodern}
\usepackage{framed}
\usepackage{algorithmic}
\usepackage{amsmath}
\usepackage[]{algorithm2e}

% Draft
%\hypersetup{draft}
%\shortcites{rasmussen2012}

\author{Reto Stauffer\\Universit\"at Innsbruck
\And Matthias Dusch\\Universit\"at Innsbruck
\AND Fabien Maussion\\Universit\"at Innsbruck
\And Georg J. Mayr\\Universit\"at Innsbruck}

\Plainauthor{Reto Stauffer, Matthias Dusch, Fabien Maussion, Georg J. Mayr}

\title{phoeton: Software for Objective Probabilistic Foehn Classification}
\Plaintitle{phoeton: Objective Probabilistic Foehn Classification}
\Shorttitle{phoeton: Objective Probabilistic Foehn Classification}

\Abstract{
    This is the abstract here \dots
}

\Keywords{meteorology, foehn, objective classification, flexible mixture model}
\Plainkeywords{meteorology, foehn, objective classification, flexible mixture model}

\Address{
  Reto Stauffer\\
  Department of Statistics\\
  Faculty of Economics and Statistics\\
  Universit\"at Innsbruck\\
  Universit\"atsstr.~15\\
  6020 Innsbruck, Austria\\
  E-mail: \email{Reto.Stauffer@uibk.ac.at}\\
  URL: \url{https://retostauffer.org}\\

  Matthias Dusch, Fabien Maussion, Georg J. Mayr\\
  Department of Atmospheric and Cryospheric Sciences\\
  Faculty of Geo- and Atmospheric Sciences\\
  Universit\"at Innsbruck\\
  Innrain~52f\\
  6020 Innsbruck, Austria\\
  E-mail: \email{Georg.Mayr@uibk.ac.at},
          \email{Matthias.Dusch@uibk.ac.at},
          \email{Fabien.Maussion@uibk.ac.at}
}

\begin{document}

% -------------------------------------------------------------------
% SECTION: INTRODUCTION
% -------------------------------------------------------------------
\section{Introduction}

Introduction why this is so great.

% -------------------------------------------------------------------
% SECTION: METHODS
% -------------------------------------------------------------------
\section{Methods}

\textit{Note by Reto:}
The original publication (\code{FlexMix}, \citealt{gruen2008})
provides general framework for flexible mixture models for multiple
clusters. For this application we will concentrate on the special
case with only two distinct clusters ($K = 2$).

\subsection{Finite Mixture Models with Two Clusters}

The density $h(\dots)$ of a finite mixture model can be written as follows:

\begin{equation}
    h(\mathit{y} ~|~ \mathbf{x}, \mathit{\omega}, \mathit{\alpha}, \mathit{\theta}) =
        \sum_{k=1}^K \pi_k(\mathbf{\omega}, \mathit{\alpha}) \cdot f_k(\mathit{y} ~|~ \mathbf{x}, \mathit{\theta}_k)l
\end{equation}

The density is the sum over the individual densities $f(\dots)$ of the $K$ clusters
times the

.....
....

$\mathit{y}$ denotes the response, $\mathbf{x}$ the predictors for the main part,
and $\mathit{\omega}$ the concomitant variables for the probability model.

... where in our case $K$ is 2 (two groups). $y$ denotes our response, $x$ the predictors
for the main part of the model, and $\omega$ the concomitant variables used for the
probability model.
The density $h(\dots)$
$f_k$ is the density of the (Gaussian) mixture model is the sum over the density
of the two groups $k=1,2$ multiplied by the probability $\pi_k$.

In case of a two-group Gaussian mixture model the density $f_\bullet$ is simply the
Gaussian density $\phi$ defined by the parameters $\theta_k = (\mu_k, \sigma_k)$ and
the probability of an observation being observed in $k=2$ is $1 - \pi_k$.
The only constraint is that $\sum_{k=1}^K = 1$, but basically any probability model
or classifier could be used. We are, for now, using a logistic regression model
with $g(\pi) = \omega^\top \alpha$. $\alpha$ are the regression coefficients from
the logistic regression model.

\subsection{Model Estimation}

To be able to estimate the model an maximum likelihood based EM algorithm is required.
It is assumed that there is an unobservable latent variable $z_n \in \{0, 1\}^K$
which exists for each observation $n = 1, \dots, N$ which defines the membership (the probability
that observation $n$ is a member of class $k$ (e.g., $z_{ik} = 1$ if observation $i$ comes from
class $k$).

Initially we have to guess $z_{nk}$.

$z_{i1} = \begin{cases} 0 & \text{if} ~~ y_i < \text{median}(y) \\ 1 & \text{else}\end{cases}$

In my case I am taking the first and third quartile
of `ff` for $\mu_1$, $\mu_2$ and the standard deviation of the data `ff` for the two
parameters $\sigma_1$, $\sigma_2$. Thus,

$\theta = (\theta_1, \theta_2) = (F^{-1}(y|p = 0.25), \text{sd}(y), F^{-1}(y|p = 0.75), \text{sd}(y))$

For the initial parameters $\alpha$ of the logistic model we can
apply a logistic regression model of the form:

$z = \frac{\exp(\omega^\top \alpha)}{1 + \exp(\omega^\top \alpha)}$

The following lines prepare the prior information $z$, $\theta$, and $\alpha$:

% -------------------------------------------------------------------
% SECTION: SOFTWARE
% -------------------------------------------------------------------
\section{Software}

Explain software.

% -------------------------------------------------------------------
% SECTION: DATA
% -------------------------------------------------------------------
\section{Data}

Maybe just combine data and the case study.

% -------------------------------------------------------------------
% SECTION: RESULTS
% -------------------------------------------------------------------
\section{Case Study/Results}

% -------------------------------------------------------------------
% SECTION: Discussion and Outlook
% -------------------------------------------------------------------
\section{Discussion and Outlook}

% -------------------------------------------------------------------
% Copernicus extra sections
% -------------------------------------------------------------------

\section*{Acknowledgments}

Acknowledgments dow here \dots

% -------------------------------------------------------------------
% Bibliography here
% -------------------------------------------------------------------
\newpage
\bibliography{references}

% -------------------------------------------------------------------
% Appendix section
% -------------------------------------------------------------------
\newpage

\begin{appendix}

\section{Appendix Section}

This is the appendix, if needed \dots

\end{appendix}

\end{document}

### 2018-11-29 R/foehndiag.R

@@ -7,7 +7,7 @@
 # -------------------------------------------------------------------
 # - EDITORIAL:   2018-11-28, RS: Created file on thinkreto.
 # -------------------------------------------------------------------
-# - L@ST MODIFIED: 2018-11-29 12:05 on marvin
+# - L@ST MODIFIED: 2018-11-29 12:52 on marvin
 # -------------------------------------------------------------------
 
 
@@ -308,7 +308,7 @@ print.summary.phoeton <- function(x, ...) {
 #       follow our naming conventions, or make it much more
 #       flexible/generig.
 # -------------------------------------------------------------------
-plot.phoeton <- function(x, start = NULL, end = NULL, ...) {
+plot.phoeton <- function(x, start = NULL, end = NULL, ..., xtra = NULL) {
 
     hold <- par(no.readonly = TRUE); on.exit(par(hold))
 
@@ -318,7 +318,7 @@ plot.phoeton <- function(x, start = NULL, end = NULL, ...) {
     print(end)
     tmp <- window(tmp, start = start, end = end)
 
-    par(mfrow = c(3,1), mar = rep(0.1, 4), xaxs = "i", oma = c(4.1, 4.1, 2, 4.1))
+    par(mfrow = c(4,1), mar = rep(0.1, 4), xaxs = "i", oma = c(4.1, 4.1, 2, 4.1))
 
     # Air temperature
     plot(tmp$t, col ="red", lwd = 2, ylab = NA, xaxt = "n")
@@ -330,6 +330,10 @@ plot.phoeton <- function(x, start = NULL, end = NULL, ...) {
     abline(h = seq(20, 100, by = 20), lty = 3, col = "gray")
     axis(side = 4, at = seq(20, 100, by = 20))
     mtext(side = 4, line = 3, "relative humidity")
+
+    # Temperature difference
+    plot(tmp$diff_t, col = "magenta", lwd = 2, xaxt = "n")
+    mtext(side = 2, line = 3, "temperature difference")
 
     # Wind speed and direction
     if ( ! is.null(x$windsector) ) {
@@ -357,6 +361,7 @@ plot.phoeton <- function(x, start = NULL, end = NULL, ...) {
     plot(tmp$prob * 100, ylab = NA, col = "blue", ylim = c(0,100), yaxs = "i") 
     abline(h = seq(20, 80, by = 20), col = "gray", lty = 3)
     mtext(side = 2, line = 3, "foehn probability")
[…]

### 2018-11-30 inst/draft/draft.tex
Updated text

@@
 \usepackage[]{algorithm2e}
+
+\newcommand{\note}[1]{\textcolor{red}{\textbf{Note:} #1}}
 
@@
 
-Introduction why this is so great.
+Introduction why this is so great. Georg might like to contribute
+here. Just to leave/include some references:
+
+\begin{itemize}
+    \item finite mixture model \citep{leisch2004,gruen2008}
+    \item automatic probabilistic foehn classification \cite{plavcan2014}
+    \item IWLS/IRLS \citep{mccullagh1989}
+\end{itemize}
 
@@
 
-\subsection{Finite Mixture Models with Two Clusters}
+\subsection{Finite mixture models}
 
-The density $h(\dots)$ of a finite mixture model can be written as follows:
+The finite mixture model in its general form is given by
 
 \begin{equation}
-    h(\mathit{y} ~|~ \mathbf{x}, \mathit{\omega}, \mathit{\alpha}, \mathit{\theta}) =
-        \sum_{k=1}^K \pi_k(\mathbf{\omega}, \mathit{\alpha}) \cdot f_k(\mathit{y} ~|~ \mathbf{x}, \mathit{\theta}_k)l
+    h(\mathit{y} | \mathit{x}, \mathit{\omega}, \mathit{\alpha}, \mathit{\theta}) =
+        \sum_{k=1}^K \pi_k(\mathit{\alpha}, \mathbf{\omega}) \cdot f_k(\mathit{y} | \mathit{x}, \mathit{\theta}_k)
+        ~~\text{for}~~k=1,\dots,K
+    \label{eqn:flexmix-density-general}
 \end{equation}
 
-The density is the sum over the individual densities $f(\dots)$ of the $K$ clusters
-times the
+where $\mathit{\theta}$ and $\mathit{\alpha}$ denote the parameters for the
+mixture density $h()$. $\mathit{y}$ denotes the response, $\mathit{x}$ the
+predictor variable (one single variable) and $\mathbf{\omega}$ the concomitant
+variables (multiple variables allowed) for the probability model.
 
-.....
-....
+For the special case of a finite mixture model with $K=2$ Gaussian clusters
+the specification as shown in Equation~\ref{eqn:flexmix-density-general} simplifies to
 
-$\mathit{y}$ denotes the response, $\mathbf{x}$ the predictors for the main part,
-and $\mathit{\omega}$ the concomitant variables for the probability model.
+\begin{equation}
+    h(\mathit{y} | \mathit{x}, \mathit{\omega}, \mathit{\alpha}, \mathit{\theta}) =
+        \underbrace{
+            (1 - \mathit{\pi}(\mathbf{\omega}, \mathit{\alpha})) \cdot
+            \phi(\mathit{y} | \mathit{\theta}_1, \mathit{x})
+        }_{\text{cluster 1}}
+        \cdot
+        \underbrace{
+            \mathit{\pi}(\mathbf{\omega}, \mathit{\alpha}) \cdot \phi(\mathit{\theta}_2|\mathit{x})
+        }_{\text{cluster 2}}
+    \label{eqn:flexmix-density-gaussian}
+\end{equation}
 
-... where in our case $K$ is 2 (two groups). $y$ denotes our response, $x$ the predictors
-for the main part of the model, and $\omega$ the concomitant variables used for the
-probability model.
-The density $h(\dots)$
-$f_k$ is the density of the (Gaussian) mixture model is the sum over the density
-of the two groups $k=1,2$ multiplied by the probability $\pi_k$.
+where $\phi()$ is the probability function of the Gaussian distribution.
+The concomitant model for $\pi()$ reduces to a binomial problem as the probability
+for a specific observation $y_i$ to be observed in the first cluster is the
+counter probability of $y_i$ being observed in the second cluster.
+Thus, for $\pi()$ any type of binomial model can be used which satisfies
+that $\pi \in [0,1] \forall i=1,\dots,N$.
 
-In case of a two-group Gaussian mixture model the density $f_\bullet$ is simply the
-Gaussian density $\phi$ defined by the parameters $\theta_k = (\mu_k, \sigma_k)$ and
-the probability of an observation being observed in $k=2$ is $1 - \pi_k$.
-The only constraint is that $\sum_{k=1}^K = 1$, but basically any probability model
-or classifier could be used. We are, for now, using a logistic regression model
-with $g(\pi) = \omega^\top \alpha$. $\alpha$ are the regression coefficients from
-the logistic regression model.
+One frequently used model for a binary response is the binomial logistic regression
+model or binomial logit model (BLM), however, any appropriate binomial probability
+model could be used for $\pi$.
+The BLM is defined as follows:
 
-\subsection{Model Estimation}
+\begin{equation}
+    \log\Big(\frac{\pi}{1 - \pi}\Big) =
+        \omega^\top \alpha = \alpha_0 + \omega_1 \alpha_1 + \dots + \omega_P \alpha_P,
+\end{equation}
 
-To be able to estimate the model an maximum likelihood based EM algorithm is required.
-It is assumed that there is an unobservable latent variable $z_n \in \{0, 1\}^K$
-which exists for each observation $n = 1, \dots, N$ which defines the membership (the probability
-that observation $n$ is a member of class $k$ (e.g., $z_{ik} = 1$ if observation $i$ comes from
-class $k$).
+where $\mathit{\pi}$ is the probability that an observation belongs to the
+second cluster (Eqn.~\ref{eqn:flexmix-density-gaussian}).
+The probabilities are modelled using multilinear regression
+$\mathit{\omega}^\top \mathit{\alpha}$ with $p=1,\dots,P$ concomitant
+variables $\mathit{\omega}$ times the corresponding parameters or regression
+coefficients $\mathit{\alpha}$.
 
-Initially we have to guess $z_{nk}$.
+As the BLM is a generalized linear model, the parameters can be estimated using
+an iterative (re-)weighted least squares (IWLS) solver.
+The IWLS procedure to estimate the parameters $\mathit{\alpha}$ of the BLM
+is shown in Algorithm~\ref{alg:iwls}.
 
-$z_{i1} = \begin{cases} 0 & \text{if} ~~ y_i < \text{median}(y) \\ 1 & \text{else}\end{cases}$
+\begin{algorithm}
+    \caption{Iterative (re-)weighted least squares (IWLS) solver for the binomial logit model (BLM).}
+    \label{alg:iwls}
 
-In my case I am taking the first and third quartile
-of `ff` for $\mu_1$, $\mu_2$ and the standard deviation of the data `ff` for the two
-parameters $\sigma_1$, $\sigma_2$. Thus,
+    \textbf{Initialization ($j = 0$):}
+    \begin{enumerate}
+        \item Initialize $\mathit{\alpha}^{0} = 0$.
+        \item Compute initial latent response $\mathit{\eta}^{(0)} = \mathbf{\omega}^\top\mathit{\alpha}^{(0)}$.
+        \item Compute initial probabilities $\pi^{(0)} = \frac{\exp(\eta^{(0)})}{1 + \exp(\eta^{(0)})}$. 
+        \item Evaluate initial log-likelihood:
+            \begin{equation*}
+                \ell^{(0)} = %(\mathit{\alpha} | \mathbf{\omega})^{(0)} =
+                \sum_{i=1}^{N} \Big(\log(1 - \pi_i^{(0)})^{(y_i = 0)} + \log(\pi_i)^{(y_i=1)}\Big)
+            \end{equation*}
 
-$\theta = (\theta_1, \theta_2) = (F^{-1}(y|p = 0.25), \text{sd}(y), F^{-1}(y|p = 0.75), \text{sd}(y))$
+    \end{enumerate}
+    \textbf{Iterative parameter estimation ($j = 1, \dots, J$):}
+    \begin{enumerate}
+        \item Compute new inverse square root weights:
+            \begin{equation*}
+                \mathit{W}^{(j)} = \Big(\big(\frac{d\eta}{d\pi}\big)^{(j-1)}\Big)^{-\frac{1}{2}} =
+                    \big(\mathit{\pi}^{(j-1)} (1 - \mathit{\pi}^{(j-1)})\big)^\frac{1}{2}
+            \end{equation*}
+        %%%\item Compute new latent response:
+        %%%    \begin{equation*}
+        %%%        \mathit{z}^{(j+1)} = \mathit{\eta}^{(j)} + (\mathit{y} - \mathit{\pi}^{(j)})
+        %%%            \Big(\frac{d \mathit{\eta}}{d \mathit{\pi}}\Big)^{(j)} =
+        %%%            \mathit{\eta}^{(j)} + \frac{\mathit{y} - \mathit{\pi}^{(j)}}{\mathit{\pi}^{(j)}(1 - \mathit{\pi}^{(j)})}
+        %%%    \end{equation*}
+        \item Update parameters:
+            \begin{equation*}
+                \mathit{\alpha}^{(j)} =
+                    \big((\mathbf{\omega} \mathit{W}^{(j)})^\top
+                    (\mathbf{\omega} \mathit{W}^{(j)})\big)^{-1}
+                    (\mathbf{\omega} \mathit{W}^{(j)})^\top
+                    \mathit{W}^{(j)} \mathit{z}^{(j)}
+            \end{equation*}
 
-For the initial parameters $\alpha$ of the logistic model we can
-apply a logistic regression model of the form:
+            with $\mathit{z}^{(j)} = \mathit{\eta}^{(j-1)} + (\mathit{y} - \mathit{\pi}^{(j-1)}) \Big(\frac{d\pi}{d\eta}\Big)^{(j-1)}$ this yields
 
-$z = \frac{\exp(\omega^\top \alpha)}{1 + \exp(\omega^\top \alpha)}$
+            \begin{equation*}
+                \mathit{\alpha}^{(j)} =
+                    \big((\mathbf{\omega} \mathit{W}^{(j)})^\top
+                    (\mathbf{\omega} \mathit{W}^{(j)})\big)^{-1}
+                    (\mathbf{\omega} \mathit{W}^{(j)})^\top
+                    \Big(\mathit{\eta}^{(j-1)} \mathit{W}^{(j)} + \frac{\mathit{y} - \mathit{\pi}^{(j-1)}}{W^{(j)}}\Big)
+            \end{equation*}
+
+        \item Update response:
+            \begin{equation*}
+                \mathit{\eta}^{(j)} = \mathbf{\omega}\mathit{\alpha}^{(j)}
+                ~~\text{and}~~
+                \mathit{\pi}^{(j)} = \frac{\exp(\mathit{\eta}^{(j)})}{1 + \exp(\mathit{\eta}^{(j)})}
+            \end{equation*}
+
+        \item Evaluate log-likelihood sum:
+            \begin{equation*}
+                \ell^{(j)} = %(\mathit{\alpha}^{(j+1)} | \mathbf{\omega}) =
+                \sum_{i=1}^{N} \Big(\log(1 - \pi_i^{(j)})^{(y_i = 0)} + \log(\pi_i^{(j)})^{(y_i=1)}\Big)
+            \end{equation*}
+
+        \item Repeat until convergence, e.g., until the log-likelihood improvement
+            $\ell^{(j)} - \ell^{(j-1)}$ falls below a certain threshold or $j$ reaches
+            $J$, the maximum number of iterations.
+
+    \end{enumerate}
+\end{algorithm}
+
+\subsection{Model estimation}
+
+The parameters $\mathit{\theta}$ and $\mathit{\alpha}$ of the finite mixture model
+can be estimated using a likelihood based EM algorithm (\note{citation?}).
+
+For the special case of a two-cluster Gaussian finite mixture model the estimation
+can be performed using iterative (re-)weighted empirical moments for the parameters
+of the Gaussian densities ($\mathit{\theta}$) and IWLS to estimate the parameters
+of the binomial logit model ($\mathit{\theta}$).
+\note{This is true for $K>2$ as well, only the probability model would have to be adjusted.}
+
+It is assumed that an unobservable latent variable $z_i \in [0,1]$ exists
+for each observation $i=1, \dots N$ which defines the membership of the observation,
+the probability that observation $n$ comes from the second cluster.
+As the membership is unknown it has to be initially guessed:
+
+\begin{equation}
+    \mathit{z}^{(0)} = \begin{cases}
+        1 & \text{if}~x_n > \text{median}(x), \\
+        0 & \text{else}.
+    \end{cases}
+    \label{eqn:init-zn}
+\end{equation}
+\note{This $z$ is basically the initial guess of the probability!}
+
+In addition, initial values for the parameters of the two Gaussian distributions
+for cluster one and cluster two have to be estimated which are used as starting
+values for the EM procedure. The parameters $\mathit{\theta}$ are thus initialized
+as follows:
+
+\begin{equation}
+    \begin{split}
+        \mathit{\theta}_1^{(0)} = (\mu_1^{(0)}, \sigma_1^{(0)}) = \Big(q_{0.25}(x), \text{sd}(x)\Big) \\
+        \mathit{\theta}_2^{(0)} = (\mu_2^{(0)}, \sigma_2^{(0)}) = \Big(q_{0.75}(x), \text{sd}(x)\Big)
+    \end{split}
+    \label{eqn:init-theta}
+\end{equation}
+
+The first ($q_{0.25}()$) and the third ($q_{0.75}()$) quartile of the predictor 
+variable $\mathit{x}$ are used as inital guess for the location parameters
+$\hat{\mu}_1$ and $\hat{\mu}_2$ while the standard deviation of $\mathit{x}$ is used as initial
+guess for the scale parameter $\hat{\sigma}_1$ and $\hat{\sigma}_2$.
+The parameters $\mathit{\alpha}$ of the binomial logit model are simply initialized
+with $0$ (\note{find a way for better initial values}).
+
+The EM algorithm can be written as follows:
+
+\begin{algorithm}
+    \caption{Parameter estimation for the two-cluster Gaussian finite mixture model
+    using iterative (re-)weighted optimization. $\Phi()$ is the cumulative distribution
+    function of the standard Gaussian distribution.}
+
+    \textbf{Initialization ($j = 0$):}
+    \begin{enumerate}
+        \item Initialize $\mathit{z}$ and $\mathit{\theta}$ as shown in
+            Equation~\ref{eqn:init-zn}\,\&\,\ref{eqn:init-theta}.\label{item:a2:1}
+
+        \item Estimate initial $\alpha^{(0)}$ parameters using $\mathit{z}$ from
+            Step~\ref{item:a2:1} as $\mathit{y}$ and the IWLS solver from
+            Algorithm~\ref{alg:iwls}.
+
+        \item Compute response of the BLM with initial parameters:
+            \begin{equation*}
+                \mathit{\eta}^{(0)} = \mathbf{\omega}^\top \mathit{\alpha}^{(0)}
+                ~~\text{and}~~
+                \mathit{\pi}^{(0)} = \frac{\exp(\mathit{\eta}^{(0)})}{1 - \exp(\mathit{\eta}^{(0)})}
+            \end{equation*}
+    \end{enumerate}
+
+    \textbf{EM iterations ($j = 1, \dots, J$):}
+    \begin{enumerate}
+
+        \item Compute finite mixture model densities for both clusters:
+            \begin{equation*}
+                    d_1^{(j)} = (1 - \mathit{\pi}^{(j)}) \Phi\Big(\frac{\mathit{y} - \mu_1^{(j-1)}}{\sigma_1^{(j-1)}}\Big),
+                    ~~
+                    d_2^{(j)} = \mathit{\pi}^{(j)} \Phi\Big(\frac{\mathit{y} - \mu_2^{(j-1)}}{\sigma_2^{(j-1)}}\Big)
+            \end{equation*}
+
+        \item Compute posterior probabilities:
+            \begin{equation*}
+                \tilde{\mathit{p}}^{(j)} = \frac{d_2^{(j)}}{d_1^{(j)} + d_2^{(j)}}
+            \end{equation*}
+
+        %%%\item Compute initial log-likelihood sum:
+        %%%    \begin{equation*}
+        %%%        \ell^{(j)} = \sum_{i=1}^N 
+        %%%        \underbrace{
+        %%%            (1 - \tilde{p}_i^{(j)}) \log(d_{1,i}^{(j)})
+        %%%        }_{\text{cluster 1}}
+        %%%        +
+        %%%        \underbrace{
+        %%%            \tilde{p}_i^{(j)} \log(d_{2,i}^{(j)})
+        %%%        }_{\text{cluster 2}}
+        %%%        +
+        %%%        \underbrace{
+        %%%            (1 - \tilde{p}_i^{(j)}) \log(1 - \pi_i^{(j)}) + \tilde{p}_i^{(j)} \log(\pi_i^{(j)})
+        %%%        }_{\text{concomitant model}}
+        %%%        \Big)
+        %%%    \end{equation*}
+
+        \item Update $\mathit{\theta}$ using weighted empirical moments:
+            \begin{equation*}
+                \begin{split}
+                    \mu_1^{(j)} = \frac{\sum \mathit{x} (1 - \tilde{p}^{(j)})}{\sum (1 - \tilde{p}^{(j)})}
+                    ~~\text{and}~~
+                    \sigma_1^{(j)} = \Big( \frac{\sum (x - \mu_1^{(j)})^2 (1 - \tilde{p}^{(j)})}{\sum(1 - \tilde{p}^{(j)})} \Big)^\frac{1}{2} \\
+                    \mu_2^{(j)} = \frac{\sum \mathit{x} \tilde{p}^{(j)}}{\sum \tilde{p}^{(j)}}
+                    ~~\text{and}~~
+                    \sigma_2^{(j)} = \Big( \frac{\sum (x - \mu_2^{(j)})^2 \tilde{p}^{(j)}}{\sum \tilde{p}^{(j)}} \Big)^\frac{1}{2} \\
+                \end{split}
+            \end{equation*}
+
+        \item Update $\mathit{\alpha}$ parameters using the new probabilities $\pi^{(j)}$
+            as $\mathit{y}$ and the IWLS solver from Algorithm~\ref{alg:iwls}.
+
+        \item Update response of the concomitant model:
+            \begin{equation*}
+                \mathit{\eta}^{(j)} = \mathbf{\omega}\mathit{\alpha}^{(j)}
+                ~~\text{and}~~
+                \mathit{\pi}^{(j)} = \frac{\exp(\mathit{\eta}^{(j)})}{1 + \exp(\mathit{\eta}^{(j)})}
+            \end{equation*}
+
+        \item Evaluate log-likelihood given the new parameters $\mathit{\alpha}$ and $\mathit{\theta}$
+            as well as the updated posterior probabilities $\mathit{\tilde{p}}$:
+            \begin{equation*}
+                \ell^{(j)} = \sum_{i=1}^N 
+                \underbrace{
+                    (1 - \tilde{p}_i^{(j)}) \log(d_{1,i}^{(j)})
+                }_{\text{cluster 1}}
+                +
+                \underbrace{
+                    \tilde{p}_i^{(j)} \log(d_{2,i}^{(j)})
+                }_{\text{cluster 2}}
+                +
+                \underbrace{
+                    (1 - \tilde{p}_i^{(j)}) \log(1 - \pi_i^{(j)}) + \tilde{p}_i^{(j)} \log(\pi_i^{(j)})
+                }_{\text{concomitant model}}
+                \Big)
+            \end{equation*}
+
+        \item Stop if $j = J$ and the maximum number of iterations is reached or
+            (for $j > 1$) if the improvements in the log-likelihood $\ell^{(j)} - \ell^{(j-1)}$
+            fall below a certain threshold.
+    \end{enumerate}
+\end{algorithm}
+
+\begin{equation}
+    z_{i1} = \begin{cases} 0 & \text{if} ~~ y_i < \text{median}(y) \\ 1 & \text{else}\end{cases}
+\end{equation}
+
+\begin{equation}
+    z = \frac{\exp(\omega^\top \alpha)}{1 + \exp(\omega^\top \alpha)}
+\end{equation}
 
@@
 % -------------------------------------------------------------------
-\section{Case Study/Results}
+\section{Case study/results}
 
@@
 % -------------------------------------------------------------------
-\section{Discussion and Outlook}
+\section{Discussion and outlook}
 

### 2018-11-30 inst/draft/draft.tex
Updated draft

@@
 provides general framework for flexible mixture models for multiple
-clusters. For this application we will concentrate on the special
-case with only two distinct clusters ($K = 2$).
+components. For this application we will concentrate on the special
+case with only two distinct components ($K = 2$).
 
@@
 
-The finite mixture model in its general form is given by
+Finite mixture models are a general model class of finite
+mixtures of regression models. It is assumed that the mixture
+consists of a set of $K$ distinct components, each described by
+a parametric distribution.
+In addition, an weight is assigned to each of the $K$ classes
+for each observation which describes the a-priori probability
+for an observation to come from this component.
+These weights can depend on multiple covariates which are
+referred to as \textit{concomitant} variables.
 
@@
 predictor variable (one single variable) and $\mathbf{\omega}$ the concomitant
-variables (multiple variables allowed) for the probability model.
+variables for the probability model.
 
-For the special case of a finite mixture model with $K=2$ Gaussian clusters
+For the special case of a finite mixture model with $K=2$ Gaussian components
 the specification as shown in Equation~\ref{eqn:flexmix-density-general} simplifies to
@@
             \phi(\mathit{y} | \mathit{\theta}_1, \mathit{x})
-        }_{\text{cluster 1}}
+        }_{\text{component 1}}
         \cdot
@@
             \mathit{\pi}(\mathbf{\omega}, \mathit{\alpha}) \cdot \phi(\mathit{\theta}_2|\mathit{x})
-        }_{\text{cluster 2}}
+        }_{\text{component 2}}
     \label{eqn:flexmix-density-gaussian}
@@
 The concomitant model for $\pi()$ reduces to a binomial problem as the probability
-for a specific observation $y_i$ to be observed in the first cluster is the
-counter probability of $y_i$ being observed in the second cluster.
+for a specific observation $y_i$ to be observed in the first component is the
+counter probability of $y_i$ being observed in the second component.
 Thus, for $\pi()$ any type of binomial model can be used which satisfies
 that $\pi \in [0,1] \forall i=1,\dots,N$.
+Figure~\ref{fig:schematic_density} shows an example of a finite mixture model with
+two Gaussian components with fixed $\mathit{\theta}$ coefficients but different
+weights or a-priori probabilities $\tilde{p}_k$.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width = .8\textwidth]{draft_schematic_density}
+    \caption{Schematic representation of the finite mixture model density
+    (shaded; red dashed line) and the two Gaussian components (black/gray). 
+    $\tilde{p}_1$ and $\tilde{p}_2$ are the a-prior weights for the components.
+    The dashed lines always show the distribution of the component with equal
+    and constant weights of $\tilde{p}_1 = \tilde{p}_2 = 0.5$.}
+    \label{fig:schematic_density}
+\end{figure}
 
@@
 where $\mathit{\pi}$ is the probability that an observation belongs to the
-second cluster (Eqn.~\ref{eqn:flexmix-density-gaussian}).
+second component (Eqn.~\ref{eqn:flexmix-density-gaussian}).
 The probabilities are modelled using multilinear regression
@@
 
-For the special case of a two-cluster Gaussian finite mixture model the estimation
+For the special case of a two-component Gaussian finite mixture model the estimation
 can be performed using iterative (re-)weighted empirical moments for the parameters
@@
 for each observation $i=1, \dots N$ which defines the membership of the observation,
-the probability that observation $n$ comes from the second cluster.
+the probability that observation $n$ comes from the second component.
 As the membership is unknown it has to be initially guessed:
@@
 In addition, initial values for the parameters of the two Gaussian distributions
-for cluster one and cluster two have to be estimated which are used as starting
+for component one and component two have to be estimated which are used as starting
 values for the EM procedure. The parameters $\mathit{\theta}$ are thus initialized
@@
 \begin{algorithm}
-    \caption{Parameter estimation for the two-cluster Gaussian finite mixture model
+    \caption{Parameter estimation for the two-component Gaussian finite mixture model
     using iterative (re-)weighted optimization. $\Phi()$ is the cumulative distribution
-    function of the standard Gaussian distribution.}
+    function of the Gaussian distribution.}
 
@@
 
-        \item Compute finite mixture model densities for both clusters:
-            \begin{equation*}
-                    d_1^{(j)} = (1 - \mathit{\pi}^{(j)}) \Phi\Big(\frac{\mathit{y} - \mu_1^{(j-1)}}{\sigma_1^{(j-1)}}\Big),
-                    ~~
-                    d_2^{(j)} = \mathit{\pi}^{(j)} \Phi\Big(\frac{\mathit{y} - \mu_2^{(j-1)}}{\sigma_2^{(j-1)}}\Big)
-            \end{equation*}
+        %%\item Compute finite mixture model densities for both components:
+        %%    \begin{equation*}
+        %%            d_1^{(j)} = (1 - \mathit{\pi}^{(j)}) \Phi\Big(\frac{\mathit{y} - \mu_1^{(j-1)}}{\sigma_1^{(j-1)}}\Big),
+        %%            ~~
+        %%            d_2^{(j)} = \mathit{\pi}^{(j)} \Phi\Big(\frac{\mathit{y} - \mu_2^{(j-1)}}{\sigma_2^{(j-1)}}\Big)
+        %%    \end{equation*}
+
+        %%\item Compute posterior probabilities:
+        %%    \begin{equation*}
+        %%        \tilde{\mathit{p}}^{(j)} = \frac{d_2^{(j)}}{d_1^{(j)} + d_2^{(j)}}
+        %%    \end{equation*}
 
@@
             \begin{equation*}
-                \tilde{\mathit{p}}^{(j)} = \frac{d_2^{(j)}}{d_1^{(j)} + d_2^{(j)}}
+                \tilde{\mathit{p}}^{(j)} = 
+                \frac{
+                    \mathit{\pi}^{(j-1)} \Phi(\mathit{y} | \mu_2^{(j-1)}, \sigma_2^{(j-1)})
+                }{
+                    (1 - \mathit{\pi}^{(j-1)}) \Phi(\mathit{y} | \mu_1^{(j-1)}, \sigma_1^{(j-1)}) +
+                    \mathit{\pi}^{(j-1)} \Phi(\mathit{y} | \mu_2^{(j-1)}, \sigma_2^{(j-1)})
+                }
             \end{equation*}
@@
         %%%            (1 - \tilde{p}_i^{(j)}) \log(d_{1,i}^{(j)})
-        %%%        }_{\text{cluster 1}}
+        %%%        }_{\text{component 1}}
         %%%        +
@@
         %%%            \tilde{p}_i^{(j)} \log(d_{2,i}^{(j)})
-        %%%        }_{\text{cluster 2}}
+        %%%        }_{\text{component 2}}
         %%%        +
@@
         %%%    \end{equation*}
+        \item Update $\mathit{\theta}$ using weighted empirical moments:
+            \begin{eqnarray*}
+                    \mu_1^{(j)} = \frac{\sum \mathit{x} (1 - \tilde{p}^{(j)})}{\sum (1 - \tilde{p}^{(j)})}
+                    &\text{and}&
+                    \sigma_1^{(j)} = \Big( \frac{\sum (x - \mu_1^{(j)})^2 (1 - \tilde{p}^{(j)})}{\sum(1 - \tilde{p}^{(j)})} \Big)^\frac{1}{2} \\
+                    \mu_2^{(j)} = \frac{\sum \mathit{x} \tilde{p}^{(j)}}{\sum \tilde{p}^{(j)}}
+                    &\text{and}&
+                    \sigma_2^{(j)} = \Big( \frac{\sum (x - \mu_2^{(j)})^2 \tilde{p}^{(j)}}{\sum \tilde{p}^{(j)}} \Big)^\frac{1}{2} \\
+            \end{eqnarray*}
 
@@
             \begin{equation*}
-                \ell^{(j)} = \sum_{i=1}^N 
+                \ell^{(j)} = \sum_{i=1}^N \Big(
                 \underbrace{
                     (1 - \tilde{p}_i^{(j)}) \log(d_{1,i}^{(j)})
-                }_{\text{cluster 1}}
+                }_{\text{component 1}}
                 +
@@
                     \tilde{p}_i^{(j)} \log(d_{2,i}^{(j)})
-                }_{\text{cluster 2}}
+                }_{\text{component 2}}
                 +

### 2018-11-30 R/foehndiag.R

@@ -7,7 +7,7 @@
 # -------------------------------------------------------------------
 # - EDITORIAL:   2018-11-28, RS: Created file on thinkreto.
 # -------------------------------------------------------------------
-# - L@ST MODIFIED: 2018-11-29 12:52 on marvin
+# - L@ST MODIFIED: 2018-11-30 14:10 on marvin
 # -------------------------------------------------------------------
 
 
@@ -144,10 +144,8 @@ foehndiag <- function(formula, data, windsector = NULL, maxit = 100L, tol = 1e-8
     # assignment.
     alpha <- iwls_logit(logitX, as.numeric(y > median(y)), maxit = tail(maxit, 1), tol = tail(tol, 1))
 
-    # Initial log-likelihood given the initial guess
-    prob <- plogis(drop(logitX %*% alpha))
-    post <- foehndiag_gauss_posterior(y, prob, theta)
-    ll0  <- foehndiag_gauss_loglik(y, post, prob, theta)$ll - 1000
+    # Initial value
+    ll0 <- NULL
 
     # Start optimization using weighted empirical moments for
     # location and scale of the two Gaussian distributions plus
@@ -171,6 +169,11 @@ foehndiag <- function(formula, data, windsector = NULL, maxit = 100L, tol = 1e-8
         #        ((1 - prob) * dnorm(y, theta$mu1, exp(theta$logsd1)) +
         #         prob * dnorm(y, theta$mu2, exp(theta$logsd2)))
         post <- foehndiag_gauss_posterior(y, prob, theta)
+
+        # Initial log-likelihood given the initial guess
+        #prob <- plogis(drop(logitX %*% alpha))
+        #post <- foehndiag_gauss_posterior(y, prob, theta)
+        if ( is.null(ll0) ) ll0  <- foehndiag_gauss_loglik(y, post, prob, theta)$ll - 1000
     
         ##### M-Step ######
         # Empirical moments for the two clusters (mu1/sd1 and mu2/sd2)

### 2018-11-30 inst/draft/draft_images.R

@@ -1,0 +1,90 @@
+# -------------------------------------------------------------------
+# - NAME:        images.R
+# - AUTHOR:      Reto Stauffer
+# - DATE:        2018-11-30
+# -------------------------------------------------------------------
+# - DESCRIPTION:
+# -------------------------------------------------------------------
+# - EDITORIAL:   2018-11-30, RS: Created file on thinkreto.
+# -------------------------------------------------------------------
+# - L@ST MODIFIED: 2018-11-30 17:16 on marvin
+# -------------------------------------------------------------------
+
+library("latex2exp")
+
+theta <- list(mu1 = 3, sd1 = 1, mu2 = 6, sd2 = 1.4)
+
+x <- seq(-2, 20, length = 1000)
+d1 <- dnorm(x, theta$mu1, theta$sd1)
+d2 <- dnorm(x, theta$mu2, theta$sd2)
+
+c1 <- "#000000"
+c2 <- "#9ea3b6"
+
+# -------------------------------------------------------------
+# -------------------------------------------------------------
+pdf(file = "draft_schematic_density.pdf", width = 8, height = 2.3)
+    par(mfrow = c(1,3), cex = 1, mar = c(2,2,.5,1))
+    
+    # --------------------------------------
+    plot(NA, xlim = c(0, 10), ylim = c(0, max(d1, d2))*1.05, yaxs = "i",
+         xaxt = "n", yaxt = "n", bty = "l", xlab = NA, ylab = NA, main = NA)
+    mtext(side = 1, line = 3, "x")
+    
+    polygon(x, 0.5 * d1 + 0.5 * d2, col = "gray90", border = NA)
+    lines(x, 0.5 * d1 + 0.5 * d2, col = 2, lty = 3, lwd = 2)
+    lines(x, 0.5 * d1, col = c1, lwd = 2)
+    lines(x, 0.5 * d2, col = c2, lwd = 2)
+    
+    axis(side = 1, col.axis = c1, col = c1, at = theta$mu1, expression(mu[1]))
+    axis(side = 1, col.axis = c2, col = c2, at = theta$mu2, expression(mu[2]))
[…]

### 2018-11-30 inst/draft/draft.tex
Updated draft and refs

@@
 here. Just to leave/include some references:
-
 \begin{itemize}
@@
     \item IWLS/IRLS \citep{mccullagh1989}
+    \item Community foehn classification experiment \cite{mayr2018unpublished}
 \end{itemize}
+
+I would go for something like:
+
+Gap winds are typically characterized by strong ncrease of wind speed with a rapid
+increase in temperature and a drop in relative humidity. Such winds can be
+found in various regions on our globe and bear different names for specific
+regions such as the Santa-Anna winds in California which became infamous due to
+their destructive effect in combination with wilde fires, the Chinook winds in
+the Rocky Mountain area, Mistral and Bora in south France and Croatia, or the
+Foehn winds in the European Alps.  Over the past decades \textit{foehn wind}
+has become a synonym for such gap flows and will be used in this article,
+however, the method as presented is applicable for all gap flow winds
+independent from their geographical location.
+
+The conceptual model of foehn winds is quite well known
+\citep{armi2007,mayr2008,mayr2011}, however, foehn cannot measured directly.
+Historically foehn was often diagnosed manually by looking at available
+meteorological records such as wind speed and direction, time series of
+temperature and humidity, the conditions of the sky (e.g., occurence of
+lenticularis clouds in the region), or observations upstream and downstream of
+the location of interest.  Some of the longest records of foehn observations
+can be found in Switzerland (Altdorf \note{i'm just guessing!}) reaching back
+to 1850 (\note{just guessing!}) for a hand full of stations.  These records are
+typically based manual classification made by experienced weather observers,
+weather enthusiasts, and meteorologists.  This task is time consuming and is
+only done for some selected sites, typically sites with long historical
+records. To be able to extend the diagnosis and to perform foehn classification
+on a large number of meteorological sites computer amed classification is
+needed.
+
+Few attepmts to automatically classify foehn winds have been proposed over the
+past years such as \note{cite this and that}. These methods are only binary
+(either \textit{yes} or \textit{no}) and reqire manual intervention to tune the
+hyper-parameters, e.g., thresholds used to classify a specific observation as
+foehn. In \citeyear{plavcan2014} \citeauthor{plavcan2014} proposed a new
+objective and probabilistic foehn classification method based on a finite
+mixture model, a statistical regression-based model for automatic foehn
+classification. In a recent study of \cite{mayr2018unpublished} this new method
+has been tested against existing methods, students, and experts to to proof the
+diagnostic power of the new approach.
+
+Overall, the automatic finite mixture model based foehn classification
+\citep{plavcan2014} is able to compete with the human classification and 
+has the advantages that it is (i) reproductable, and (ii) can easily be
+applied to large volumes of data. This allows one to perform foehn classification
+for a wide range of different sites.
+
+The article by \citep{plavcan2014} introduced the method using a powerful
+package called \code{FlexMix} \citep{leisch2004,gruen2007,gruen2008} available
+for the statistical software \textit{R}.  However, only little guidance was
+given by \citep{plavcan2014} on how to apply this method to own data and an
+implementation of the method outside the \textit{R} programming language has
+was not available until today.
+Thus, the aim of this article is to introduce the algorithm behind the
+automated probabilistic foehn classification \citep{plavcan2014} in more
+detail and introducing new open-source software for \textit{R} and \textit{Python}
+which provide the required methods to perform the classification and a set
+of diagnostic tools for this specific task.
+
+Section~\ref{sec:methods} describes the method itself followed by
+Section~\ref{sec:software}  introducing the software with a set of
+examples. Section~\dots \note{case study? Required for this article?}
 
@@
 % -------------------------------------------------------------------
-\section{Methods}
+\section{Methods}\label{sec:methods}
 
@@
 case with only two distinct components ($K = 2$).
+
+Thus, this section introduces a simplified approach to estimate 
+a finite mixture model for the special case with two distinct
+Gaussian distributions and a binomial logit model to estimate
+the coefficients for the concomitant variables.
+
+\note{\dots ... and so far, and so on.}
 
@@
 % -------------------------------------------------------------------
-\section{Software}
+\section{Software}\label{sec:software}
 

### 2018-12-01 inst/draft/draft.tex
Fixed some of the serious typos, still very early note file and not really a draft

--- 2018-11-30 inst/draft/draft.tex
@@
     \item IWLS/IRLS \citep{mccullagh1989}
+    \item Community foehn classification experiment \cite{mayr2018unpublished}
 \end{itemize}
+
+I would go for something like:
+
+Gap winds are typically characterized by strong increase of wind speed with a rapid
+rise in temperature and a drop in relative humidity during the onset. Such winds can be
+found in various regions on our globe and bear different names for specific
+regions such as the Santa-Anna winds in California which became infamous due to
+their destructive effect in combination with wild fires, the Chinook winds in
+the Rocky Mountain area, Mistral and Bora in south France and Croatia, or the
+Foehn winds in the European Alps.  Over the past decades \textit{foehn} wind
+has become a synonym for such gap flows and will be used in this article.
+However, the method as presented is applicable for all gap flow winds
+independent from their geographical location.
+
+The conceptual model of foehn winds is quite well known
+\citep{armi2007,mayr2008,armi2011}, however, foehn cannot measured directly.
+Historically, foehn was often diagnosed manually by looking at available
+meteorological records such as wind speed and direction, time series of
+temperature and humidity, the conditions of the sky (e.g., occurrence of
+lenticularis clouds in the region), and observations upstream and downstream of
+the location of interest.  Some of the longest records of foehn observations
+can be found in Switzerland (Altdorf \note{i'm just guessing!}) reaching back
+to 1850 (\note{just guessing!}) for a hand full of stations.  These records are
+typically based on manual classification made by experienced weather observers,
+weather enthusiasts, and meteorologists.  This task is very time consuming and
+is only done for some selected sites, typically sites with long historical
+records. To be able to extend the diagnosis and to perform foehn classification
+on a large number of meteorological sites, computer aimed classification is
+needed.
+
+Few attempts to automatically classify foehn winds have been proposed over the
+past years such as \note{cite this and that}. These methods are mostly binary
+(either \textit{yes} or \textit{no}) and require manual tuning to specify the
+hyper-parameters, e.g., thresholds used to classify a specific observation as
+foehn. \cite{plavcan2014} proposed a new automated probabilistic foehn
+classification method based on finite mixture models, a statistical
+regression-based model for automatic foehn classification. In a recent study of
+\cite{mayr2018unpublished} this new method has been tested against existing
+classification methods and manual foehn classification made by students and
+experts to proof the diagnostic power of the new approach.
+
+Overall, the automatic finite mixture model based foehn classification
+\citep{plavcan2014} is able to compete with the human classification
+\citep{mayr2018unpublished} and has the advantages that it is (i) reproducible,
+and (ii) can easily be applied to large volumes of data. This allows one to
+efficiently perform foehn classification for a wide range of different sites.
+
+The article by \citep{plavcan2014} introduced the method using a powerful
+package called \code{FlexMix} \citep{leisch2004,gruen2007,gruen2008} available
+for the statistical software \textit{R}.  However, only little guidance was
+given by \citep{plavcan2014} on how to apply this method to own data and an
+implementation of the method outside the \textit{R} programming language was
+not available until today.  Thus, the aim of this article is to introduce the
+algorithm behind the automated probabilistic foehn classification
+\citep{plavcan2014} in more detail and introducing new open-source software for
+\textit{R} and \textit{Python} which provides the required methods to perform
+the classification.  Furthermore, the software comes with a set of examples and
+diagnostic tools for this specific task which allows everyone to use the method
+on their own.
+
+Section~\ref{sec:methods} describes the method itself followed by
+Section~\ref{sec:software}  introducing the software with a set of
+examples. Section~\dots \note{case study? Required for this article?}
 

### 2018-12-12 R/foehnix.R
Add binning now ...

@@ -11,7 +11,7 @@
 # -------------------------------------------------------------------
 # - EDITORIAL:   2018-11-28, RS: Created file on thinkreto.
 # -------------------------------------------------------------------
-# - L@ST MODIFIED: 2018-12-12 09:16 on marvin
+# - L@ST MODIFIED: 2018-12-12 12:35 on marvin
 # -------------------------------------------------------------------
 
 # -------------------------------------------------------------------
@@ -116,29 +116,6 @@ foehnix <- function(formula, data, windsector = NULL, maxit = 100L, tol = 1e-8, 
     # Attach to current environment to be used in the EM algorithm.
     for ( n in names(x) ) eval(parse(text = sprintf("%1$s <- x$%1$s", n))); rm(x)
 
-
-    #TODO: Should be moved into the prepare function.
-    # If lambda.min is not loglik: ridge penalization
-    if ( lambda.min %in% c("AIC", "BIC") ) {
-        # If nlambda is not a positive integer: stop.
-        stopifnot(inherits(nlambda, c("integer", "numeric")))
-        stopifnot(nlambda >= 0)
-        # Find large lambda where all parameters are close to 0.
-        # Trying lambdas between exp(6) and exp(12).
-        lambdas <- exp(seq(6, 12, by = 2))
-        # Fitting logistic regression models with different lambdas.
-        coef_sum_fun <- function(lambda, logitX, post, maxit, tol) {
-            # As response a first guess y >= median(y) is used.
-            m <- iwls_logit(logitX, as.numeric(y >= median(y)), standardize = FALSE,
-                            lambda = lambda, maxit = maxit, tol = tol)
-            sum(abs(m$beta[which(!grepl("^\\(Intercept\\)$", rownames(m$beta))),]))
-        }
-        x <- sapply(lambdas, coef_sum_fun, logitX = logitX, post = post, maxit = maxit, tol = tol)
-        # Pick the lambda where sum of parameters is smaller than a 
-        # certain threshold OR take maximum of lambdas tested.
-        lambdas <- exp(seq(min(which(x < 0.1), length(x)), -8, length = as.numeric(nlambda)))
-    } else { lambdas <- NULL }
-
     # Start optimization using weighted empirical moments for
     # location and scale of the two Gaussian distributions plus
     # an IWLS solver for the logistic regression model given the
[…]

### 2018-12-12 R/foehnix_functions.R

@@ -11,7 +11,7 @@
 # -------------------------------------------------------------------
 # - EDITORIAL:   2018-11-28, RS: Created file on thinkreto.
 # -------------------------------------------------------------------
-# - L@ST MODIFIED: 2018-12-12 10:24 on marvin
+# - L@ST MODIFIED: 2018-12-12 14:35 on marvin
 # -------------------------------------------------------------------
 
 # -------------------------------------------------------------------
@@ -122,7 +122,9 @@ foehnix_prepare <- function(formula, data, windsector, maxit, tol, lambda.min,
     # as.numeric(y > median(y)) is the initial best guess component assignment.
     # Force standardize = FALSE as logitX is already standardized if input
     # standardize = TRUE for the main function foehnix and advanced_foehnix.
-    ccmodel <- iwls_logit(logitX, as.numeric(y > median(y)), standardize = FALSE,
+    post <- do.call(sprintf("foehnix_%s_posterior", family),
+                    list(y = y, prob = rep(.5, length(y)), theta = theta))
+    ccmodel <- iwls_logit(logitX, post, standardize = FALSE,
                           maxit = tail(maxit, 1), tol = tail(tol, 1), nlambda = nlambda)
     cat("\nInitial parameters:\n")
     print(matrix(c(theta$mu1, exp(theta$logsd1), theta$mu2, exp(theta$logsd2)), ncol = 2,
@@ -199,6 +201,38 @@ destandardize_coefficients <- function(beta, X) {
     return(beta)
 }
 
+# -------------------------------------------------------------------
+# Trying to estimate a reasonably high upper bound for lambda
+# if ridge penalization is needed.
+# -------------------------------------------------------------------
+get_lambdas <- function(nlambda, logitX, post, maxit, tol) {
+
+    # If nlambda is not a positive integer: stop.
+    stopifnot(inherits(nlambda, c("integer", "numeric")))
+    stopifnot(nlambda >= 0)
+
+    # Fitting logistic regression models with different lambdas.
+    csum_fun <- function(lambda, logitX, post, maxit, tol) {
+        # As response a first guess y >= median(y) is used.
+        # Force standardize = FALSE as logitX is already standardized if
[…]

### 2018-12-19 R/families.R
Updated modular, moving to master now

@@ -7,7 +7,7 @@
 # -------------------------------------------------------------------
 # - EDITORIAL:   2018-12-13, RS: Created file on thinkreto.
 # -------------------------------------------------------------------
-# - L@ST MODIFIED: 2018-12-18 10:38 on marvin
+# - L@ST MODIFIED: 2018-12-19 08:30 on marvin
 # -------------------------------------------------------------------
 
 
@@ -134,7 +134,7 @@ foehnix_clogistic <- function(left = -Inf, right = Inf) {
                   package = "foehnix")
         },
         # Random sample function for two-component censored logistic distribution
-        r = function(n, mu, sigma) {
+        r = function(n, mu, sigma, lower.tail = TRUE) {
             stopifnot(length(mu) == 2)
             stopifnot(length(sigma) == 2)
             if ( length(n) == 1 ) n <- c(floor(n/2), ceiling(n/2))

### 2018-12-19 R/foehnix.R

@@ -11,7 +11,7 @@
 # -------------------------------------------------------------------
 # - EDITORIAL:   2018-11-28, RS: Created file on thinkreto.
 # -------------------------------------------------------------------
-# - L@ST MODIFIED: 2018-12-18 10:45 on marvin
+# - L@ST MODIFIED: 2018-12-19 15:53 on marvin
 # -------------------------------------------------------------------
 
 
@@ -79,7 +79,7 @@ foehnix.noconcomitant.fit <- function(y, family,
         # specified tolerance we assume that the algorithm
         # converged: stop iteration.
         if( is.na(llpath[[iter]]$full)) {
-            print("kaputte likelihood")
+            print("Likelihood got NA! Start environment browser for debugging ...")
             browser()
         }
 
@@ -300,7 +300,8 @@ print.foehnix.control <- function(x, ...) str(x)
 #       destandardize function should technically be ready to support
 #       this.
 # -------------------------------------------------------------------
-foehnix <- function(formula, data, windfilter = NULL, family = "gaussian",
+foehnix <- function(formula, data, switch = FALSE, windfilter = NULL,
+                    family = "gaussian",
                     control = foehnix.control(family, ...), ...) { 
 
     # Start timing (execution time of foehnix)
@@ -347,7 +348,7 @@ foehnix <- function(formula, data, windfilter = NULL, family = "gaussian",
     # If a wind sector is given: identify observations with a wind direction
     # outside the user defined sector. These will not be considered in the
     # statistical models.
-    idx_wind <- windfilter_get_indizes(data, windfilter)
+    idx_wind <- foehnix_filter(data, windfilter)
 
     # Take all elements which are not NA and are within the
     # defined wind sectors (if wind filter specified).
@@ -358,6 +359,7 @@ foehnix <- function(formula, data, windfilter = NULL, family = "gaussian",
     # all valid rows (without missing values on the mandatory columns
     # and, if a wind sector is given, with valid wind direction observations).
     mf <- matrix(unlist(mf[idx_take,]), ncol = ncol(mf), dimnames = list(NULL, names(mf)))
+    y  <- y[idx_take]
 
     # Check whether regularization is preferred over unpenalized
 #  regression estimation (only if lambda.min is "auto")
[…]

### 2018-12-19 R/foehnix_filter.R

@@ -1,0 +1,117 @@
+# -------------------------------------------------------------------
+# - NAME:        foehnix_filter.R
+# - AUTHOR:      Reto Stauffer
+# - DATE:        2018-12-19
+# -------------------------------------------------------------------
+# - DESCRIPTION:
+# -------------------------------------------------------------------
+# - EDITORIAL:   2018-12-19, RS: Created file on thinkreto.
+# -------------------------------------------------------------------
+# - L@ST MODIFIED: 2018-12-19 15:15 on marvin
+# -------------------------------------------------------------------
+
+apply_foehnix_filter <- function(x, filter, name) {
+
+    # Length of input data set
+    N <- nrow(x)
+
+    # If filter has to be applied to one specific column: subset
+    if ( ! is.null(name) ) x <- x[,name]
+
+    # Apply filter
+    res <- filter(x)
+
+    # Else check if we got what we expect
+    if ( inherits(res, "zoo") ) {
+        if ( ! inherits(coredata(res), "logical") )
+            stop("Custom filter function returns unexpected results.")
+    } else if ( ! inherits(coredata(res), "logical") ) {
+        stop("Custom filter function returns unexpected results.")
+    }
+    if ( length(res) != N )
+        stop(sprintf("Custom filter function returned result of wrong length (N != %d)\n", N))
+
+    # Contains NA values?
+    if ( any(is.na(res)) )
+        stop("Custom filter function returned NA values. Invalid filter.")
+
+    # Else return
+    return(res)
+
+}
+
+# -------------------------------------------------------------------
+# Returns the indizes of all rows in x where the wind direction
+# lies within the wind sector. Used to filter for wind directions.
+# -------------------------------------------------------------------
+foehnix_filter <- function(x, filter) {
+
+    # Wrong input 'x' 
+    if ( ! inherits(x, c("zoo", "data.frame")) )
+        stop("Input \"x\" to foehnix_filter has to be of class zoo or data.frame.")
+
+    # ---------------
+    # If NULL: return NULL
+    if ( is.null(filter) ) return(NULL)
+
+    # ---------------
+    # If windfilter is a function: execute function
+    # and check if the return is what we have expected, a list
+    # of integers in the range 1:nrow(x). If not, raise an error.
+    if ( is.function(filter) )
+        return(as.integer(which(apply_foehnix_filter(x, filter, NULL))))
+
[…]

### 2018-12-19 index.Rmd

### Toolbox for Automated Foehn Classification based on Mixture Models

The **foehnix** package provides a toolbox for automated probabilistic foehn
wind classification based on two-component mixture models (**foehn**
m**ix**ture models).  This method has first been proposed by [Plavcan et al.
(2015)](#plavcan2014) and compared to another semi-automatic classification,
foehn experts, students, and weather enthusiasts in the "Community Foehn
Classification Experiment" ([Mayr 2019](mayr2019)).

Foehn mixture models are a special case of the general flexible mixture model
class ([Fraley 2002](#fraley2000), [Leisch 2004](#leisch2004), [Gr&uuml;n
2007](#gruen2007), [Gr&uuml;n 2008](#gruen2008)), an unsupervised statistical
model to identify unobserveable clusters or components in data sets.
**foehnix** allows to estimate two-component mixture
models with additional concomitants.

Some of the features:

* Gaussian or logistic components with optional censoring or truncation.
* Concomitant variables for the probability model.
* Model assessment based on graphical output and information criteria.
* Automatic handling of missing values in the data set.

```{r, echo = FALSE, fig = TRUE, results = 'hide', fig.width = 12, fig.height = 8}
suppressPackageStartupMessages(library("foehnix"))
# Loading observations (data.frame), convert to zoo
data("ellboegen", package = "foehnix")
data("sattelberg", package = "foehnix")
timefun    <- function(x) as.POSIXct(x,  origin = "1970-01-01", tz = "UTC")
ellboegen  <- zoo(ellboegen[,-1],  timefun(ellboegen$timestamp))
sattelberg <- zoo(sattelberg[,-1], timefun(sattelberg$timestamp))
names(sattelberg) <- paste0("crest_", names(sattelberg))
data <- merge(ellboegen, sattelberg)
data$diff_t <- data$crest_t + 10.27 - data$t
windfilter <- list(dd = c(43, 223), crest_dd = c(90, 270))
mod <- foehnix(diff_t ~ ff, data = data, windfilter = windfilter, switch = TRUE)
# Plot
tsplot(mod, start = "2010-02-10", end = "2010-02-20")
```

### References

<p id="mayr2019">
Mayr GJ, Plavcan D, Laurence A, Elvidge A, Grisogono B, Horvath K, Jackson P, Neururer A, Seibert P, Steenburgh JW, Stiperski I, Sturman A, Ve&#269;enaj &#381;, Vergeiner J, Vosper S, Z&auml;ngl G (2019).
The Community Foehn Classification Experiment.
_Bulletin of the American Meteorological Society_, **NA**, NA&mdash;NA,
<a href="https://doi.org/10.1175/BAMS-D-17-0200.1" target="_blank">10.1175/BAMS-D-17-0200.1</a>
</p>

<p id="plavcan2014">
Plavcan D, Mayr GJ, Zeileis A (2014).
Automatic and Probabilistic Foehn Diagnosis with a Statistical Mixture Model.
_Journal of Applied Meteorology and Climatology_, **53**(3), 652&mdash;659,
<a href="https://dx.doi.org/10.1175/JAMC-D-13-0267.1" target="_blank">10.1175/JAMC-D-13-0267.1</a>
</p>

<p id="gruen2008">
Gr&uuml;n B, Friedrich L (2008).
FlexMix Version 2: Finite Mixtures with Concomitant Variables and Varying and Constant Parameters.
_Journal of Statistical Software, Articles_, **28**(4), 1&mdash;35,
doi:<a href="https://dx.doi.org/10.18637/jss.v028.i04" target="_blank">10.18637/jss.v028.i04</a>
</p>

<p id="gruen2007">
Gr\"un B, Leisch F (2007).
Fitting Finite Mixtures of Generalized Linear Regressions in _R_.
_Computational Statistics \& Data Analysis_, **51**(11),
doi:<a href="https://dx.doi.org/10.1016/j.csda.2006.08.014" target="_blank">10.1016/j.csda.2006.08.014</a>
</p>

<p id="leisch2004">
Friedrich L (2004).
FlexMix: A General Framework for Finite Mixture Models and Latent Class Regression in _R_.
_Journal of Statistical Software, Articles_, **11**(8), 1&mdash;18,
doi:<a href="https://dx.doi.org/10.18637/jss.v011.i08" target="_blank">10.18637/jss.v011.i08</a>
</p>

<p id="fraley2000">
Fraley C, Raftery AE (2000).
Model-Based Clustering, Discriminant Analysis, and Density Estimation.
_Journal of the American Statistical Association_, **97**(458), 611&mdash;631,
doi:<a href="https://dx.doi.org/10.1198/016214502760047131" target="_blank">10.1198/016214502760047131</a>
</p>

### 2018-12-19 R/iwls_logit.R
Pushing updated docs, README

@@ -11,7 +11,7 @@
 # -------------------------------------------------------------------
 # - EDITORIAL:   2018-11-28, RS: Created file on thinkreto.
 # -------------------------------------------------------------------
-# - L@ST MODIFIED: 2018-12-16 16:57 on marvin
+# - L@ST MODIFIED: 2018-12-19 18:36 on marvin
 # -------------------------------------------------------------------
 
 # -------------------------------------------------------------------
@@ -19,6 +19,10 @@
 # -------------------------------------------------------------------
 iwls_logit <- function(X, y, beta = NULL, lambda = NULL, standardize = TRUE,
                        maxit = 100L, tol = 1e-8, verbose = FALSE, ...) {
+
+    # NAns?
+    if ( any(is.na(X)) ) stop("Input \"X\" contains NA values. Stop.")
+    if ( any(is.na(y)) ) stop("Input \"y\" contains NA values. Stop.")
 
     # Checking inputs. Constant covariates (concomitant variables)
     # are not allowed except one column (intercept).

### 2018-12-20 index.Rmd
update

@@
 foehn experts, students, and weather enthusiasts in the "Community Foehn
-Classification Experiment" ([Mayr 2019](mayr2019)).
+Classification Experiment" ([Mayr 2019](mayr2018)).
 
@@
 
-<p id="mayr2019">
-Mayr GJ, Plavcan D, Laurence A, Elvidge A, Grisogono B, Horvath K, Jackson P, Neururer A, Seibert P, Steenburgh JW, Stiperski I, Sturman A, Ve&#269;enaj &#381;, Vergeiner J, Vosper S, Z&auml;ngl G (2019).
+<p id="mayr2018">
+Mayr GJ, Plavcan D, Laurence A, Elvidge A, Grisogono B, Horvath K, Jackson P, Neururer A, Seibert P, Steenburgh JW, Stiperski I, Sturman A, Ve&#269;enaj &#381;, Vergeiner J, Vosper S, Z&auml;ngl G (2018).
 The Community Foehn Classification Experiment.
-_Bulletin of the American Meteorological Society_, **NA**, NA&mdash;NA,
+_Bulletin of the American Meteorological Society_, **99**(11), 2229&mdash;2235,
 <a href="https://doi.org/10.1175/BAMS-D-17-0200.1" target="_blank">10.1175/BAMS-D-17-0200.1</a>

### 2018-12-20 R/advanced_foehnix.R

@@ -1,313 +1,0 @@
-# -------------------------------------------------------------------
-# - NAME:        advanced_foehnix.R
-# - AUTHOR:      Reto Stauffer
-# - DATE:        2018-11-28
-# -------------------------------------------------------------------
-# - DESCRIPTION: Code to estimate Gaussian and logistic two-component
-#                mixture models with optional truncation and
-#                censoring. Gradient based optimization based on
-#                BFGS optimizer.
-# -------------------------------------------------------------------
-# - EDITORIAL:   2018-11-28, RS: Created file on thinkreto.
-# -------------------------------------------------------------------
-# - L@ST MODIFIED: 2018-12-12 17:07 on marvin
-# -------------------------------------------------------------------
-
-
-
-#### -------------------------------------------------------------------
-#### IWLS optimizer for logistic regression model (concomitant model)
-#### -------------------------------------------------------------------
-###bfgs_logit <- function(X, y, beta = NULL, lambda = NULL, standardize = TRUE,
-###                       maxit = 100L, tol = 1e-8, ...) {
-###
-###    # Checking inputs. Constant covariates (concomitant variables)
-###    # are not allowed except one column (intercept).
-###    if(sum(apply(X, 2, sd) == 0) > 1) stop("Multiple columns with constant values!")
-###    if ( min(y) < 0 | max(y) > 1 ) stop("y values out of range. Have to be within ]0,1[.")
-###
-###    # Standardize design matrix?
-###    if ( standardize ) X <- standardize_model_matrix(X)
-###
-###    # initialize regression coefficients if needed
-###    if ( is.null(beta) ) beta <- rep.int(0, ncol(X)) # FIXME: there is surely a better solution!
-###
-###    # Calculate logistic regression likelihood
-###    loglik <- function(beta, post, X, lambda) {
-###        eta <- drop(X %*% beta) 
-###        # Penalize all non-intercept coefficients
-###        idx <- which(!grepl("^\\(Intercept\\)$", colnames(X)))
-###        if ( is.null(lambda) | length(idx) == 0 ) return(sum(y * eta - log(1 + exp(eta))))
-###        # Penalized log-likelihood sum
-###        sum(y * eta - log(1 + exp(eta))) - lambda / 2 * sum(beta[idx]^2)
-###    }
-###    # Gradient for the 
[…]

### 2018-12-20 R/families.R

@@ -7,7 +7,7 @@
 # -------------------------------------------------------------------
 # - EDITORIAL:   2018-12-13, RS: Created file on thinkreto.
 # -------------------------------------------------------------------
-# - L@ST MODIFIED: 2018-12-19 08:30 on marvin
+# - L@ST MODIFIED: 2018-12-20 09:53 on marvin
 # -------------------------------------------------------------------
 
 
@@ -68,14 +68,11 @@ foehnix_logistic <- function() {
             # Calculate/trace loglik
             eps  <- sqrt(.Machine$double.eps)
             prob <- pmax(eps, pmin(1-eps, prob))
-            # plog function and eps are used to void log(0)
-            plog <- function(x, eps) ifelse(x < eps, log(eps), log(x))
-            eps <- sqrt(.Machine$double.eps)
             # Calculate log-likelihood
             ll <- data.frame(
                     component = sum(post       * dlogis(y, theta$mu2, exp(theta$logsd2), log = TRUE))
                               + sum((1 - post) * dlogis(y, theta$mu1, exp(theta$logsd1), log = TRUE)),
-                    concomitant = sum((1 - post) * plog(1 - prob, eps) + post * plog(prob, eps))
+                    concomitant = sum((1 - post) * log(1 - prob) + post * log(prob))
                   )
             ll$full <- sum(unlist(ll))
             return(ll)
@@ -340,14 +337,11 @@ foehnix_gaussian <- function() {
             # Calculate/trace loglik
             eps  <- sqrt(.Machine$double.eps)
             prob <- pmax(eps, pmin(1-eps, prob))
-            # plog function and eps are used to void log(0)
-            plog <- function(x, eps) ifelse(x < eps, log(eps), log(x))
-            eps <- sqrt(.Machine$double.eps)
             # Calculate log-likelihood
             ll <- data.frame(
                     component = sum(post       * dnorm(y, theta$mu2, exp(theta$logsd2), log = TRUE))
                               + sum((1 - post) * dnorm(y, theta$mu1, exp(theta$logsd1), log = TRUE)),
-                    concomitant = sum((1 - post) * plog(1 - prob, eps) + post * plog(prob, eps))
+                    concomitant = sum((1 - post) * log(1 - prob) + post * log(prob))
                   )
             ll$full <- sum(unlist(ll))
             return(ll)

### 2018-12-20 R/foehnix.R

@@ -11,7 +11,7 @@
 # -------------------------------------------------------------------
 # - EDITORIAL:   2018-11-28, RS: Created file on thinkreto.
 # -------------------------------------------------------------------
-# - L@ST MODIFIED: 2018-12-19 18:44 on marvin
+# - L@ST MODIFIED: 2018-12-20 11:32 on marvin
 # -------------------------------------------------------------------
 
 
@@ -411,6 +411,7 @@ foehnix <- function(formula, data, switch = FALSE, windfilter = NULL,
 
     # If inversion has been requested: switch coefficients
     if ( switch ) {
+        print(rval$theta)
         rval$ccmodel$coef <- -rval$ccmodel$coef
         rval$ccmodel$beta <- -rval$ccmodel$beta
         rval$coefpath     <- -rval$coefpath
@@ -421,6 +422,7 @@ foehnix <- function(formula, data, switch = FALSE, windfilter = NULL,
         rval$post <- 1 - rval$post
         # Invert coefficients if switch = TRUE
         coef <- -coef
+        print(rval$theta)
     }
 
     # Create the return list object (foehnix object)
@@ -484,7 +486,7 @@ foehnix <- function(formula, data, switch = FALSE, windfilter = NULL,
     return(res)
 }
 
-predict.foehnix <- function(x, newdata = NULL, type = "response") {
+predict.foehnix <- function(object, newdata = NULL, type = "response", ...) {
 
     # Allowed input types
     type <- match.arg(type, c("response", "all"))
@@ -495,21 +497,21 @@ predict.foehnix <- function(x, newdata = NULL, type = "response") {
 
     # If no newdata is provided: take the data set on
     # which the model has been estimated.
-    if ( is.null(newdata) ) newdata <- x$data
+    if ( is.null(newdata) ) newdata <- object$data
 
     # Probability model
-    if ( is.null(x$coef$concomitants) ) {
-        prob <- mean(x$optimizer$prob)
+    if ( is.null(object$coef$concomitants) ) {
+        prob <- mean(object$optimizer$prob)
     } else {
-        logitX <- model.matrix(x$formula, newdata)
-        prob   <- plogis(drop(logitX %*% x$coef$concomitants))
+        logitX <- model.matrix(object$formula, newdata)
+        prob   <- plogis(drop(logitX %*% object$coef$concomitants))
     }
 
     # Calculate density
-    y    <- model.response(model.frame(x$formula, newdata))
[…]

### 2018-12-20 R/foehnix_filter.R

@@ -7,7 +7,7 @@
 # -------------------------------------------------------------------
 # - EDITORIAL:   2018-12-19, RS: Created file on thinkreto.
 # -------------------------------------------------------------------
-# - L@ST MODIFIED: 2018-12-19 15:15 on marvin
+# - L@ST MODIFIED: 2018-12-19 19:55 on marvin
 # -------------------------------------------------------------------
 
 apply_foehnix_filter <- function(x, filter, name) {
@@ -91,8 +91,6 @@ foehnix_filter <- function(x, filter) {
     # the wind filter rule(s).
     fun <- function(name, filter, x) {
         # Picking data and wind filter rule
-        print(name)
-        print(filter[[name]])
         if ( is.function(filter[[name]]) ) {
             apply_foehnix_filter(x, filter[[name]], name)
         } else {

### 2018-12-20 R/foehnix_functions.R

@@ -9,7 +9,7 @@
 # -------------------------------------------------------------------
 # - EDITORIAL:   2018-12-16, RS: Created file on thinkreto.
 # -------------------------------------------------------------------
-# - L@ST MODIFIED: 2018-12-18 19:30 on marvin
+# - L@ST MODIFIED: 2018-12-20 11:37 on marvin
 # -------------------------------------------------------------------
 
 
@@ -73,43 +73,43 @@ destandardize_coefficients <- function(beta, X) {
 # Information criteria: logLik, AIC, BIC, and effective degrees of
 # freedom
 # -------------------------------------------------------------------
-logLik.foehnix <- function(x, ...)  structure(x$optimizer$loglik, names = "loglik")
-AIC.foehnix <- function(x, ...)     structure(x$optimizer$AIC, names = "AIC")
-BIC.foehnix <- function(x, ...)     structure(x$optimizer$BIC, names = "BIC")
-edf <- function(x, ...) UseMethod("edf")
-edf.foehnix <- function(x, ...)     structure(x$optimizer$edf, names = "edf")
+logLik.foehnix <- function(object, ...) structure(object$optimizer$loglik, names = "loglik")
+AIC.foehnix <- function(object, ...)    structure(object$optimizer$AIC, names = "AIC")
+BIC.foehnix <- function(object, ...)    structure(object$optimizer$BIC, names = "BIC")
+edf <- function(object, ...) UseMethod("edf")
+edf.foehnix <- function(object, ...)     structure(object$optimizer$edf, names = "edf")
 
 # -------------------------------------------------------------------
 # Print foehnix model object.
 # TODO: better default print method required?
 # -------------------------------------------------------------------
-print.foehnix <- function(x, ...) print(summary(x))
+print.foehnix <- function(object, ...) print(summary(object, ...))
 
 # -------------------------------------------------------------------
 # Estimated regression coefficients
 # -------------------------------------------------------------------
-coef.foehnix <- function(x, type = "parameter", ...) {
+coef.foehnix <- function(object, type = "parameter", ...) {
 
     # One of the two types: parameter (destandardized if required),
     # or coefficient (standardized coefficients if standardized was TRUE).
     type <- match.arg(type, c("parameter", "coefficient"))
     if ( type == "parameter" ) {
[…]

### 2018-12-20 R/shiny.R

@@ -127,7 +127,7 @@ shiny_check <- function(x, ...) {
         closest <- index(x$prob)[which.min(abs(as.numeric(index(x$prob)) - e$x))]
         showNotification(sprintf("Closest is %s", as.character(closest)))
     }
-    library("shiny")
+    requireNamespace("shiny")
     # User interface
     ui <- basicPage(
         tags$head(

### 2018-12-20 R/families.R
Updated docs

@@ -7,7 +7,7 @@
 # -------------------------------------------------------------------
 # - EDITORIAL:   2018-12-13, RS: Created file on thinkreto.
 # -------------------------------------------------------------------
-# - L@ST MODIFIED: 2018-12-20 09:53 on marvin
+# - L@ST MODIFIED: 2018-12-20 13:13 on marvin
 # -------------------------------------------------------------------
 
 
@@ -181,10 +181,10 @@ foehnix_clogistic <- function(left = -Inf, right = Inf) {
                 theta$mu1 <- sum(y * (1 - post)) / sum(1 - post)
                 theta$mu2 <- sum(y * post) / sum(post)
                 if ( init ) {
-                    theta$logsd1 <- theta$logsd2 <- log(sd(y))
+                    theta$logsd1 <- theta$logsd2 <- log(sd(y) * sqrt(3) / pi)
                 } else {
-                    theta$logsd1 <- log(sqrt(sum((y-mu1)^2 * (1-post)) / sum(1-post)))
-                    theta$logsd2 <- log(sqrt(sum((y-mu2)^2 * (post)) / sum(post)))
+                    theta$logsd1 <- log(sqrt(sum((y-mu1)^2 * (1-post)) / sum(1-post))* sqrt(3) / pi)
+                    theta$logsd2 <- log(sqrt(sum((y-mu2)^2 * (post))   / sum(post)) * sqrt(3) / pi)
                 }
             }
 
@@ -288,10 +288,10 @@ foehnix_tlogistic <- function(left = -Inf, right = Inf) {
                 theta$mu1 <- sum(y * (1 - post)) / sum(1 - post)
                 theta$mu2 <- sum(y * post) / sum(post)
                 if ( init ) {
-                    theta$logsd1 <- theta$logsd2 <- log(sd(y))
+                    theta$logsd1 <- theta$logsd2 <- log(sd(y) * sqrt(3) / pi)
                 } else {
-                    theta$logsd1 <- log(sqrt(sum((y-mu1)^2 * (1-post)) / sum(1-post)))
-                    theta$logsd2 <- log(sqrt(sum((y-mu2)^2 * (post)) / sum(post)))
+                    theta$logsd1 <- log(sqrt(sum((y-mu1)^2 * (1-post)) / sum(1-post))* sqrt(3) / pi)
+                    theta$logsd2 <- log(sqrt(sum((y-mu2)^2 * (post))   / sum(post)) * sqrt(3) / pi)
                 }
             }
 

### 2018-12-20 R/datasets.R
Roxygen!

@@ -1,0 +1,87 @@
+
+#' Meteorological Observation Data Station Ellboegen
+#'
+#' The \code{\link{foehnix}} package comes with two data sets of meteorological
+#' observations for two sites in Tyrol, Austria.  One station is located nearby
+#' Ellboegen, a small town close to Innsbruck, the second station (Sattelberg)
+#' is located south of Ellboegen close to the border between Austria and Italy
+#' on top of a mountain close to the crest of the European Alps and is used as
+#' 'crest station' (see \code{\link{sattelberg}}).
+#'
+#' @usage data("ellboegen")
+#'
+#' @source Station operated by the Department of Atmospheric and Cryospheric
+#' Sciences (\url{http://acinn.uibk.ac.at}) of the University of Innsbruck.
+#' The data is available under the
+#' \href{https://creativecommons.org/licenses/by-sa/4.0/}{CC BY-SA 4.0 license}.
+#'
+#' @details
+#' Ellboegen:
+#' \itemize{
+#'     \item Location: 11.42889 E/47.18694 N
+#'     \item Altitude: 1080 meters above mean sea level
+#' }
+#' 
+#' Variable description (columns):
+#' \itemize{
+#'    \item \code{timestamp}: UNIX time stamp, seconds since 1970-01-01 00:00:00
+#'    \item \code{dd}: wind direction in degrees, meteorological format
+#'        (0/360 correspond to 'wind from north', 90 to 'wind from east',
+#'        180 'wind from south', and 270 'wind from west'
+#'    \item \code{ff}: wind speed in meters per second
+#'    \item \code{p}: station pressure in hectopascal
+#'    \item \code{rh}: relative humidity in percent
+#'    \item \code{t}: dry air temperature in degrees Celsius
+#' }
+#'
+#' @seealso \code{\link{sattelberg}}.
+#'
+#' @name ellboegen
+#' @docType data
+#' @keywords data ellboegen
+#' @author Reto Stauffer
+NULL
+
+#' Meteorological Observation Data Station Sattelberg
+#'
+#' The \code{\link{foehnix}} package comes with two data sets of meteorological
+#' observations for two sites in Tyrol, Austria.  One station is located nearby
+#' Ellboegen (see \code{\link{ellboegen}}), a small town close to Innsbruck,
+#' the second station (Sattelberg) is located south of Ellboegen close to the
[…]

### 2018-12-20 R/foehnix.R

@@ -11,21 +11,40 @@
 # -------------------------------------------------------------------
 # - EDITORIAL:   2018-11-28, RS: Created file on thinkreto.
 # -------------------------------------------------------------------
-# - L@ST MODIFIED: 2018-12-20 11:32 on marvin
+# - L@ST MODIFIED: 2018-12-20 18:33 on marvin
 # -------------------------------------------------------------------
 
 
 
-
-# -------------------------------------------------------------------
-# -------------------------------------------------------------------
-# -------------------------------------------------------------------
-# -------------------------------------------------------------------
-# -------------------------------------------------------------------
-# -------------------------------------------------------------------
-# -------------------------------------------------------------------
-# -------------------------------------------------------------------
-# -------------------------------------------------------------------
+#' Fitting foehnix Mixture Model Without Concomitant Model.
+#'
+#' Fitting method for two-component foehnix mixture model without
+#' concomitants.
+#' Typically not called directly, interfaced via \code{\link{foehnix}}.
+#' @param y numeric vector, covariate for the components of the
+#'        mixture model, dimension \code{N}.
+#' @param family object of class \code{\link{foehnix.family}}.
+#' @param maxit positive integer, or vector of length 2 with positive
+#'        integer values. Maximum number of iterations of the EM algorithm.
+#'        Check manual of \code{\link{foehnix.control}} for more details.
+#' @param tol numeric, or vector of length 2 containing numeric values.
+#'        Tolerance for the EM algorithm.
+#'        Check manual of \code{\link{foehnix.control}} for more details.
+#' @param verbose logical, default \code{TRUE}. If set to \code{FALSE}
+#'        verbose output will be suppressed.
+#' @param \dots additional arguments, unused.
+#'
+#' @seealso
+#' \code{\link{foehnix.family}},
+#' \code{\link{foehnix}},
+#' \code{\link{foehnix.control}},
+#' \code{\link{foehnix.noconcomitant.fit}},
+#' \code{\link{foehnix.reg.fit}},
+#' \code{\link{iwls_logit}}.
+#'
+#' @author Reto Stauffer
+#' @import stats
+#' @import utils
 foehnix.noconcomitant.fit <- function(y, family,
                     maxit = 100L, tol = 1e-5, verbose = TRUE, ...) {
 
@@ -117,15 +136,40 @@ if ( inherits(y, "binned") ) stop("Stop, requires changes on computation of BIC!
 }
 
 
-# -------------------------------------------------------------------
[…]

### 2018-12-20 R/foehnix_filter.R

@@ -7,43 +7,165 @@
 # -------------------------------------------------------------------
 # - EDITORIAL:   2018-12-19, RS: Created file on thinkreto.
 # -------------------------------------------------------------------
-# - L@ST MODIFIED: 2018-12-19 19:55 on marvin
+# - L@ST MODIFIED: 2018-12-20 16:29 on marvin
 # -------------------------------------------------------------------
 
-apply_foehnix_filter <- function(x, filter, name) {
 
-    # Length of input data set
-    N <- nrow(x)
-
-    # If filter has to be applied to one specific column: subset
-    if ( ! is.null(name) ) x <- x[,name]
-
-    # Apply filter
-    res <- filter(x)
-
-    # Else check if we got what we expect
-    if ( inherits(res, "zoo") ) {
-        if ( ! inherits(coredata(res), "logical") )
-            stop("Custom filter function returns unexpected results.")
-    } else if ( ! inherits(coredata(res), "logical") ) {
-        stop("Custom filter function returns unexpected results.")
-    }
-    if ( length(res) != N )
-        stop(sprintf("Custom filter function returned result of wrong length (N != %d)\n", N))
-
-    # Contains NA values?
-    if ( any(is.na(res)) )
-        stop("Custom filter function returned NA values. Invalid filter.")
-
-    # Else return
-    return(res)
-
-}
-
-# -------------------------------------------------------------------
-# Returns the indizes of all rows in x where the wind direction
-# lies within the wind sector. Used to filter for wind directions.
-# -------------------------------------------------------------------
+#' Evaluates Data Filter Rules for foehnix Mixture Moel Calls.
+#' 
+#' \code{\link{foehnix}} models allow to specify an optional
+#' \code{foehnix_filter}. If a filter is given only a subset
+#' of the data set provided to \code{\link{foehnix}} is used
+#' for the foehn classification.
+#' 
+#' A typical example is a wind direction filter such that
+#' only observations (times) are used where the observed
+#' wind direction was within a user defined wind sector
+#' corresponding to the wind direction during foehn events
+#' for a specific location.
+#' 
+#' However, the filter option allows to even implement complex
+#' filter rules if required. The 'Details' section contains
+#' further information and examples how this filter rules can
[…]

### 2018-12-20 R/foehnix_functions.R

@@ -9,24 +9,49 @@
 # -------------------------------------------------------------------
 # - EDITORIAL:   2018-12-16, RS: Created file on thinkreto.
 # -------------------------------------------------------------------
-# - L@ST MODIFIED: 2018-12-20 11:37 on marvin
+# - L@ST MODIFIED: 2018-12-20 18:34 on marvin
 # -------------------------------------------------------------------
 
 
-# -------------------------------------------------------------------
-# Small helper function to check whether or not the matrix is
-# standardized or not. Considered to be standardized if the matrix
-# contains the two additional attributes "scaled:center" and
-# "scaled:scale".
-# -------------------------------------------------------------------
-is.standardized <- function(x, ...) UseMethod("is.standardized")
-is.standardized.matrix <- function(x, ...) inherits(x, "standardized")
-
-
-# -------------------------------------------------------------------
-# Standardize (model) matrix
-# -------------------------------------------------------------------
+#' Standardize (Model) Matrix
+#'
+#' Function to standardize the columns of a matrix, used to
+#' standardize the model matrix before estimating the regression
+#' coefficients of a generalized linear model (\code{\link{iwls_logit}}).
+#'
+#' @param x matrix of dimension \code{N x p}.
+#' @param ... additional arguments, ignored.
+#' @return Returns a matrix of the same dimension as input \code{x}
+#' but with standardized data. The return object is of class
+#' \code{c("standardized", "matrix")} which comes with some handy
+#' S3 methods.
+#'
+#' @examples
+#' # Example data set
+#' data("airquality")
+#' airquality <- na.omit(airquality)
+#'
+#' # Create model matrix
+#' X <- model.matrix(Ozone ~ ., data = airquality)
+#' print(head(X))
+#'
+#' # Standardize
+#' S <- standardize(X)
+#' print(head(S))
+#'
+#' is.standardized(X)
+#' is.standardized(X)
+#'
+#' # Get parameters used for standardization
+#' center(S)
+#' scale(S)
+#'
+#' @author Reto Stauffer
+#' @export
 standardize <- function(x, ...) UseMethod("standardize")
+
+#' @rdname standardize
+#' @export
 standardize.matrix <- function(x, ...) {
     # Scale covariates
     scaled_center <- structure(rep(0, ncol(x)), names = colnames(x))
@@ -42,16 +67,53 @@ standardize.matrix <- function(x, ...) {
[…]

### 2018-12-20 R/iwls_logit.R

@@ -1,22 +1,117 @@
 # -------------------------------------------------------------------
-# - NAME:        foehnix.R
+# - NAME:        iwls_logit.R
 # - AUTHOR:      Reto Stauffer
 # - DATE:        2018-11-28
 # -------------------------------------------------------------------
-# - DESCRIPTION: Scripts to estimate Gaussian and logistic
-#                two-component mixture models with IWLS and
-#                weighted empirical moments.
-#                "simple" here referrs to non-gradient non-optimizer
-#                based model estimates.
+# - DESCRIPTION: Iterative weighted least squares solver for logistic
+#                regression models. Used to estimate the concomitant
+#                models in 'foehnix'.
 # -------------------------------------------------------------------
 # - EDITORIAL:   2018-11-28, RS: Created file on thinkreto.
 # -------------------------------------------------------------------
-# - L@ST MODIFIED: 2018-12-19 18:36 on marvin
+# - L@ST MODIFIED: 2018-12-20 17:46 on marvin
 # -------------------------------------------------------------------
 
-# -------------------------------------------------------------------
-# IWLS optimizer for logistic regression model (concomitant model)
-# -------------------------------------------------------------------
+#' IWLS Solver for Binary Logistic Regression Model
+#' 
+#' Iterative weighted least squares solver for a logistic regression
+#' model. Used by \code{\link{foehnix.unreg.fit}} to estimate the
+#' concomitant model of the two-component foehnix mixture models.
+#' 
+#' @param X model matrix including intercept (if required).
+#' @param y response, can be binary or probabilities (values in \code{]0,1[}).
+#' @param beta initial regression coefficients. If not set (\code{beta = NULL}, default),
+#'        all coefficients will be initialized with \code{0}.
+#' @param lambda if set to \code{NULL} (default) no penalty is used during optimization.
+#'        A \code{float} can be provided for regularization (ridge/L2 penalty).
+#' @param maxit integer, maximum number of iterations, default \code{100}.
+#' @param tol float, tolerance for the improvement to check for convergence.
+#' @param standardize logical. If set to \code{TRUE} (default) the model matrix
+#'        containing the concomitant variables will be standardized.
[…]

### 2018-12-20 R/shiny.R

@@ -1,7 +1,10 @@
 
 
+# TODO: Neither well coded, nor tested, nor documented.
+#' @import zoo
 shiny_tsplot <- function(output, x, start, end) {
 
+    requireNamespace("shiny")
     tmp <- merge(window(x$prob, start = start, end = end),
                  window(x$data, start = start, end = end))
     names(tmp)[1L] <- "prob"
@@ -77,7 +80,7 @@ shiny_tsplot <- function(output, x, start, end) {
         axis(side = 4, at = seq(20, 100, by = 20))
         mtext(side = 4, line = 3, "rel humidity")
     }
-    output$ts_trh <- renderPlot(ts_trh_plot(tmp), height = 300, width = 1400)
+    output$ts_trh <- shiny::renderPlot(ts_trh_plot(tmp), height = 300, width = 1400)
 
     # -----------------------------------------------------------
     ts_dd_plot <- function(x, ...) {
@@ -98,7 +101,7 @@ shiny_tsplot <- function(output, x, start, end) {
         axis(side = 4, at = seq(60, 300, by = 60)) 
         mtext(side = 4, line = 3, "wind dir")
     }
-    output$ts_dd <- renderPlot(ts_dd_plot(tmp), height = 300, width = 1400)
+    output$ts_dd <- shiny::renderPlot(ts_dd_plot(tmp), height = 300, width = 1400)
 
     # -----------------------------------------------------------
     ts_prob_plot <- function(x, ...) {
@@ -111,11 +114,22 @@ shiny_tsplot <- function(output, x, start, end) {
         abline(h = seq(25, 75, by = 25), lty = 5, col = "#CC3300")
         add_midnight_lines(x)
     }
-    output$ts_prob <- renderPlot(ts_prob_plot(tmp), height = 300, width = 1400)
+    output$ts_prob <- shiny::renderPlot(ts_prob_plot(tmp), height = 300, width = 1400)
 }
 
 
+#' Shiny Application to Check foehnix Classification (Time Series Plots)
+#'
+#' TODO: Neither well coded, nor tested, nor documented.
+#'
+#' @param x a \code{\link{foehnix}} mixture model object.
+#' @param ... additional arguments, not in use.
+#'
+#' @import zoo
+#' @author Reto Stauffer
 shiny_check <- function(x, ...) {
+
+    requireNamespace("shiny")
 
     # Click string
     xy_str <- function(e) {
[…]

### 2018-12-20 R/tsplot.R

@@ -3,23 +3,44 @@
 # - AUTHOR:      Reto Stauffer
 # - DATE:        2018-12-16
 # -------------------------------------------------------------------
-# - DESCRIPTION:
+# - DESCRIPTION: Time series plot of observations plus classification
+#                based on a foehnix mixture model.
 # -------------------------------------------------------------------
 # - EDITORIAL:   2018-12-16, RS: Created file on thinkreto.
 # -------------------------------------------------------------------
-# - L@ST MODIFIED: 2018-12-16 17:30 on marvin
+# - L@ST MODIFIED: 2018-12-20 18:09 on marvin
 # -------------------------------------------------------------------
 
+#' Time Series Plot of foehnix Models
+#' 
+#' Development time series plots of estimated \code{\link{foehnix}}
+#' models. TODO: they are very specific at the moment!
+#' 
+#' @param x object of type \code{foehnix}.
+#' @param start POSIXt object or an object which can be converted to POSIXt.
+#' @param end POSIXt object or an object which can be converted to POSIXt.
+#' @param ndays integer, number of days used when looping trough the time series.
+#' @param ... additional arguments, ignored.
+#' @param xtra optional zoo object with probabilities in \code{]0,1[} to compare
+#'         two classification algorithms.
+#' @param ask logical, default is \code{TRUE}.
+#' 
+#' @details Development method to access plausability of the estimated foehn
+#' proabilities.  For software release this method should either be removed or
+#' made much more general. At the moment the method heavily depends on the
+#' names of the data used as input for \code{\link{foehnix}}.
+#'
+#' TODO: this is very specific to our data and our variable names.
+#'       either provide something like this and force the users to
+#'       follow our naming conventions, or make it much more
+#'       flexible/generig.
+#'
+#' @author Reto Stauffer
+#' @export
+tsplot <- function(x, start = NULL, end = NULL, ndays = 10, ..., xtra = NULL, ask = TRUE) {
 
-# -------------------------------------------------------------------
-# Development time series plot routine.
[…]

### 2018-12-20 R/windrose.R

@@ -7,10 +7,28 @@
 # -------------------------------------------------------------------
 # - EDITORIAL:   2018-12-16, RS: Created file on thinkreto.
 # -------------------------------------------------------------------
-# - L@ST MODIFIED: 2018-12-19 21:54 on marvin
+# - L@ST MODIFIED: 2018-12-20 18:21 on marvin
 # -------------------------------------------------------------------
 
-windrose <- function(x, ...) UseMethod("windrose")
+#' Windrose Plot
+#'
+#' Plotting a windrose using meteorological wind direction and
+#' wind speed (or gust speed).
+#'
+#' @param x \code{zoo} object or \code{data.frame} or \code{\link{foehnix}}
+#'        mixture model object.
+#' @param which character, one of \code{"unconditional"}, \code{"nofoehn"}, 
+#'        \code{"foehn"}, or \code{"combined"}.
+#' @param ddvar character, name of the wind direction varaible in \code{x} or as used
+#'        for training the \code{\link{foehnix}} mixture model.
+#' @param ffvar character, name of the wind speed varaible in \code{x} or as used
+#'        for training the \code{\link{foehnix}} mixture model.
+#' @param ... additional arguments, currently unused.
+#'
+#' @details TODO: not yet finished and/or tested!
+windrose <- function(x, which = NULL, ddvar = "dd", ffvar = "ff", ...) UseMethod("windrose")
+
+#' @rdname windrose
 windrose.foehnix <- function(x, which = NULL, ddvar = "dd", ffvar = "ff", ...) {
 
     # First: checking argument "which"
@@ -68,7 +86,7 @@ windrose.foehnix <- function(x, which = NULL, ddvar = "dd", ffvar = "ff", ...) {
         stop("Data sets do not match, no overlapping data found for probabilities of foehn, dd, and ff.")
 
     # Final step: combine dd, ff, probabilities, and start plotting.
-    data <- na.omit(merge(dd, ff, x$prob))
+    data <- na.omit(merge(dd, ff, x$prob)); names(data)[ncol(data)] <- "prob"
     if ( nrow(data) == 0 )
         stop("No data left after combining \"dd\", \"ff\", and \"foehn probabilities\".")
 
[…]

### 2018-12-21 index.Rmd
Update to version 0.0-4

@@
 data$diff_t <- data$crest_t + 10.27 - data$t
-windfilter <- list(dd = c(43, 223), crest_dd = c(90, 270))
-mod <- foehnix(diff_t ~ ff, data = data, windfilter = windfilter, switch = TRUE)
+filter <- list(dd = c(43, 223), crest_dd = c(90, 270))
+mod <- foehnix(diff_t ~ ff, data = data, filter = filter, switch = TRUE)
 # Plot

### 2018-12-21 R/foehnix.R

@@ -11,7 +11,7 @@
 # -------------------------------------------------------------------
 # - EDITORIAL:   2018-11-28, RS: Created file on thinkreto.
 # -------------------------------------------------------------------
-# - L@ST MODIFIED: 2018-12-20 18:44 on marvin
+# - L@ST MODIFIED: 2018-12-21 17:28 on marvin
 # -------------------------------------------------------------------
 
 
@@ -66,8 +66,8 @@ foehnix.noconcomitant.fit <- function(y, family, switch = FALSE,
 
     # Initial probability (fifty fifty) and inital prior
     # probabilites for the component membership.
-    prob  <- rep(.5, length(y))
-    post  <- family$posterior(y, mean(prob), theta)
+    prob  <- mean(z)
+    post  <- family$posterior(y, prob, theta)
 
     # EM algorithm: estimate probabilities (prob; E-step), update the model
     # given the new probabilities (M-step). Always with respect to the
@@ -77,10 +77,7 @@ foehnix.noconcomitant.fit <- function(y, family, switch = FALSE,
         iter <- iter + 1;
 
         # M-step: update probabilites and theta
-        #prob  <- as.numeric(post >= .5)
-        ##TODO: prob <- post
-        #prob <- post
-        prob <- mean(post)
+        prob  <- mean(post)
         theta <- family$theta(y, post, theta = theta)
 
         # E-step: calculate a-posteriori probability
@@ -93,12 +90,6 @@ foehnix.noconcomitant.fit <- function(y, family, switch = FALSE,
         if ( verbose )
             cat(sprintf("EM iteration %d/%d, ll = %10.2f\r", iter, maxit[1L], llpath[[iter]]))
 
-        # If the log-likelihood decreases: proceed!
-        if ( iter == 1 ) next
-
-        # Improvement < 0 (model got worse): continue
-        ##if ( (llpath[[iter]]$full - llpath[[iter - 1]]$full) < 0 ) next
-
         # If the log-likelihood improvement falls below the
         # specified tolerance we assume that the algorithm
         # converged: stop iteration.
@@ -107,7 +98,17 @@ foehnix.noconcomitant.fit <- function(y, family, switch = FALSE,
             browser()
         }
 
-        if ( (llpath[[iter]]$full - llpath[[iter - 1]]$full) < tol[1L] ) break
+        # If the log-likelihood decreases: proceed!
+        if ( iter == 1 ) next
+
+        # Likelihood improvement felt below the threshold:
[…]

### 2018-12-21 R/foehnix_functions.R

@@ -9,7 +9,7 @@
 # -------------------------------------------------------------------
 # - EDITORIAL:   2018-12-16, RS: Created file on thinkreto.
 # -------------------------------------------------------------------
-# - L@ST MODIFIED: 2018-12-20 18:34 on marvin
+# - L@ST MODIFIED: 2018-12-21 16:30 on marvin
 # -------------------------------------------------------------------
 
 
@@ -79,6 +79,15 @@ scale.standardized <- function(x, ...) return(attr(x, "scaled:scale"))
 center <- function(x, ...) UseMethod("center")
 #' @export
 center.standardized <- function(x, ...) return(attr(x, "scaled:center"))
+
+destandardize <- function(x, ...) UseMethod("destandardize")
+destandardize.standardized <- function(x, ...) {
+    center <- center(x)
+    scale  <- scale(x)
+    for ( i in seq_along(center) ) x[,i] <- x[,i] * scale[i] + center[i]
+    matrix(x, ncol = ncol(x), nrow = nrow(x), dimnames = dimnames(x))
+}
+
 
 
 #' Check if Matrix is Standardized
@@ -248,14 +257,15 @@ summary.foehnix <- function(object, ...) {
     rval$coef    <- coef(object, type = "parameter")
 
     # Optimizer statistics
-    rval$time      <- object$time
-    rval$logLik    <- logLik(object)
-    rval$AIC       <- AIC(object)
-    rval$BIC       <- BIC(object)
-    rval$edf       <- edf(object)
-    rval$n.iter    <- object$optimizer$n.iter
-    rval$maxit     <- object$optimizer$maxit
-    rval$converged <- object$optimizer$converged
+    rval$filter_obj <- object$filter_obj
+    rval$time       <- object$time
+    rval$logLik     <- logLik(object)
+    rval$AIC        <- AIC(object)
+    rval$BIC        <- BIC(object)
+    rval$edf        <- edf(object)
+    rval$n.iter     <- object$optimizer$n.iter
+    rval$maxit      <- object$optimizer$maxit
+    rval$converged  <- object$optimizer$converged
 
     class(rval) <- "summary.foehnix"
     return(rval)
@@ -265,7 +275,7 @@ summary.foehnix <- function(object, ...) {
 print.summary.foehnix <- function(x, ...) {
 
     # Model call
[…]

### 2018-12-21 R/iwls_logit.R

@@ -9,7 +9,7 @@
 # -------------------------------------------------------------------
 # - EDITORIAL:   2018-11-28, RS: Created file on thinkreto.
 # -------------------------------------------------------------------
-# - L@ST MODIFIED: 2018-12-20 17:46 on marvin
+# - L@ST MODIFIED: 2018-12-21 15:24 on marvin
 # -------------------------------------------------------------------
 
 #' IWLS Solver for Binary Logistic Regression Model
@@ -125,6 +125,7 @@ iwls_logit <- function(X, y, beta = NULL, lambda = NULL, standardize = TRUE,
     if ( min(y) < 0 | max(y) > 1 ) stop("y values out of range. Have to be within ]0,1[.")
 
     # Standardize design matrix?
+    holdX <- X
     if ( standardize ) X <- standardize(X)
 
     # initialize regression coefficients if needed
@@ -139,8 +140,9 @@ iwls_logit <- function(X, y, beta = NULL, lambda = NULL, standardize = TRUE,
     # Apply link function on linear predictor to get response mu (probabilities)
     mu  <- plogis(eta) 
 
-    iter   <- 0L
-    llpath <- list()
+    iter     <- 0L
+    llpath   <- list()
+    coefpath <- list()
     while( iter < maxit ) {
         # Increase iteration counter
         iter <- iter + 1L
@@ -159,6 +161,7 @@ iwls_logit <- function(X, y, beta = NULL, lambda = NULL, standardize = TRUE,
 
         # Update log-likelihood sum
         llpath[[iter]]   <- sum(y * eta - log(1 + exp(eta)))
+        coefpath[[iter]] <- beta
 
         # Continue if iter == 1 and/or is.null(tol)
         if ( iter == 1 | is.null(tol) ) next
@@ -166,8 +169,14 @@ iwls_logit <- function(X, y, beta = NULL, lambda = NULL, standardize = TRUE,
         if ( verbose) cat(sprintf("Iteration %d, ll = %15.4f, %s\r", iter, ll,
             ifelse(is.null(lambda), "unregularized", sprintf("lambda = %10.4f", lambda))))
 
-        # Check if improvement falls below tolerance
-        if ( (llpath[[iter]] - llpath[[iter - 1]]) < tol ) break
+        # Check if improvement falls below tolerance: early stopping,
+        # remove latest likelihood/coefficients and stop iteration.
[…]

### 2018-12-21 R/tsplot.R

@@ -8,7 +8,7 @@
 # -------------------------------------------------------------------
 # - EDITORIAL:   2018-12-16, RS: Created file on thinkreto.
 # -------------------------------------------------------------------
-# - L@ST MODIFIED: 2018-12-20 18:09 on marvin
+# - L@ST MODIFIED: 2018-12-21 17:27 on marvin
 # -------------------------------------------------------------------
 
 #' Time Series Plot of foehnix Models
@@ -52,7 +52,9 @@ tsplot <- function(x, start = NULL, end = NULL, ndays = 10, ..., xtra = NULL, as
             y <- par()$usr[3:4]
             for ( i in seq(1, length(up))) {
                 if ( length(isna) > 0 ) {
-                    to <- min(min(isna[isna > up[i]]), down[i])
+                    if ( any(isna > up[i]) ) {
+                        to <- min(min(isna[isna > up[i]]), down[i])
+                    } else { to <- down[i]; }
                 } else {
                     to <- down[i]
                 }

### 2018-12-21 R/foehnix.R
Updated to devel version v0.0-5

@@ -11,7 +11,7 @@
 # -------------------------------------------------------------------
 # - EDITORIAL:   2018-11-28, RS: Created file on thinkreto.
 # -------------------------------------------------------------------
-# - L@ST MODIFIED: 2018-12-21 17:28 on marvin
+# - L@ST MODIFIED: 2018-12-21 21:14 on marvin
 # -------------------------------------------------------------------
 
 
@@ -557,6 +557,10 @@ foehnix <- function(formula, data, switch = FALSE, filter = NULL,
                     family = "gaussian",
                     control = foehnix.control(family, switch, ...), ...) { 
 
+    if ( "windfilter" %in% names(list(...)) )
+        stop("STOP! The \"windfilter\" option has been replaced by the \"filter\" option!",
+             "Please adjust your calls (replace \"windfilter\" with \"filter\"!")
+
     # Start timing (execution time of foehnix)
     timing <- Sys.time() # Measure execution time
 
@@ -792,15 +796,17 @@ predict.foehnix <- function(object, newdata = NULL, type = "response", ...) {
     # any foehn).
     idx_isna <- which(is.na(y) | apply(logitX, 1, function(x) sum(is.na(x)) > 0))
     # Inverse wind filter
-    idx_wind <- which(! 1:nrow(newdata) %in% foehnix_filter(newdata, object$filter))
+    filter_obj <- foehnix_filter(newdata, object$filter)
+    print(filter_obj)
 
     # Create return object of type zoo. By default:
     # - prob is the a-posteriory probability, flag is 1.
     # - For rows removed by the filter option: set prob = 0 and flag = 0
     # - For rows where input contained NA: set prob = NA and flag = NA
     res <- zoo(data.frame(prob = post, flag = rep(1, length(post))), index(newdata))
-    if ( length(idx_wind) > 0 ) res[idx_wind,] <- 0
-    if ( length(idx_isna) > 0 ) res[idx_isna,] <- NA
+    if ( length(filter_obj$ugly) > 0 ) res[filter_obj$ugly,] <- NA
+    if ( length(filter_obj$bad)  > 0 ) res[filter_obj$bad,]  <- 0
+
 
     # If type is response: return foehn probability
     if ( type == "response" ) return(res)
@@ -824,9 +830,8 @@ predict.foehnix <- function(object, newdata = NULL, type = "response", ...) {
 #'
 #' @param object a \code{\link{foehnix}} model object.
 #' @param which what to get as return. Can a character string, one of
[…]

### 2018-12-21 R/foehnix_filter.R

@@ -7,7 +7,7 @@
 # -------------------------------------------------------------------
 # - EDITORIAL:   2018-12-19, RS: Created file on thinkreto.
 # -------------------------------------------------------------------
-# - L@ST MODIFIED: 2018-12-21 17:15 on marvin
+# - L@ST MODIFIED: 2018-12-21 21:27 on marvin
 # -------------------------------------------------------------------
 
 
@@ -178,7 +178,7 @@
 #' print(filter_obj)
 #' 
 #' # Subsetting the 'good' rows
-#' data <- data[idx$good,]
+#' data <- data[filter_obj$good,]
 #' 
 #' summary(subset(data, select = c(dd, crest_dd)))
 #' 

### 2018-12-21 R/foehnix_functions.R

@@ -9,7 +9,7 @@
 # -------------------------------------------------------------------
 # - EDITORIAL:   2018-12-16, RS: Created file on thinkreto.
 # -------------------------------------------------------------------
-# - L@ST MODIFIED: 2018-12-21 16:30 on marvin
+# - L@ST MODIFIED: 2018-12-21 18:39 on marvin
 # -------------------------------------------------------------------
 
 
@@ -229,6 +229,10 @@ coef.foehnix <- function(object, type = "parameter", ...) {
 }
 
 #' @export
+coefficients.foehnix <- function(object, type = "parameter", ...)
+    coef.foehnix(object, type = type, ...)
+
+#' @export
 print.coef.foehnix <- function(x, ...) {
     cat("Coefficients of foehnix model\n")
     cat(sprintf("Model formula:           %s\n",
@@ -295,8 +299,6 @@ print.summary.foehnix <- function(x, ...) {
                 sum_1, sum_1 / nrow(x$prob) * 100))
     cat(sprintf("\nClimatological foehn occurance %.2f percent (on n = %d)\n", mean_occ, mean_n))
     cat(sprintf("Mean foehn probability %.2f percent (on n = %d)\n", mean_prob, mean_n))
-
-    print(x$filter_obj)
 
     cat(sprintf("\nLog-likelihood: %.1f, %d effective degrees of freedom\n",   x$logLik, x$edf))
     cat(sprintf("Corresponding AIC = %.1f, BIC = %.1f\n", x$AIC, x$BIC))

### 2018-12-21 R/plot.foehnix.R

@@ -7,7 +7,7 @@
 # -------------------------------------------------------------------
 # - EDITORIAL:   2018-12-16, RS: Created file on thinkreto.
 # -------------------------------------------------------------------
-# - L@ST MODIFIED: 2018-12-20 18:07 on marvin
+# - L@ST MODIFIED: 2018-12-21 19:05 on marvin
 # -------------------------------------------------------------------
 
 
@@ -18,10 +18,12 @@
 #' @param x a \code{\link{foehnix}} mixture model object.
 #' @param which \code{NULL} (default), character, character string,
 #'         integer, or numeric. Allowed characters: \code{loglik},
-#'         \code{loglikcontribution}, and \code{coef}. If \code{which}
+#'         \code{loglikcontribution}, \code{coef}, and \code{hist}. If \code{which}
 #'         is numeric/integer or a vector of numerics/integers the numbers
 #'         correspond to \code{loglik} (\code{1}), \code{loglikpath} (\code{2}),
 #'         or \code{coef} (\code{3}).
+#' @param log logical, if \code{TRUE} the x-axis is shown on the log scale,
+#'        else on the iteration scale.
 #' @param ... additional arguments, unused.
 #' 
 #' @details
@@ -42,7 +44,7 @@
 #' @import graphics
 #' @author Reto Stauffer
 #' @export
-plot.foehnix <- function(x, which = NULL, ...) {
+plot.foehnix <- function(x, which = NULL, log = TRUE, ...) {
 
     # Define plot type
     allowed <- c("loglik","loglikcontribution", "coef", "hist")
@@ -51,7 +53,7 @@ plot.foehnix <- function(x, which = NULL, ...) {
     } else if ( inherits(which, c("integer", "numeric")) ) {
         which <- allowed[as.integer(which)]
     } else {
-        which <- match.arg(which, allowed, several.ok = TRUE)
+        which <- match.arg(tolower(which), allowed, several.ok = TRUE)
     }
 
     # Keep user params
@@ -67,49 +69,50 @@ plot.foehnix <- function(x, which = NULL, ...) {
     # Plotting likelihood path if requested
     if ( "loglik" %in% which ) {
         ll <- x$optimizer$loglikpath
-        ylim <- range(ll) + c(0, 0.1) * diff(range(ll))
+        ylim <- range(ll) + c(0, 0.2) * diff(range(ll))
         # Plot
[…]

### 2018-12-21 R/wind_functions.R

@@ -1,0 +1,150 @@
+# -------------------------------------------------------------------
+# - NAME:        wind_functions.R
+# - AUTHOR:      Reto Stauffer
+# - DATE:        2018-12-21
+# -------------------------------------------------------------------
+# - DESCRIPTION:
+# -------------------------------------------------------------------
+# - EDITORIAL:   2018-12-21, RS: Created file on thinkreto.
+# -------------------------------------------------------------------
+# - L@ST MODIFIED: 2018-12-21 21:23 on marvin
+# -------------------------------------------------------------------
+
+#' Convert U and V Wind Components to Wind Speed and Direction
+#'
+#' Takes U/V wind components as input and returns wind speed
+#' and meteorological wind direction.
+#'
+#' @param u \code{numeric} vector with U components or a 
+#'    \code{data.frame} or \code{zoo} object containing
+#'    a column named \code{u} and a column named \code{v}.
+#' @param v \code{NULL} (default) or \code{numeric} vector.
+#'    Iv input \code{u} is a single vector \code{v} has to be given.
+#' @param rad \code{logical}, default is \code{FALSE}. Returns wind
+#'    direction in radiant rather than in degrees.
+#'
+#' @return Returns a data.frame with two columns named \code{ff}
+#'    and \code{dd} containing wind speed (same physical unit as
+#'    input \code{u}/\code{v}) and wind direction. Wind direction
+#'    is either in meteorological degrees (0 from North, from 90 East,
+#'    180 from South, and 270 from West) or in mathematical radiant
+#'    if input \code{rad = TRUE}.
+#'
+#' @seealso ddff2uv
+#' @examples
+#' ## Generate data.frame with u/v components for all 4 main wind directions
+#' data <- data.frame( name = c("N","E","S","W"),
+#'                     u    = c( 0, -1 , 0,  1 ),
+#'                     v    = c(-1,  0,  1,  0 ) )
+#' ## Use data.frame input
+#' ddff <- uv2ddff( data )
+#' cbind( data, ddff )
+#' ## Use u/v components as two separate vector inputs
+#' ddff <- uv2ddff( data$u, data$v )
+#' cbind( data, ddff )
+#' ## Radiant
[…]

### 2018-12-21 R/windrose.R

@@ -7,7 +7,7 @@
 # -------------------------------------------------------------------
 # - EDITORIAL:   2018-12-16, RS: Created file on thinkreto.
 # -------------------------------------------------------------------
-# - L@ST MODIFIED: 2018-12-20 18:21 on marvin
+# - L@ST MODIFIED: 2018-12-21 21:05 on marvin
 # -------------------------------------------------------------------
 
 #' Windrose Plot
@@ -23,13 +23,16 @@
 #'        for training the \code{\link{foehnix}} mixture model.
 #' @param ffvar character, name of the wind speed varaible in \code{x} or as used
 #'        for training the \code{\link{foehnix}} mixture model.
+#' @param maxpp maximum number of plot per page. If not all plots fit
+#'        on one page multiple pages will be shown as a sequence of plots.
 #' @param ... additional arguments, currently unused.
 #'
 #' @details TODO: not yet finished and/or tested!
 windrose <- function(x, which = NULL, ddvar = "dd", ffvar = "ff", ...) UseMethod("windrose")
 
 #' @rdname windrose
-windrose.foehnix <- function(x, which = NULL, ddvar = "dd", ffvar = "ff", ...) {
+windrose.foehnix <- function(x, which = NULL, ddvar = "dd", ffvar = "ff",
+                             maxpp = Inf, ...) {
 
     # First: checking argument "which"
     allowed <- c("unconditional", "nofoehn", "foehn", "combined")
@@ -43,14 +46,10 @@ windrose.foehnix <- function(x, which = NULL, ddvar = "dd", ffvar = "ff", ...) {
                                paste(seq_along(allowed), collapse = ", "))))
         which <- allowed[which]
     }
-    if ( ! all(which %in% allowed) ) {
-        stop(paste("Nonsuitable input \"which\". Out of range. Allowed:",
-                   sprintf("%s (or %s) or combinations of them.",
-                           paste(sprintf("\"%s\"", allowed), collapse = ", "),
-                           paste(seq_along(allowed), collapse = ", "))))
-    }
+    which <- match.arg(which, allowed, several.ok = TRUE)
 
-    # Checking dd
+    # -------------------
[…]

### 2018-12-22 R/windrose.R
Update to v0.0-6

@@ -7,7 +7,7 @@
 # -------------------------------------------------------------------
 # - EDITORIAL:   2018-12-16, RS: Created file on thinkreto.
 # -------------------------------------------------------------------
-# - L@ST MODIFIED: 2018-12-21 21:05 on marvin
+# - L@ST MODIFIED: 2018-12-22 15:30 on marvin
 # -------------------------------------------------------------------
 
 #' Windrose Plot
@@ -15,38 +15,54 @@
 #' Plotting a windrose using meteorological wind direction and
 #' wind speed (or gust speed).
 #'
-#' @param x \code{zoo} object or \code{data.frame} or \code{\link{foehnix}}
-#'        mixture model object.
-#' @param which character, one of \code{"unconditional"}, \code{"nofoehn"}, 
-#'        \code{"foehn"}, or \code{"combined"}.
-#' @param ddvar character, name of the wind direction varaible in \code{x} or as used
-#'        for training the \code{\link{foehnix}} mixture model.
-#' @param ffvar character, name of the wind speed varaible in \code{x} or as used
-#'        for training the \code{\link{foehnix}} mixture model.
-#' @param maxpp maximum number of plot per page. If not all plots fit
-#'        on one page multiple pages will be shown as a sequence of plots.
-#' @param ... additional arguments, currently unused.
+#' @param x forwarded to windrose S3 methods.
+#' @param ... forwarded as well.
+#'
+#' @seealso windrose.default windrose.foehnix
 #'
 #' @details TODO: not yet finished and/or tested!
-windrose <- function(x, which = NULL, ddvar = "dd", ffvar = "ff", ...) UseMethod("windrose")
+#' @export
+windrose <- function(x, ...) UseMethod("windrose")
 
+
+#' foehnix Mixture Model Windrose Plot
+#'
+#' Windrose plot based on a \code{\link{foehnix}} mixture model object.
+#'
+#' @param x object of class \code{\link{foehnix}}.
+#' @param type \code{NULL} or character, one of \code{density} or \code{histogram}.
+#'        If \code{NULL} both types will be plotted.
+#' @param which \code{NULL} or character, one of \code{unconditional}, \code{nofoehn},
+#'        or \code{foehn}. If \code{NULL} all three will be plotted.
[…]

### 2019-01-02 index.Rmd
Update to v0.0-6
Local commit

@@
 
+```{r, echo = FALSE, fig = TRUE, results = 'hide', fig.width = 12, fig.height = 4}
+windrose(mod, type = "histogram", mfcol = 3)
+```
+
 ### References
@@
 <a href="https://dx.doi.org/10.1175/JAMC-D-13-0267.1" target="_blank">10.1175/JAMC-D-13-0267.1</a>
+</p>
+
+<p id="hastie2009">
+Hastie T, Tibshirani R, Friedman J (2009).
+Fitting Logistic Regression Models. In _The Elements of Statistical Learning_
+(Chapter 4.4.1), 2_nd_ edition, ISBN 978-0387848570.
+<a href="https://web.stanford.edu/~hastie/ElemStatLearn/" target="_blank">PDF download</a>
 </p>
@@
 </p>
+
+<p id="mccullagh1999">
+McCullagh P, Nelder JA (1999).
+Likelihood functions for binary data. In _Generalized Linear Models_ (Chapter 4.4),
+2_nd_ edition, ISBN 0-412-31760-5.
+</p>

### 2019-01-02 R/foehnix.R

@@ -11,7 +11,7 @@
 # -------------------------------------------------------------------
 # - EDITORIAL:   2018-11-28, RS: Created file on thinkreto.
 # -------------------------------------------------------------------
-# - L@ST MODIFIED: 2018-12-21 21:14 on marvin
+# - L@ST MODIFIED: 2019-01-02 15:39 on marvin
 # -------------------------------------------------------------------
 
 
@@ -24,7 +24,7 @@
 #' @param y numeric vector, covariate for the components of the
 #'        mixture model, dimension \code{N}.
 #' @param family object of class \code{\link{foehnix.family}}.
-#' @param logical wether or not the two components should be switched.
+#' @param switch logical whether or not the two components should be switched.
 #'        By default (\code{switch = FALSE}) the component which shows
 #'        higher values of \code{y} is assumed to be the foehn cluster!
 #'        Depending on what your covariate is you might need to switch
@@ -153,7 +153,7 @@ if ( inherits(y, "binned") ) stop("Stop, requires changes on computation of BIC!
 #' @param logitX numeric matrix of dimension \code{N x P}, covariates
 #'        for the concomitant model (logistic regression model matrix).
 #' @param family object of class \code{\link{foehnix.family}}.
-#' @param logical wether or not the two components should be switched.
+#' @param switch logical whether or not the two components should be switched.
 #'        By default (\code{switch = FALSE}) the component which shows
 #'        higher values of \code{y} is assumed to be the foehn cluster!
 #'        Depending on what your covariate is you might need to switch
@@ -290,7 +290,7 @@ if ( inherits(y, "binned") ) stop("Stop, requires changes on computation of BIC!
 #' @param logitX numeric matrix of dimension \code{N x P}, covariates
 #'        for the concomitant model (logistic regression model matrix).
 #' @param family object of class \code{\link{foehnix.family}}.
-#' @param logical wether or not the two components should be switched.
+#' @param switch logical whether or not the two components should be switched.
 #'        By default (\code{switch = FALSE}) the component which shows
 #'        higher values of \code{y} is assumed to be the foehn cluster!
 #'        Depending on what your covariate is you might need to switch
@@ -329,7 +329,7 @@ foehnix.reg.fit <- function(y, logitX, family, switch = FALSE,
 #' @param family character specifying the distribution of the components in the
 #'        mixture model. Allowed: \code{"gaussian"} and \code{"logistic"}.  For
 #'        experts: custom \code{foehnix.family} objects can be provided as well.
[…]

### 2019-01-02 R/foehnix_filter.R

@@ -272,7 +272,7 @@ foehnix_filter <- function(x, filter) {
 
 #' @rdname foehnix_filter
 #' @export
-print.foehnix.filter <- function(x) {
+print.foehnix.filter <- function(x, ...) {
     cat("\nFoehnix Filter Object:\n")
     cat("Call: "); print(x$call)
     cat(sprintf("Total data set length:          %8d\n", x$total))

### 2019-01-02 R/foehnix_functions.R

@@ -9,7 +9,7 @@
 # -------------------------------------------------------------------
 # - EDITORIAL:   2018-12-16, RS: Created file on thinkreto.
 # -------------------------------------------------------------------
-# - L@ST MODIFIED: 2018-12-21 18:39 on marvin
+# - L@ST MODIFIED: 2019-01-02 15:35 on marvin
 # -------------------------------------------------------------------
 
 
@@ -171,7 +171,7 @@ edf.foehnix <- function(object, ...)     structure(object$optimizer$edf, names =
 
 #' @rdname foehnix
 #' @export
-print.foehnix <- function(object, ...) print(summary(object, ...))
+print.foehnix <- function(x, ...) print(summary(object, ...))
 
 
 #' Get Estimated Mixture Model Coefficients
@@ -248,7 +248,7 @@ print.coef.foehnix <- function(x, ...) {
 
 #' @rdname foehnix
 #' @export
-formula.foehnix <- function(object, ...) object$formula
+formula.foehnix <- function(x, ...) x$formula
 
 
 #' @rdname foehnix
@@ -311,3 +311,129 @@ print.summary.foehnix <- function(x, ...) {
     }
 }
 
+#' foehnix Image Plot - Hoevmoeller Diagram
+#'
+#' The default \code{\link{image}} plot of a \code{\link{foehnix}} object
+#' is a Hoevmoeller diagram.
+#'
+#' @param x object of class \code{\link{foehnix}}.
+#' @param FUN character string or a custom aggregation function. See 'Details'
+#'        section for more information.
+#' @param contours logical \code{TRUE} or \code{FALSE}, whether or not the
+#'        concours should be added.
+#' @param deltat integer, interval in seconds for the time of day axis. Has to be
+#'        a fraction of 86400 (24 hours in seconds). It \code{NuLL} (default) the
+#'        interval of the time series object will be used.
+#' @param ... additional arguments, currently unused.
+#'
+#' @details Plotting a Hoevmoeller diagram based on the \code{\link{zoo}} time
+#' series object of the \code{\link{foehnix}} classification. Different plot
+#' types are available. The default functions (see list below) use \code{na.rm = TRUE}.
+#'
[…]

### 2019-01-02 R/windrose.R

@@ -7,7 +7,7 @@
 # -------------------------------------------------------------------
 # - EDITORIAL:   2018-12-16, RS: Created file on thinkreto.
 # -------------------------------------------------------------------
-# - L@ST MODIFIED: 2018-12-22 15:30 on marvin
+# - L@ST MODIFIED: 2019-01-02 15:34 on marvin
 # -------------------------------------------------------------------
 
 #' Windrose Plot
@@ -15,12 +15,11 @@
 #' Plotting a windrose using meteorological wind direction and
 #' wind speed (or gust speed).
 #'
-#' @param x forwarded to windrose S3 methods.
+#' @param x forwarded to windrose methods.
 #' @param ... forwarded as well.
 #'
 #' @seealso windrose.default windrose.foehnix
 #'
-#' @details TODO: not yet finished and/or tested!
 #' @export
 windrose <- function(x, ...) UseMethod("windrose")
 
@@ -38,9 +37,29 @@ windrose <- function(x, ...) UseMethod("windrose")
 #'        the wind direction information.
 #' @param ffvar character, name of the column in the training data set which contains
 #'        the wind speed (or gust speed) data.
-#' @param mfcolinteger, number of columns of subplots.
+#' @param mfcol integer, number of columns of subplots.
 #' @param maxpp integer (\code{>0}), maximum plots per page. Not all plots fit on one
 #'        page the script loops trough.
+#'
+#' @details Allows to draw windrose plots from a \code{\link{foehnix}} mixture model
+#' object. Two \code{type}s are available: circular density plots and circular
+#' histograms, both for \code{which = "unconditional"}, \code{which = "nofoehn"},
+#' and \code{which = "foehn"}, details below.
+#' By default (\code{type = NULL}, \code{which = NULL}) all combinations will be plotted.
+#'
+#' \itemize{
+#'    \item \code{which = "unconditional"}: unconditional windrose (all observations
+#'          of the data set used to estimate the \code{\link{foehnix}} model).
+#'    \item \code{which = "nofoehn"}: windrose of all observations which have not been
+#'          classified as foehn (probability \code{< 0.5}).
+#'    \item \code{which = "foehn"}: windrose of all observations classified as foehn events
+#'          (probability \code{>= 0.5}).
+#' }
+#'
[…]

### 2019-01-03 index.Rmd
Pushing 0.0-6

@@
 windrose(mod, type = "histogram", mfcol = 3)
+```
+
+```{r, echo = FALSE, fig = TRUE, results = 'hide', fig.width = 12, fig.height = 6.5}
+col <- rev(colorspace::sequential_hcl(21, h = c(-50, 120), c. = c(80, 10)))
+image(mod, FUN = "freq", deltad = 7L,
+      col = col, contours = TRUE, drawlabels = FALSE,
+      main = "Hourly foehn Frequency Diagram, Weekly Aggregation")
 ```

### 2019-01-03 R/windrose.R

@@ -7,7 +7,7 @@
 # -------------------------------------------------------------------
 # - EDITORIAL:   2018-12-16, RS: Created file on thinkreto.
 # -------------------------------------------------------------------
-# - L@ST MODIFIED: 2019-01-02 15:34 on marvin
+# - L@ST MODIFIED: 2019-01-03 13:58 on marvin
 # -------------------------------------------------------------------
 
 #' Windrose Plot
@@ -226,6 +226,8 @@ windrose_get_density <- function(x, dd.breaks = seq(0, 360, by = 30),
 #' dimension as \code{x} with colors for the different bins based on the
 #' counts, and \code{legend} with levels and colors for the color legend of the
 #' plot.
+#'
+#' @importFrom grDevices rgb
 windrose_get_cols <- function(x, col, p = 1, ncol = 50L) {
     # Extend colors if needed
     if ( length(col) == 1 ) col <- rep(col, ncol)

### 2019-01-04 R/foehnix.R
Moving to 0.0-8

@@ -11,7 +11,7 @@
 # -------------------------------------------------------------------
 # - EDITORIAL:   2018-11-28, RS: Created file on thinkreto.
 # -------------------------------------------------------------------
-# - L@ST MODIFIED: 2019-01-02 15:39 on marvin
+# - L@ST MODIFIED: 2019-01-04 13:31 on marvin
 # -------------------------------------------------------------------
 
 
@@ -670,6 +670,18 @@ foehnix <- function(formula, data, switch = FALSE, filter = NULL,
 
     # Store coefficients
     res$coef <- rval$theta; res$coef$concomitants = coef
+
+    # Calculate the weightes standard error of the estimated
+    # coefficients for the test statistics.
+    # First: calculate weighted sum of squared residuals for
+    # both components
+    res_c1 <- (y - res$coef$mu1) * (1 - rval$post)
+    res_c2 <- (y - res$coef$mu2) *      rval$post
+    mu1.se <- sqrt((sum(res_c1^2) / (sum((1 - rval$post)^2) * (sum(1 - rval$post) - 1))))
+    mu2.se <- sqrt((sum(res_c2^2) / (sum(     rval$post^2 ) * (sum(    rval$post) - 1))))
+
+    # Standard errors for intercept of mu1 (component1) and mu2 (component 2)
+    res$mu.se  <- setNames(c(mu1.se, mu2.se), c("mu1.se", "mu2.se"))
 
     # The final result, the foehn probability. Creates an object
     # of the same class as the input "data" (currently only zoo!)

### 2019-01-16 index.Rmd
Small updates

@@
 # Loading observations (data.frame), convert to zoo
-data("ellboegen", package = "foehnix")
-data("sattelberg", package = "foehnix")
-timefun    <- function(x) as.POSIXct(x,  origin = "1970-01-01", tz = "UTC")
-ellboegen  <- zoo(ellboegen[,-1],  timefun(ellboegen$timestamp))
-sattelberg <- zoo(sattelberg[,-1], timefun(sattelberg$timestamp))
-names(sattelberg) <- paste0("crest_", names(sattelberg))
-data <- merge(ellboegen, sattelberg)
-data$diff_t <- data$crest_t + 10.27 - data$t
+data <- demodata()
 filter <- list(dd = c(43, 223), crest_dd = c(90, 270))

### 2019-01-16 R/datasets.R

@@ -34,7 +34,7 @@
 #'    \item \code{t}: dry air temperature in degrees Celsius
 #' }
 #'
-#' @seealso \code{\link{sattelberg}}.
+#' @seealso \code{\link{sattelberg}}, \code{\link{demodata}}.
 #'
 #' @name ellboegen
 #' @docType data
@@ -77,7 +77,7 @@ NULL
 #'    \item \code{t}: dry air temperature in degrees Celsius
 #' }
 #'
-#' @seealso \code{\link{ellboegen}}.
+#' @seealso \code{\link{ellboegen}}, \code{\link{demodata}}.
 #'
 #' @name sattelberg
 #' @docType data

### 2019-01-16 R/foehnix_filter.R

@@ -7,7 +7,7 @@
 # -------------------------------------------------------------------
 # - EDITORIAL:   2018-12-19, RS: Created file on thinkreto.
 # -------------------------------------------------------------------
-# - L@ST MODIFIED: 2019-01-07 21:47 on marvin
+# - L@ST MODIFIED: 2019-01-16 23:44 on marvin
 # -------------------------------------------------------------------
 
 
@@ -128,10 +128,8 @@
 #' @examples
 #' # Loading example data set and conver to zoo time series
 #' # time series object.
-#' data("ellboegen", package = "foehnix")
-#' library("zoo")
-#' ellboegen <- zoo(subset(ellboegen, select = -timestamp),
-#'                  as.POSIXct(ellboegen$timestamp, origin = "1970-01-01", tz = "UTC"))
+#' # Loading combined demo data set
+#' ellboegen <- demodata("ellboegen")
 #' 
 #' # Case 1:
 #' # -----------------
@@ -160,17 +158,8 @@
 #' # has to show southerly flows with a wind direction from
 #' # 90 degrees (East) to 270 degrees (West).
 #' 
-#' # Loading Sattelberg station data and convert to zoo
-#' # time series object.
-#' data("sattelberg", package = "foehnix")
-#' require("zoo")
-#' sattelberg <- zoo(subset(sattelberg, select = -timestamp),
-#'                  as.POSIXct(sattelberg$timestamp, origin = "1970-01-01", tz = "UTC"))
-#' names(sattelberg) <- paste0("crest_", names(sattelberg)) # Renaming variables
-#' 
-#' # Combine Ellboegen observations with Sattelberg observations
-#' data <- merge(ellboegen, sattelberg)
-#' print(head(data))
+#' # Loading combined demo data set
+#' data <- demodata()
 #' 
 #' # Now apply a wind filter
 #' my_filter  <- list(dd = c(43, 223), crest_dd = c(90, 270))

### 2019-01-16 R/windrose.R

@@ -7,7 +7,7 @@
 # -------------------------------------------------------------------
 # - EDITORIAL:   2018-12-16, RS: Created file on thinkreto.
 # -------------------------------------------------------------------
-# - L@ST MODIFIED: 2019-01-16 22:19 on marvin
+# - L@ST MODIFIED: 2019-01-16 23:41 on marvin
 # -------------------------------------------------------------------
 
 #' Windrose Plot
@@ -71,21 +71,8 @@ windrose <- function(x, ...) UseMethod("windrose")
 #' (\code{type = "histogram"} or \code{type = "density"}).
 #'
 #' @examples
-#' # Loading observations (data.frame), convert to zoo
-#' data("ellboegen",  package = "foehnix")
-#' data("sattelberg", package = "foehnix")
-#' ellboegen  <- zoo(ellboegen[,-1],  as.POSIXct(ellboegen[,1],  origin = "1970-01-01"))
-#' sattelberg <- zoo(sattelberg[,-1], as.POSIXct(sattelberg[,1], origin = "1970-01-01"))
-#' 
-#' # Modify sattelberg variable names (crest_ identifies Sattelberg
-#' # observations, our crest station) and combine both data sets.
-#' names(sattelberg) <- paste0("crest_", names(sattelberg))
-#' data <- merge(ellboegen, sattelberg)
-#' 
-#' # Dry adiabatic temperature difference between 
-#' # Sattelberg (data$crest_t) and Ellboegen (data$t) corrected by
-#' # 1027/10 degrees.
-#' data$diff_t <- data$crest_t + 10.27 - data$t
+#' # Loading combined demo data set
+#' data <- demodata()
 #'
 #' # Before estimating a model: plot a wind rose for all observations
 #' windrose(data$dd, data$ff, type = "histogram")

### 2019-01-19 R/foehnix_filter.R
Update to v0.0-10

@@ -7,7 +7,7 @@
 # -------------------------------------------------------------------
 # - EDITORIAL:   2018-12-19, RS: Created file on thinkreto.
 # -------------------------------------------------------------------
-# - L@ST MODIFIED: 2019-01-16 23:44 on marvin
+# - L@ST MODIFIED: 2019-01-19 19:34 on marvin
 # -------------------------------------------------------------------
 
 
@@ -35,6 +35,8 @@
 #'        function operating on \code{x}, or a named list with a simple
 #'        filter rule (\code{numeric} of length two) or custom filter
 #'        functions. Details provided in the 'Details' section.
+#' @param ... currently unused.
+#'
 #' @return Returns a vector of integers corresponding to those rows in
 #' the data set \code{x} which fulfill all filte criteria. If input
 #' \code{filter = NULL} an integer vector \code{1:nrow(x)} is returned.

### 2019-01-19 R/foehnix_functions.R

@@ -9,9 +9,10 @@
 # -------------------------------------------------------------------
 # - EDITORIAL:   2018-12-16, RS: Created file on thinkreto.
 # -------------------------------------------------------------------
-# - L@ST MODIFIED: 2019-01-07 21:29 on marvin
+# - L@ST MODIFIED: 2019-01-19 19:33 on marvin
 # -------------------------------------------------------------------
 
+utils::globalVariables(c("time_mid", "yday_mid", "value"))
 
 #' Standardize (Model) Matrix
 #'
@@ -362,7 +363,7 @@ print.summary.foehnix <- function(x, ...) {
 #' @param deltat integer, interval in seconds for the time of day axis. Has to be
 #'        a fraction of 86400 (24 hours in seconds). It \code{NuLL} (default) the
 #'        interval of the time series object will be used.
-#' @param deltat integer, similar to \code{deltat}, the interval in days for the
+#' @param deltad integer, similar to \code{deltat}, the interval in days for the
 #'        grid on the x-axis. Default is \code{7L} (aggregate to weekly values).
 #' @param col vector of colors forwarded to \code{image.default}. By default
 #'        a gray scale color map is used.

### 2019-01-19 R/tsplot.R

@@ -8,7 +8,7 @@
 # -------------------------------------------------------------------
 # - EDITORIAL:   2018-12-16, RS: Created file on thinkreto.
 # -------------------------------------------------------------------
-# - L@ST MODIFIED: 2019-01-18 19:31 on marvin
+# - L@ST MODIFIED: 2019-01-19 19:36 on marvin
 # -------------------------------------------------------------------
 
 
@@ -77,7 +77,8 @@ tsplot.control <- function(...) {
     diff_t; diff_t;  orange;  temperature difference [C]
     dd;     dd;      black;   wind direction [deg]
     ff;     ff;      #005ce6; wind speed [m/s]
-    ffx;    ffx;     #5c00e6; gust speed [m/s]"
+    ffx;    ffx;     #5c00e6; gust speed [m/s]
+    prob;   ----;    #FF6666; probability"
     # Reading data.frame definition from the string above.
     def <- read.table(textConnection(def), sep = ";",
                       header = TRUE,
@@ -87,7 +88,7 @@ tsplot.control <- function(...) {
     # User arguments
     arg <- list(...)
     for ( n in names(arg) ) {
-        idx <- which(def$name == n)
+        idx <- which(def$var == n)
         if ( length(idx) == 0 ) next
         # Else modify the definition.
         # If input is a character: modify name
@@ -105,13 +106,17 @@ tsplot.control <- function(...) {
         }
     }
 
+    # Polygons use hex + alpha. Thus, we do have to convert
+    # all colors to hex.
+    tmp  <- grDevices::col2rgb(def$color) / 255
+    def$color <- colorspace::hex(colorspace::sRGB(tmp[1,], tmp[2,], tmp[3,]))
+
     # Add custom class and return
     class(def) <- c("tsplot.control", "data.frame")
     return(def)
 }
 
 #' @rdname tsplot.control
-get <- function(x, ...) UseMethod("get")
 get.tsplot.control <- function(x, var, property) {
     # Check if we have the property (match.arg)
     property <- match.arg(property, names(x))
@@ -126,14 +131,16 @@ get.tsplot.control <- function(x, var, property) {
 #' Development time series plots of estimated \code{\link{foehnix}}
 #' models. TODO: they are very specific at the moment!
 #' 
-#' @param x object of type \code{foehnix}.
+#' @param x object of type \code{foehnix} or a list with \code{foehnix} and
+#'     univariate \code{zoo} objects, see 'Details'.
[…]

### 2019-08-06 R/windrose.R
Update tsplot

@@ -425,18 +425,6 @@ windrose.default <- function(x, ff,
         filter_obj <- NULL
     }
 
-    # Convert input "windsector" to list if needed.
-    windsector_convert <- function(x) {
-        if (inherits(x, c("NULL", "list"))) return(x)
-        # If matrix convert to list
-        if (inherits(x, "data.frame")) x <- as.matrix(x)
-        hadnames <- !is.null(rownames(x))
-        x <- as.list(as.data.frame(t(x)))
-        if (!hadnames) names(x) <- NULL
-        return(x)
-    }
-    if (!inherits(windsector, c("NULL", "list", "matrix", "data.frame")))
-        stop("input \"windsector\" of unexpected class. Stop.")
     windsector <- windsector_convert(windsector)
 
     # If ff is missing but x is of class zoo or data.frame with

### 2019-08-06 tests/testthat/test_foehnix_functions.R

@@ -3,6 +3,26 @@
 library("testthat")
 library("foehnix")
 
+
+test_that("Convert windsector information", {
+
+    # Allowed: NULL, matrix, list, data.frame
+    expect_is(foehnix:::windsector_convert(NULL), "NULL")
+    expect_is(foehnix:::windsector_convert(matrix(c(10, 30, 90, 140), byrow = TRUE, ncol = 2)), "list")
+    expect_is(foehnix:::windsector_convert(matrix(c(10, 30, 90, 140), byrow = TRUE, ncol = 2,
+                                        dimnames = list(c("A", "B"), c("from", "to")))), "list")
+    expect_is(foehnix:::windsector_convert(list(c(10, 30), c(90, 140))), "list")
+    expect_is(foehnix:::windsector_convert(list(A = c(10, 30), B = c(90, 140))), "list")
+    expect_is(foehnix:::windsector_convert(data.frame(from = c(30, 90), to = c(30, 140))), "list")
+    expect_is(foehnix:::windsector_convert(structure(data.frame(from = c(30, 90), to = c(30, 140)),
+                                           row.names = c("foo", "bar"))), "list")
+
+    # Not allowed
+    expect_error(foehnix:::windsector_convert(c(0, 100)))
+    expect_error(foehnix:::windsector_convert(100, 200))
+})
+
+
 test_that("Foehnix functions: standardize/destandardize", {
 
     # Wrong inputs

### 2019-08-06 tests/testthat/test_plot.R

@@ -210,9 +210,6 @@ test_that("Testing tsplot.foehnix function", {
     modX <- mod0; modX$prob$prob <- NA
     expect_silent(tsplotfun(modX, start = start, end = end, t = "myt", rh = "myrh",
               diff_t = "mydiff_t", dd = "mydd", ff = "myff", prob = "myprob", ask = FALSE))
-    tsplot(modX, start = start, end = end, t = "myt", rh = "myrh",
-              diff_t = "mydiff_t", dd = "mydd", ff = "myff", prob = "myprob", ask = FALSE)
-
     # Should result in an error as probabilities out of range
     modX$prob$prob <- 10
     expect_error(tsplotfun(modX, start = start, end = end, t = "myt", rh = "myrh",

### 2019-08-08 index.Rmd
Update master to current devel version

@@
 col <- rev(colorspace::sequential_hcl(21, h = c(-50, 120), c. = c(80, 10)))
+devtools::load_all()
+image(mod)
 image(mod, FUN = "freq", deltad = 7L,

### 2019-08-08 R/datasets.R

@@ -2,7 +2,7 @@
 #' Meteorological Observation Data, Station Ellboegen
 #'
 #' The \code{\link{foehnix}} package comes with two data sets of meteorological
-#' observations for two sites in Tyrol, Austria.  One station is located nearby
+#' observations for two sites in Tyrol, Austria.  One station is located near
 #' Ellboegen, a small town close to Innsbruck, the second station (Sattelberg)
 #' is located south of Ellboegen close to the border between Austria and Italy
 #' on top of a mountain close to the crest of the European Alps and is used as
@@ -22,30 +22,31 @@
 #'     \item Altitude: 1080 meters above mean sea level
 #' }
 #' 
-#' Variable description (columns):
+#' \code{zoo} time series object with an hourly temporal resolution (time zone: UTC)
+#' including the following variables:
 #' \itemize{
-#'    \item \code{timestamp}: UNIX time stamp, seconds since 1970-01-01 00:00:00
 #'    \item \code{dd}: wind direction in degrees, meteorological format
 #'        (0/360 correspond to 'wind from north', 90 to 'wind from east',
-#'        180 'wind from south', and 270 'wind from west'
+#'        180 'wind from south', and 270 'wind from west')
 #'    \item \code{ff}: wind speed in meters per second
 #'    \item \code{p}: station pressure in hectopascal
 #'    \item \code{rh}: relative humidity in percent
 #'    \item \code{t}: dry air temperature in degrees Celsius
 #' }
 #'
-#' @seealso \code{\link{sattelberg}}, \code{\link{demodata}}.
+#' @seealso \code{\link{sattelberg}}, \code{\link{demodata}},
+#'          \code{\link{viejas}}, \code{\link{luckyfive}}.
 #'
 #' @name ellboegen
 #' @docType data
 #' @keywords data ellboegen
-#' @author Reto Stauffer
+#' @author Reto Stauffer, Deborah Detka
 NULL
 
 #' Meteorological Observation Data, Station Sattelberg
 #'
 #' The \code{\link{foehnix}} package comes with two data sets of meteorological
-#' observations for two sites in Tyrol, Austria.  One station is located nearby
+#' observations for two sites in Tyrol, Austria.  One station is located near
 #' Ellboegen (see \code{\link{ellboegen}}), a small town close to Innsbruck,
 #' the second station (Sattelberg) is located south of Ellboegen close to the
[…]

### 2019-08-08 R/demodata.R

@@ -1,15 +1,3 @@
-# -------------------------------------------------------------------
-# - NAME:        demodata.R
-# - AUTHOR:      Reto Stauffer
-# - DATE:        2019-01-16
-# -------------------------------------------------------------------
-# - DESCRIPTION: A small function which returns the demo data set.
-# -------------------------------------------------------------------
-# - EDITORIAL:   2019-01-16, RS: Created file on thinkreto.
-# -------------------------------------------------------------------
-# - L@ST MODIFIED: 2019-01-24 12:26 on marvin
-# -------------------------------------------------------------------
-
 
 #' foehnix Demo Data Set
 #'
@@ -24,63 +12,103 @@
 #' crest station as concomitant variables and for custom wind
 #' filters (\code{\link{foehnix_filter}}).
 #'
-#' @param which either \code{"combined"} (returns a combined data set),
-#'    \code{"sattelberg"} (only observations from station Sattelberg),
-#'    or \code{"ellboegen"} (only observations rom station Ellboegen).
+#' @param which the \code{foehnix} package comes with two demo data sets.
+#'      Each consists of meteorological observations of two stations (one
+#'      in the valley; target station) and one near the crest upstream of
+#'      typical foehn direction (crest station).
+#'      \code{which = "tyrol"} returns the tyrolean data set (A),
+#'      \code{which = "california"} the californian data set (USA), a time series
+#'      object which contains measurements from both stations (valley/crest).
+#'      The function can also be used to get the demo data sets for specific sites
+#'      (see 'Examples').
 #'
-#' @source Station operated by the Department of Atmospheric and Cryospheric
+#' @source
+#' Ellboegen and Sattelberg:
+#' Operated by the Department of Atmospheric and Cryospheric
 #' Sciences (\url{http://acinn.uibk.ac.at}) of the University of Innsbruck.
 #' The data is available under the
 #' \href{https://creativecommons.org/licenses/by-sa/4.0/}{CC BY-SA 4.0 license}.
+#'
+#' Viejas and the Lucky Five Ranch:
+#' Operated by #TODO add correct source (GM).
 #' 
 #' @examples
-#' # Loading the combined demo data set. Variables starting with
-#' # "crest_" are observations from station Sattelberg.
[…]

### 2019-08-08 R/foehnix.R

@@ -11,7 +11,7 @@
 # -------------------------------------------------------------------
 # - EDITORIAL:   2018-11-28, RS: Created file on thinkreto.
 # -------------------------------------------------------------------
-# - L@ST MODIFIED: 2019-04-18 17:11 on marvin
+# - L@ST MODIFIED: 2019-08-07 16:39 on marvin
 # -------------------------------------------------------------------
 
 
@@ -773,7 +773,6 @@ foehnix <- function(formula, data, switch = FALSE, filter = NULL,
     # outside the user defined sector. These will not be considered in the
     # statistical models.
     filter_obj <- foehnix_filter(data, filter)
-
 
     # Take all elements which are not NA and where the
     # foehnix_filter routine returned an index on "good" (not

### 2019-08-08 R/foehnix_filter.R

@@ -7,7 +7,7 @@
 # -------------------------------------------------------------------
 # - EDITORIAL:   2018-12-19, RS: Created file on thinkreto.
 # -------------------------------------------------------------------
-# - L@ST MODIFIED: 2019-04-15 17:43 on marvin
+# - L@ST MODIFIED: 2019-08-07 16:31 on marvin
 # -------------------------------------------------------------------
 
 

### 2019-08-08 R/image.R

@@ -1,0 +1,339 @@
+
+#' foehnix Image Plot - Hovmoeller Diagram
+#'
+#' The default \code{\link{image}} plot of a \code{\link{foehnix}} object
+#' is a Hovmoeller diagram.
+#'
+#' @param x object of class \code{\link{foehnix}}.
+#' @param FUN character string or a custom aggregation function. See 'Details'
+#'        section for more information.
+#' @param deltat integer, interval in seconds for the time of day axis. Has to be
+#'        a fraction of 86400 (24 hours in seconds). It \code{NuLL} (default) the
+#'        interval of the time series object will be used.
+#' @param deltad integer, similar to \code{deltat}, the interval in days for the
+#'        grid on the x-axis. Default is \code{7L} (aggregate to weekly values).
+#' @param col vector of colors forwarded to \code{image.default}. By default
+#'        a gray scale color map is used.
+#' @param contours logical \code{TRUE} or \code{FALSE}, whether or not the
+#'        concours should be added.
+#' @param contour.col color for the contour lines, only used if \code{contours = TRUE}.
+#' @param xlim limits of the x-axis, vector of length two. The limits have to lie
+#'          within \code{0} and \code{365} (Julian day). Can be decreasing
+#'          (e.g., \code{xlim = c(300, 100)}) to plot a period over new years eve.
+#'
+#' @details Plotting a Hovmoeller diagram based on the \code{\link{zoo}} time
+#' series object of the \code{\link{foehnix}} classification. Different plot
+#' types are available. The default functions (see list below) use \code{na.rm = TRUE}.
+#'
+#' Input \code{FUN} can be one of the following character strings:
+#' \itemize{
+#'    \item \code{freq}: plotting frequencies (default).
+#'    \item \code{mean}: mean probability.
+#'    \item \code{occ}: plotting occurrence of foehn (where probability \code{>= 0.5}).
+#'    \item \code{noocc}: contrary to \code{occ}: occurrence of no foehn (probability \code{< 0.5}).
+#' }
+#'
+#' \code{FUN} can also be a custom function used for time series aggregation
+#' (see \code{\link{aggregate.zoo}}).
+#' 
+#' @importFrom grDevices gray.colors
[…]

### 2019-08-08 R/windrose.R

@@ -60,7 +60,7 @@ utils::globalVariables(c("prob", "maxpp"))
 #'
 #' @examples
 #' # Loading combined demo data set
-#' data <- demodata()
+#' data <- demodata("tyrol") # default
 #'
 #' # Before estimating a model: plot a wind rose for all observations
 #' windrose(data$dd, data$ff, type = "histogram")
@@ -302,7 +302,7 @@ windrose.foehnix <- function(x, type = NULL, which = NULL, ddvar = "dd", ffvar =
 #' # Loading observation data for station Ellboegen.
 #' # The object returned is a zoo time series object
 #' # containing (see ?ellboegen).
-#' data <- demodata("Ellboegen")
+#' data <- demodata("ellboegen")
 #' class(data)
 #' head(data)
 #' 

### 2019-08-08 data-raw/ellboegen.R

@@ -1,6 +1,11 @@
 
+# Read csv data set
 ellboegen  <- utils::read.table("ellboegen.csv",  sep = ";",
                                 na.strings = "NA",
                                 header = TRUE, strip.white = TRUE)
-usethis::use_data(ellboegen)
+# Convert to zoo
+ellboegen <- zoo::zoo(ellboegen[, -1L],
+                 as.POSIXct(ellboegen[, 1L], tz = "UTC", origin = "1970-01-01"))
+# Add to package data folder
+usethis::use_data(ellboegen, version = 2)
 

### 2019-08-08 data-raw/luckyfive.R

@@ -1,6 +1,11 @@
 
+# Read csv data set
 luckyfive  <- utils::read.table("luckyfive.csv",  sep = ";",
                                 na.strings = "NA",
                                 header = TRUE, strip.white = TRUE)
-usethis::use_data(luckyfive)
+# Convert to zoo
+luckyfive <- zoo::zoo(luckyfive[, -1L],
+                 as.POSIXct(luckyfive[, 1L], tz = "UTC", origin = "1970-01-01"))
+# Add to package data folder
+usethis::use_data(luckyfive, version = 2)
 

### 2019-08-08 data-raw/sattelberg.R

@@ -1,6 +1,11 @@
 
-sattelberg <- utils::read.table("sattelberg.csv", sep = ";", 
+# Read csv data set
+sattelberg  <- utils::read.table("sattelberg.csv",  sep = ";",
                                 na.strings = "NA",
                                 header = TRUE, strip.white = TRUE)
-usethis::use_data(sattelberg)
+# Convert to zoo
+sattelberg <- zoo::zoo(sattelberg[, -1L],
+                 as.POSIXct(sattelberg[, 1L], tz = "UTC", origin = "1970-01-01"))
+# Add to package data folder
+usethis::use_data(sattelberg, version = 2)
 

### 2019-08-08 data-raw/viejas.R

@@ -1,6 +1,11 @@
 
+# Read csv data set
 viejas  <- utils::read.table("viejas.csv",  sep = ";",
                                 na.strings = "NA",
                                 header = TRUE, strip.white = TRUE)
-usethis::use_data(viejas)
+# Convert to zoo
+viejas <- zoo::zoo(viejas[, -1L],
+                 as.POSIXct(viejas[, 1L], tz = "UTC", origin = "1970-01-01"))
+# Add to package data folder
+usethis::use_data(viejas, version = 2)
 

### 2019-08-08 tests/testthat/test_foehnix.R

@@ -1,229 +1,0 @@
-
-
-library("foehnix")
-library("testthat")
-library("zoo")
-
-test_that("Testing foehnix function (main calls)", {
-
-    # Loading test data
-    expect_silent(data <- demodata("Ellboegen"))
-    expect_silent(data <- demodata("Sattelberg"))
-    expect_silent(data <- demodata())
-
-    # Specify our wind filter rule(s)
-    expect_silent(filter <- list(dd = c(43, 223), crest_dd = c(90, 270)))
-
-    # Testing wrong calls
-    # Input 'formula' is not a formula
-    expect_error(foehnix(3, data, TRUE))
-    expect_error(foehnix("wrong formula", data, TRUE))
-    # Data ste is not a 'zoo' object
-    expect_error(foehnix(ff ~ rh, data = data.frame(ff = c(1,2), rh = c(50,90))))
-    expect_error(foehnix(ff ~ rh, data = matrix(rnorm(10), ncol = 2)))
-    # Wrong switch argument
-    expect_error(foehnix(ff ~ rh, data, switch = "foo"))
-    # Wrong filter
-    expect_error(foehnix(ff ~ rh, data, filter = "foo"))
-    expect_error(foehnix(ff ~ rh, data, filter = c(1,2,3)))
-    # Wrong foehnix control
-    expect_error(foehnix(ff ~ rh, data, control = "foo"))
-    expect_error(foehnix(ff ~ rh, data, control = list(a = 1, b = 2)))
-    # Inexistent family
-    expect_error(foehnix(ff ~ rh, data, family = "foo family"))
-    # Formula contains variables not in 'data'
-    expect_error(foehnix("foo ~ bar", data, TRUE))
-
-
-    # Model: model will be added to the global 
-    # environment such that we can use it in the
-    # other test_that calls.
-    expect_silent(mod <- foehnix(diff_t ~ ff + rh, data = data, switch = TRUE,
-                                 filter = filter, verbose = FALSE))
-    # When maxit = 1 the algorithm cannot converge. Expect
-    # a message to be shown and the foehnix object will be returned.
-    expect_warning(x <- foehnix(diff_t ~ ff + rh, data = head(data,10000), 
-                                switch = TRUE, filter = filter,
[…]

### 2019-08-08 tests/testthat/test_foehnix_families.R

@@ -1,0 +1,66 @@
+
+
+library("foehnix")
+library("testthat")
+library("zoo")
+
+
+test_that("Testing foehnix families (simple test)", {
+
+    # Loading test data
+    expect_silent(data <- demodata("ellboegen"))
+
+    # Gaussian
+    expect_output(g1 <- foehnix(ff ~ rh, data = head(data, 1000), family = "gaussian"))
+    expect_output(g2 <- foehnix(ff ~ rh, data = head(data, 1000), family = foehnix:::foehnix_gaussian()))
+    expect_output(g3 <- foehnix(ff ~ rh, data = head(data, 1000), family = foehnix:::foehnix_cgaussian()))
+    expect_output(g4 <- foehnix(ff ~ rh, data = head(data, 1000), family = foehnix:::foehnix_tgaussian()))
+
+    expect_identical(c(AIC(g1), BIC(g1), logLik(g1)), c(AIC(g2), BIC(g2), logLik(g1)))
+    expect_identical(c(AIC(g3), BIC(g3), logLik(g3)), c(AIC(g4), BIC(g4), logLik(g4)))
+    expect_identical(sapply(list(g1, g2, g3, g4), class), rep("foehnix", 4))
+
+    expect_true(all(is.na(fitted(g1)) | (fitted(g1) >= 0 & fitted(g1) <= 1)))
+    expect_true(all(is.na(fitted(g2)) | (fitted(g2) >= 0 & fitted(g2) <= 1)))
+    expect_true(all(is.na(fitted(g3)) | (fitted(g3) >= 0 & fitted(g3) <= 1)))
+    expect_true(all(is.na(fitted(g4)) | (fitted(g4) >= 0 & fitted(g4) <= 1)))
+
+
+    # Logistic
+    expect_output(l1 <- foehnix(ff ~ rh, data = head(data, 1000), family = "logistic"))
+    expect_output(l2 <- foehnix(ff ~ rh, data = head(data, 1000), family = foehnix:::foehnix_logistic()))
+    expect_output(l3 <- foehnix(ff ~ rh, data = head(data, 1000), family = foehnix:::foehnix_clogistic()))
+    expect_output(l4 <- foehnix(ff ~ rh, data = head(data, 1000), family = foehnix:::foehnix_tlogistic()))
+
+    expect_identical(c(AIC(l1), BIC(l1), logLik(l1)), c(AIC(l2), BIC(l2), logLik(l1)))
[…]

### 2019-08-08 tests/testthat/test_foehnix_main.R

@@ -1,0 +1,51 @@
+
+
+library("foehnix")
+library("testthat")
+library("zoo")
+
+test_that("Testing foehnix function (main calls)", {
+
+    # Loading test data
+    expect_silent(data <- demodata("ellboegen"))
+    expect_silent(data <- demodata("sattelberg"))
+    expect_silent(data <- demodata("tyrol"))
+
+    # Specify our wind filter rule(s)
+    expect_silent(filter <- list(dd = c(43, 223), crest_dd = c(90, 270)))
+
+    # Testing wrong calls
+    # Input 'formula' is not a formula
+    expect_error(foehnix(3, data, TRUE))
+    expect_error(foehnix("wrong formula", data, TRUE))
+    # Data ste is not a 'zoo' object
+    expect_error(foehnix(ff ~ rh, data = data.frame(ff = c(1,2), rh = c(50,90))))
+    expect_error(foehnix(ff ~ rh, data = matrix(rnorm(10), ncol = 2)))
+    # Wrong switch argument
+    expect_error(foehnix(ff ~ rh, data, switch = "foo"))
+    # Wrong filter
+    expect_error(foehnix(ff ~ rh, data, filter = "foo"))
+    expect_error(foehnix(ff ~ rh, data, filter = c(1,2,3)))
+    # Wrong foehnix control
+    expect_error(foehnix(ff ~ rh, data, control = "foo"))
+    expect_error(foehnix(ff ~ rh, data, control = list(a = 1, b = 2)))
+    # Inexistent family
+    expect_error(foehnix(ff ~ rh, data, family = "foo family"))
+    # Formula contains variables not in 'data'
+    expect_error(foehnix("foo ~ bar", data, TRUE))
+
+
+    # Model: model will be added to the global 
+    # environment such that we can use it in the
+    # other test_that calls.
+    expect_silent(mod <- foehnix(diff_t ~ ff + rh, data = data, switch = TRUE,
+                                 filter = filter, verbose = FALSE))
+    # When maxit = 1 the algorithm cannot converge. Expect
+    # a message to be shown and the foehnix object will be returned.
+    expect_warning(x <- foehnix(diff_t ~ ff + rh, data = head(data,10000), 
+                                switch = TRUE, filter = filter,
+                                verbose = FALSE, maxit = 1))
+    expect_is(x, "foehnix")
+
+})
+

### 2019-08-08 tests/testthat/test_plot.R

@@ -6,7 +6,7 @@ library("testthat")
 # Prepare the models in the "global env" such that we only have
 # to run the model estimation twice for all kinds of plots!
 # Loading test data
-data <- demodata()
+data <- demodata("tyrol")
 filter <- list(dd = c(43, 223), crest_dd = c(90, 270))
 
 # Model: model will be added to the global 

### 2019-08-08 R/datasets.R
Update devel 0.1-3

@@ -166,7 +166,6 @@ NULL
 #'          \code{\link{viejas}}, \code{\link{luckyfive}}.
 #'
 #' @name luckyfive
-#' @docType data
-#' @keywords data luckyfive
-#' @author Reto Stauffer, Georg J. Mayr
 NULL
+
+

### 2019-08-08 tests/testthat/test_md5.R

@@ -7,12 +7,13 @@ library("testthat")
 # ------------------------------------------------------------
 test_that("md5 sums of CSV data sets", {
 
-    test <- list("ellboegen.csv"    = "a48b770830a688cb31dd56b9900bd46a",
-                 "sattelberg.csv"   = "2cc74fcb6c02361f0fed26e18be247a2",
-                 "viejas.csv"       = "f1a19ec04637fc231bc33b052ca2aeb5",
-                 "luckyfive.csv"    = "529ef6465418c45a7656c8dc2c027c73")
-    for (file in names(test)) {
-        expect_true(tools::md5sum(sprintf("../../data-raw/%s", file)) == test[[file]])
+    files <- list("ellboegen.csv"    = "a48b770830a688cb31dd56b9900bd46a",
+                  "sattelberg.csv"   = "2cc74fcb6c02361f0fed26e18be247a2",
+                  "viejas.csv"       = "f1a19ec04637fc231bc33b052ca2aeb5",
+                  "luckyfive.csv"    = "529ef6465418c45a7656c8dc2c027c73")
+    for (file in names(files)) {
+        x <- expect_true(tools::md5sum(sprintf("../../data-raw/%s", file)) == files[[file]])
+        if(!x) print(file)
     }
 
 })
@@ -23,12 +24,13 @@ test_that("md5 sums of CSV data sets", {
 # ------------------------------------------------------------
 test_that("md5 sums of CSV data sets", {
 
-    test <- list("ellboegen.rda"    = "9ab817dd24b80297c06d7108c241c45e",
-                 "sattelberg.rda"   = "cc65d0e08ab8be6dd882a98d5bf2766b",
-                 "viejas.rda"       = "842165bf047d8d8a9a51b039b7dda100",
-                 "luckyfive.rda"    = "7c939315ef2fb3b7f01ff9e98d05068c")
-    for (file in names(test)) {
-        expect_true(tools::md5sum(sprintf("../../data/%s", file)) == test[[file]])
+    files <- list("ellboegen.rda"    = "9ab817dd24b80297c06d7108c241c45e",
+                  "sattelberg.rda"   = "cc65d0e08ab8be6dd882a98d5bf2766b",
+                  "viejas.rda"       = "842165bf047d8d8a9a51b039b7dda100",
+                  "luckyfive.rda"    = "7c939315ef2fb3b7f01ff9e98d05068c")
+    for (file in names(files)) {
[…]

### 2019-08-09 R/demodata.R
Update; see if travis is fine with it so far

@@ -1,37 +1,22 @@
 
 #' foehnix Demo Data Set
 #'
-#' The foehnix package comes with two demo data sets containing
-#' observations from two automated weather stations in Tyrol,
-#' Austria. One station (Sattelberg) is located close to the main
-#' alpine ridge and is used as 'crest' station. The second station
-#' (Ellboegen) is located in the Wipp valley.
-#' The package demos and examples will estimate automated foehn
-#' classification models (\code{\link{foehnix}} models) for 
-#' station Ellboegen using additional information from the
-#' crest station as concomitant variables and for custom wind
-#' filters (\code{\link{foehnix_filter}}).
+#' The \code{\link{foehnix}} package comes with two sets of meteorological
+#' observations one for Tyrol, Austria, and one for Southern California, USA.
+#' Each region comes with observations from two stations, one valley station
+#' (or target station) and one station further upstream of the main wind direction
+#' (crest station) used to filter the data (see \code{\link{foehnix_filter}}).
+#' For Tyrol, observations for station Ellbögen (valley) and station Sattelberg
+#' (crest) are included, the Californian data set consists of the crest station
+#' 'Lucky Five Ranch' and the valley station 'Viejas Casino and Resort'.
+#' More details are provided in the 'Details' section.
 #'
-#' @param which the \code{foehnix} package comes with two demo data sets.
-#'      Each consists of meteorological observations of two stations (one
-#'      in the valley; target station) and one near the crest upstream of
-#'      typical foehn direction (crest station).
-#'      \code{which = "tyrol"} returns the tyrolean data set (A),
-#'      \code{which = "california"} the californian data set (USA), a time series
+#' @param which \code{which = "tyrol"} returns the combined tyrolean data set (A) and
+#'      \code{which = "california"} the combined californian data set (USA), a time series
 #'      object which contains measurements from both stations (valley/crest).
 #'      The function can also be used to get the demo data sets for specific sites
-#'      (see 'Examples').
+#'      (see 'Examples'/'Usage').
 #'
-#' @source
-#' Ellboegen and Sattelberg:
-#' Operated by the Department of Atmospheric and Cryospheric
-#' Sciences (\url{http://acinn.uibk.ac.at}) of the University of Innsbruck.
-#' The data is available under the
-#' \href{https://creativecommons.org/licenses/by-sa/4.0/}{CC BY-SA 4.0 license}.
-#'
[…]

### 2019-08-09 R/foehnix.R

@@ -11,7 +11,7 @@
 # -------------------------------------------------------------------
 # - EDITORIAL:   2018-11-28, RS: Created file on thinkreto.
 # -------------------------------------------------------------------
-# - L@ST MODIFIED: 2019-08-07 16:39 on marvin
+# - L@ST MODIFIED: 2019-08-09 16:36 on marvin
 # -------------------------------------------------------------------
 
 
@@ -37,8 +37,7 @@
 #'        Check manual of \code{\link{foehnix.control}} for more details.
 #' @param verbose logical, default \code{TRUE}. If set to \code{FALSE}
 #'        verbose output will be suppressed.
-#' @param \dots additional arguments used when calling \code{image.foehnix},
-#'        forwarded to \code{image} and \code{contour}.
+#' @param ... currently unused.
 #'
 #' @seealso
 #' \code{\link{foehnix.family}},
@@ -168,7 +167,7 @@ foehnix.noconcomitant.fit <- function(y, family, switch = FALSE,
 #'        Check manual of \code{\link{foehnix.control}} for more details.
 #' @param verbose logical, default \code{TRUE}. If set to \code{FALSE}
 #'        verbose output will be suppressed.
-#' @param ... additional arguments, unused.
+#' @param ... currently unused.
 #'
 #' @seealso
 #' \code{\link{foehnix.family}},
@@ -182,8 +181,7 @@ foehnix.noconcomitant.fit <- function(y, family, switch = FALSE,
 #' @import stats
 #' @import utils
 foehnix.unreg.fit <- function(y, logitX, family, switch = FALSE,
-                    maxit = 100L, tol = 1e-5, verbose = TRUE,
-                    ...) {
+                    maxit = 100L, tol = 1e-5, verbose = TRUE, ...) {
 
     # Lists to trace log-likelihood path and the development of
     # the coefficients during EM optimization.
@@ -307,7 +305,7 @@ foehnix.unreg.fit <- function(y, logitX, family, switch = FALSE,
 #'        containing the arguments for the glmnet function.
 #' @param verbose logical, default \code{TRUE}. If set to \code{FALSE}
 #'        verbose output will be suppressed.
-#' @param ... additional arguments, unused.
+#' @param ... currently unused.
 #'
 #' @seealso \code{\link{foehnix}}, \code{\link{foehnix.control}},
 #' \code{\link{foehnix.noconcomitant.fit}}, \code{\link{foehnix.reg.fit}},
[…]

### 2019-08-09 R/foehnix_functions.R

@@ -186,6 +186,94 @@ edf.foehnix <- function(object, ...)     structure(object$optimizer$edf, names =
 print.foehnix <- function(x, ...) print(summary(x, ...))
 
 
+#' Write Estimated Probabilities to CSV File
+#'
+#' Write the results of a \code{\link{foehnix}} model into a CSV
+#' text file. Custom date/time information can be specified using the
+#' input argument \code{format}. By default UNIX timestamp will be used.
+#'
+#' @param x a \code{\link{foehnix}} object
+#' @param file character, name of the target file
+#' @param info logical, whether or not header information
+#'      should be written
+#' @param format NULL or a character to specify the format in which the
+#'      date/time information should be written to the \code{file} (forwarded to
+#'      \code{strftime})
+#' @param ... currently ignored
+#'
+#' @return Invisible return of the data.frame written to the output file.
+#' @author Reto Stauffer
+#' @export
+write.csv.foehnix <- function(x, file, info = TRUE, format = NULL, ...) {
+
+    # Prepare the data set
+    if (is.null(format)) {
+        datetime <- data.frame(timestamp = as.numeric(index(x$prob)))
+        fmt_data <- "%10d; %6.4f; %4d"
+        fmt_head <- "%10s; %6s; %4s"
+    } else {
+        datetime <- data.frame(datetime = strftime(index(x$prob), format), 
+                               stringsAsFactors = FALSE)
+        nmax <- max(nchar(datetime$datetime), na.rm = TRUE)
+        fmt_data <- paste0("%", nmax, "s; %6.4f; %4d")
+        fmt_head <- paste0("%", nmax, "s; %6s; %4s")
+        if (anyDuplicated(datetime$datetime))
+            warning(sprintf("Custom format \"%s\" results in non-unique datetime information.",
+                    format))
+    }
+    # Combine
+    data <- cbind(datetime, as.data.frame(x$prob, row.names = 1:nrow(x$prob)))
+
+    # Custom information
+    if (info) {
+        formula_to_str <- function(x)
+            sprintf("%s ~ %s", as.character(x)[2L], as.character(x)[3L])
+        info <- list(
+             paste(rep("-", 70), collapse = ""),
+             "Results from foehnix classification model",
+             sprintf("Object name:     %s", deparse(substitute(x))),
+             sprintf("Model formula:   %s", formula_to_str(formula(x))),
+             sprintf("Output created:  %s", Sys.time()),
+             "Description:",
[…]

### 2019-08-09 R/image.R

@@ -17,9 +17,7 @@
 #' @param contours logical \code{TRUE} or \code{FALSE}, whether or not the
 #'        concours should be added.
 #' @param contour.col color for the contour lines, only used if \code{contours = TRUE}.
-#' @param xlim limits of the x-axis, vector of length two. The limits have to lie
-#'          within \code{0} and \code{365} (Julian day). Can be decreasing
-#'          (e.g., \code{xlim = c(300, 100)}) to plot a period over new years eve.
+#' @param ... additional arguments (see 'Details' section).
 #'
 #' @details Plotting a Hovmoeller diagram based on the \code{\link{zoo}} time
 #' series object of the \code{\link{foehnix}} classification. Different plot
@@ -35,28 +33,45 @@
 #'
 #' \code{FUN} can also be a custom function used for time series aggregation
 #' (see \code{\link{aggregate.zoo}}).
+#'
+#' Additional arguments which can be set:
+#' \itemize{
+#'      \item \code{xlim}: limits of abscissa. Numeric vector of length 2 with
+#'            Julian days (0 - 365). If the values are decreasing (e.g., \code{xlim = c(300, 100)})
+#'            the abscissa will be adjusted to show continuous data around new years eve.
+#'      \item \code{ylim}: limits of ordinate. Numeric vector of length 2 with
+#'            seconds (seconds of the day; 0 = 00:00:00 UTC, 86400 = 24:00:00 UTC).
+#'      \item \code{xlab}, \code{ylab}, \code{main}: axis labels and title of the plot.
+#' }
 #' 
 #' @importFrom grDevices gray.colors
-#' @rdname foehnix
+#' @name image
 #' @export
 image.foehnix <- function(x, FUN = "freq", deltat = NULL, deltad = 7L,
                           col = rev(gray.colors(20)), contours = FALSE,
-                          contour.col = "black", xlim = c(0, 364),
-                          ylim = c(0, 86400), ...) {
+                          contour.col = "black", ...) {
 
     stopifnot(inherits(x, "foehnix"))
 
+    # Some arguments for the plot
+    arg <- list(...)
[…]

### 2019-08-09 R/tsplot.R

@@ -1,7 +1,7 @@
 
 
 # Some global variable(s) to make R CMD check happy
-utils::globalVariables("vars")
+utils::globalVariables(c("vars", "enabled"))
 
 
 #' foehnix Time Series Plot Config
@@ -22,7 +22,7 @@ utils::globalVariables("vars")
 #' @param style character, name of the style template (\code{default}, \code{advanced},
 #'      \code{bw}) or path to a file containing the required information.
 #' @param windsector vector or list to highlight specific wind sectors.
-#'      See \code{\link[foehnix]{foehnix:::windsector_convert}} for details
+#'      See \code{foehnix:::windsector_convert} for details
 #' @param var used when calling the \code{\link[foehnix]{tsplot_get_control}}
 #'     function. Name of the (original!) variable name
 #' @param property the property which should be returned by
@@ -459,12 +459,11 @@ tsplot <- function(x, start = NULL, end = NULL, ndays = 10,
     for (k in seq_along(start)) {
 
         # Parameters of the graphical output device
-        #par(mfrow = c(Nplots, 1), ask = FALSE, mar = rep(0.1, 4),
-        #    xaxs = "i", oma = c(4.1, 4.1, 2, 5.1))
-        par(ask = FALSE, mar = rep(0.1, 4), xaxs = "i", oma = c(4.1, 4.1, 2, 5.1))
+        par(ask = FALSE, mar = rep(0.1, 4), xaxs = "i", oma = c(4.1, 5.1, 2.6, 5.1))
         hgt <- ifelse(names(doplot)[unlist(doplot)] == "tempdiff", 1, 2)
         layout(matrix(1:Nplots, ncol = 1), heights = hgt)
-        # Pick subset to plot
+
+        # Pick data subset to plot
         tmp <- window(data, start = start[k], end = end[k])
 
         # Calculate the limits for the gray boxes (where prob >= .5)
@@ -501,7 +500,7 @@ tsplot <- function(x, start = NULL, end = NULL, ndays = 10,
 
 }
 
-#' Helper function to add the gray boxes (background)
+# Helper function to add the gray boxes (background)
 tsplot_add_boxes <- function(x, col = "gray90") {
     # Loaded from parent env
     if ( length(x$up) > 0 ) {
@@ -550,6 +549,7 @@ tsplot_calc_prob_boxes <- function(x) {
 }
 
 
+# Add legend to time series plot (tsplot)
[…]

### 2019-08-09 tests/testthat/test_md5.R

@@ -2,35 +2,49 @@
 library("foehnix")
 library("testthat")
 
-# ------------------------------------------------------------
-# ------- testing the csv files in data-raw (demodata) -------
-# ------------------------------------------------------------
-test_that("md5 sums of CSV data sets", {
+# We cannot do these checks during R CMD check as the files
+# will not be available (see .Rbuildignore; data will be packed).
+if (!is_testing()) {
+    
+    # ------------------------------------------------------------
+    # ------- testing the csv files in data-raw (demodata) -------
+    # ------------------------------------------------------------
+    test_that("md5 sums of CSV data sets", {
+    
+        files <- list("ellboegen.csv"    = "a48b770830a688cb31dd56b9900bd46a",
+                      "sattelberg.csv"   = "2cc74fcb6c02361f0fed26e18be247a2",
+                      "viejas.csv"       = "f1a19ec04637fc231bc33b052ca2aeb5",
+                      "luckyfive.csv"    = "529ef6465418c45a7656c8dc2c027c73")
+        cat("\n\n", getwd(), "\n\n")
+        for (file in names(files)) {
+            # Path relative to this file.
+            fullfile <- sprintf("../../data-raw/%s", file)
+            ##cat("\n--------------------\n", fullfile, "\n",
+            ##    getwd(), "\n", tools::md5sum(fullfile), "->", files[[file]], "\n")
+            expect_true(tools::md5sum(fullfile) == files[[file]])
+        }
+    
+    })
+    
+    
+    # ------------------------------------------------------------
+    # --------- testing the rda files in data (demodata) ---------
+    # ------------------------------------------------------------
+    test_that("md5 sums of CSV data sets", {
+    
+        files <- list("ellboegen.rda"    = "9ab817dd24b80297c06d7108c241c45e",
+                      "sattelberg.rda"   = "cc65d0e08ab8be6dd882a98d5bf2766b",
+                      "viejas.rda"       = "842165bf047d8d8a9a51b039b7dda100",
+                      "luckyfive.rda"    = "7c939315ef2fb3b7f01ff9e98d05068c")
+        for (file in names(files)) {
+            # Path relative to the "tests/testthat.R" script
+            # Path relative to this file.
+            fullfile <- sprintf("../../data/%s", file)
[…]

### 2019-08-09 tests/testthat/test_results.R

@@ -1,0 +1,40 @@
+
+
+library("foehnix")
+library("testthat")
+
+test_that("Testing results for models, Tyrolean data set", {
+
+    # Loading test data
+    expect_silent(data_ell   <- demodata("ellboegen"))
+    expect_silent(data_tyrol <- demodata("tyrol"))
+
+    # Specify our wind filter rule(s)
+    expect_silent(filter_ell   <- list(dd = c(43, 223)))
+    expect_silent(filter_tyrol <- list(dd = c(43, 223), crest_dd = c(90, 270)))
+
+    # Temporary file to store the results
+    tmpfile <- tempfile()
+
+    expect_silent(ell1 <- foehnix(ff ~ 1, data = data_ell, filter = filter_ell, verbose = FALSE))
+    expect_silent(write.csv(ell1, file = tmpfile, info = FALSE))
+    #print(tools::md5sum(tmpfile))
+    expect_true(tools::md5sum(tmpfile) == "02d5c0c4ab6afafb27e178977e89fe1b")
+
+    expect_silent(ell2 <- foehnix(ff ~ rh, data = data_ell, filter = filter_ell, verbose = FALSE))
+    expect_silent(write.csv(ell2, file = tmpfile, info = FALSE))
+    #print(tools::md5sum(tmpfile))
+    expect_true(tools::md5sum(tmpfile) == "1aca3601dfe1926be5cb54d0739bbf84")
+
+    expect_silent(tyr1 <- foehnix(ff ~ 1, data = data_tyrol, filter = filter_tyrol, verbose = FALSE))
+    expect_silent(write.csv(tyr1, file = tmpfile, info = FALSE))
+    #print(tools::md5sum(tmpfile))
+    expect_true(tools::md5sum(tmpfile) == "0825a50c768fca512590f1ce49d0f37d")
+
+    expect_silent(tyr2 <- foehnix(ff ~ 1, data = data_tyrol, filter = filter_tyrol, verbose = FALSE))
+    expect_silent(write.csv(tyr2, file = tmpfile, info = FALSE))
+    #print(tools::md5sum(tmpfile))
+    expect_true(tools::md5sum(tmpfile) == "0825a50c768fca512590f1ce49d0f37d")
+
+})
+

### 2019-08-11 R/image.R
Viejas vignette, small fixes image, custom breaks windrose

@@ -191,8 +191,8 @@ image.foehnix <- function(x, FUN = "freq", deltat = NULL, deltad = 7L,
     # 0 - 354 (0 based Julian day) along the x-axis, and
     # 0 - 86400 (one full day in seconds) along the y-axis.
     plot(NA, bty = "n",
-         xlim = arg$xlim, xaxt = "n", xaxs = "i",
-         ylim = arg$ylim, yaxt = "n", yaxs = "i",
+         xlim = sort(arg$xlim), xaxt = "n", xaxs = "i",
+         ylim = arg$ylim,       yaxt = "n", yaxs = "i",
          xlab = NA, ylab = NA, main = NA)
     mtext(side = 1, line = 3.0, arg$xlab)
     mtext(side = 2, line = 3.3, arg$ylab)

### 2019-08-11 R/windrose.R

@@ -22,6 +22,7 @@ utils::globalVariables(c("prob", "maxpp"))
 #'        the wind direction information.
 #' @param ffvar character, name of the column in the training data set which contains
 #'        the wind speed (or gust speed) data.
+#' @param breaks \code{NULL} or a numeric vector to define the wind speed (\code{ff}) breaks.
 #' @param mfcol integer, number of columns of subplots.
 #' @param maxpp integer (\code{>0}), maximum plots per page. Not all plots fit on one
 #'        page the script loops trough.
@@ -109,7 +110,7 @@ windrose <- function(x, ...) UseMethod("windrose")
 #' @rdname windrose
 #' @export
 windrose.foehnix <- function(x, type = NULL, which = NULL, ddvar = "dd", ffvar = "ff",
-                             mfcol = 2L, maxpp = Inf, ...) {
+                             breaks = NULL, mfcol = 2L, maxpp = Inf, ...) {
 
     # Just to be on the very safe side.
     stopifnot(inherits(x, "foehnix"))
@@ -169,7 +170,7 @@ windrose.foehnix <- function(x, type = NULL, which = NULL, ddvar = "dd", ffvar =
                        "For more information please read see ?windrose.foehnix."))
         } else {
             # Else picking ff from x$data
-            ff <- x$data[,ffvar]
+            ff <- x$data[, ffvar]
         }
     } else if ( ! inherits(ddvar, "zoo") ) {
         stop("Input ddvar has to be of class 'zoo' or 'character'.")
@@ -259,9 +260,9 @@ windrose.foehnix <- function(x, type = NULL, which = NULL, ddvar = "dd", ffvar =
 #' @param dd can be used if univariate objects or vectors are provided for
 #'        wind speed and meteorological wind direction (see 'Details' section).
 #' @param filter a custom set of filter rules (see \code{\link{foehnix_filter}}).
-#' @param var.dd string, custom name of the wind direction variable in \code{x}
+#' @param ddvar string, custom name of the wind direction variable in \code{x}
 #'        if \code{x} is a multivariate object (see 'Details' section).
-#' @param var.ff string, custom name of the wind speed variable in \code{x}
+#' @param ffvar string, custom name of the wind speed variable in \code{x}
 #'        if \code{x} is a multivariate object (see 'Details' section).
 #'
[…]

### 2019-08-13 index.Rmd
Fixed Umlaut
Update 0.1-4

@@
+# foehnix
+
+[![Build Status](https://api.travis-ci.org/retostauffer/Rfoehnix.svg?branch=master)](https://api.travis-ci.org/retostauffer/Rfoehnix.svg?branch=master)
+[![codecov](https://codecov.io/gh/retostauffer/Rfoehnix/branch/master/graph/badge.svg)](https://codecov.io/gh/retostauffer/Rfoehnix)
+[![Python](https://img.shields.io/badge/Python-githubio-yellowgreen)](https://matthiasdusch.github.io/foehnix-python/_build/html/)
+
 ### Toolbox for Automated Foehn Classification based on Mixture Models
@@
 <p id="gruen2007">
-Gr\"un B, Leisch F (2007).
+Gr&uuml;n B, Leisch F (2007).
 Fitting Finite Mixtures of Generalized Linear Regressions in _R_.

### 2019-08-13 R/foehnix_filter.R

@@ -35,6 +35,9 @@
 #'        function operating on \code{x}, or a named list with a simple
 #'        filter rule (\code{numeric} of length two) or custom filter
 #'        functions. Details provided in the 'Details' section.
+#' @param cols \code{NULL} or a character vector containing the names in 'x'
+#'        which are not allowed to contain missing values.
+#'        If \code{NULL} all elements have to be non-missing.
 #' @param ... currently unused.
 #'
 #' @return Returns a vector of integers corresponding to those rows in
@@ -175,16 +178,15 @@
 #' 
 #' @author Reto Stauffer
 #' @export
-foehnix_filter <- function(x, filter) {
+foehnix_filter <- function(x, filter, cols = NULL) {
 
     # Wrong input 'x' 
-    if ( ! inherits(x, c("zoo", "data.frame")) )
+    if (! inherits(x, c("zoo", "data.frame")))
         stop("Input \"x\" to foehnix_filter has to be of class zoo or data.frame.")
 
     # ---------------
-    # If NULL: return integer sequence corresponding to
-    # the number of rows in 'x'.
-    if ( is.null(filter) ) {
+    # If filter is NULL
+    if (is.null(filter)) {
         tmp <- rep(TRUE, nrow(x))
 
     # ---------------
@@ -246,8 +248,17 @@ foehnix_filter <- function(x, filter) {
         tmp <- apply(tmp, 1, function(x) ifelse(any(is.na(x)), NA, all(x)))
     }
 
+    # Set rows with missing values to 'NA'
+    if (is.null(cols)) {
+        isnan <- which(apply(x, 1, function(x) any(is.na(x))))
+    } else {
+        isnan <- which(apply(subset(x, select = cols), 1, function(x) any(is.na(x))))
+    }
+    if (length(isnan) > 0) tmp[isnan] <- NA
+
+
     # Wrong length? Stop!
-    if ( ! length(tmp) == nrow(x) )
+    if (!length(tmp) == nrow(x))
         stop("Filter did not return expected length (not 'nrow(x)'). Stop.")
 
     # Else return a list with indizes corresponding to the filter

### 2019-08-13 R/foehnix_functions.R

@@ -362,21 +362,21 @@ summary.foehnix <- function(object, eps=1e-4, detailed = FALSE, ...) {
     rval$coef     <- coef(object, type = "parameter")
 
     # Posteriors
-    tab <- matrix(NA, ncol = 4, nrow = 2,
-                  dimnames = list(c("Component 1 (foehn)", "Component 2 (no foehn)"),
-                                  c("prior", "size", "post>0", "ratio")))
+    separation <- matrix(NA, ncol = 4, nrow = 2,
+                         dimnames = list(c("Component 1 (foehn)", "Component 2 (no foehn)"),
+                                         c("prior", "size", "post>0", "ratio")))
     priorfun <- function(x) { x <- mean(x, na.rm = TRUE); c(x, 1 - x) }
     sizefun  <- function(x) { N <- sum(!is.na(x)); x <- sum(x >= 0.5, na.rm = TRUE); c(x, N - x) }
     postfun  <- function(x, eps) { x <- na.omit(x); c(sum(x > eps), sum((1 - x) > eps)) }
 
     # Analog to flexmix summary
-    tab[, "prior"]   <- priorfun(object$optimizer$prob) # mean prior probability
-    tab[, "size"]    <- sizefun(object$optimizer$post)  # on posterior probability
-    tab[, "post>0"]  <- postfun(object$optimizer$post, eps = eps) # on posterior prob
-    tab[, "ratio"]   <- tab[, "size"] / tab[, "post>0"]
+    separation[, "prior"]   <- priorfun(object$optimizer$prob) # mean prior probability
+    separation[, "size"]    <- sizefun(object$optimizer$post)  # on posterior probability
+    separation[, "post>0"]  <- postfun(object$optimizer$post, eps = eps) # on posterior prob
+    separation[, "ratio"]   <- separation[, "size"] / separation[, "post>0"]
 
     # Optimizer statistics
-    rval$csummary   <- tab # cluster summary (separation)
+    rval$separation <- separation # cluster summary (separation)
     rval$filter_obj <- object$filter_obj
     rval$time       <- object$time
     rval$logLik     <- logLik(object)
@@ -447,7 +447,7 @@ print.summary.foehnix <- function(x, ...) {
         # non-vanishing posteriors pnk should also be assigned to the corresponding
         # cluster, giving a ratio close to 1.
         cat("\nCluster separation (ratios close to 1 indicate\nwell separated clusters):\n")
-        print(round(x$csummary, 2))
[…]

### 2019-08-13 R/plot.foehnix.R

@@ -7,7 +7,7 @@
 # -------------------------------------------------------------------
 # - EDITORIAL:   2018-12-16, RS: Created file on thinkreto.
 # -------------------------------------------------------------------
-# - L@ST MODIFIED: 2019-08-12 17:11 on marvin
+# - L@ST MODIFIED: 2019-08-13 11:15 on marvin
 # -------------------------------------------------------------------
 
 
@@ -187,23 +187,32 @@ plot.foehnix <- function(x, which = NULL, log = TRUE, ..., ask = TRUE) {
             # Position where to draw the density
             at   <- seq(min(tmp$y), max(tmp$y), length = 501)
 
+            # Breaks
             bk   <- seq(min(tmp$y), max(tmp$y), length = 50)
+
+            # Calculate the histograms
             h1   <- hist(tmp$y[which(tmp$prob <  .5)], plot = FALSE, breaks = bk, include.lowest = TRUE)
             h2   <- hist(tmp$y[which(tmp$prob >= .5)], plot = FALSE, breaks = bk, include.lowest = TRUE)
-            ylim <- c(0, max(h1$density, h2$density, na.rm = TRUE))
+            # Calculate density
+            d1   <-  x$control$family$d(at, x$coef$mu1, exp(x$coef$logsd1))
+            d2   <-  x$control$family$d(at, x$coef$mu2, exp(x$coef$logsd2))
+
+            ylim <- c(0, max(h1$density, h2$density, d1, d2, na.rm = TRUE))
             xlim <- range(bk)
 
             # Plotting conditional component 1 histogram
             plot(h1, freq = FALSE, xlim = xlim, ylim = ylim,
                  main = "Conditional Histogram\nComponent 1 (no foehn)",
                  border = "gray50", xlab = expression(paste("y[",pi < 0.5,"]")))
-            lines(at, x$control$family$d(at, x$coef$mu1, exp(x$coef$logsd1)), col = 2, lwd = 2)
+            lines(at, d2, col = "gray50", lwd = .5, lty = 5)
+            lines(at, d1, col = 2, lwd = 2)
 
             # Plotting conditional component 2 histogram
             plot(h2, freq = FALSE, xlim = xlim, ylim = ylim,
                  main = "Conditional Histogram\nComponent 2 (foehn)",
                  border = "gray50", xlab = expression(paste("y[",pi >= 0.5,"]")))
[…]

### 2019-08-13 tests/testthat/test_results.R

@@ -24,7 +24,7 @@ test_that("Testing results for models, Tyrolean data set", {
     expect_silent(ell2 <- foehnix(ff ~ rh, data = data_ell, filter = filter_ell, verbose = FALSE))
     expect_silent(write.csv(ell2, file = tmpfile, info = FALSE))
     #print(tools::md5sum(tmpfile))
-    expect_true(tools::md5sum(tmpfile) == "1aca3601dfe1926be5cb54d0739bbf84")
+    expect_true(tools::md5sum(tmpfile) == "1807cfb60f0b28438ab0fe80621a8ca1")
 
     expect_silent(tyr1 <- foehnix(ff ~ 1, data = data_tyrol, filter = filter_tyrol, verbose = FALSE))
     expect_silent(write.csv(tyr1, file = tmpfile, info = FALSE))

### 2019-08-15 index.Rmd
Update to 0-1.4

--- 2019-08-08 index.Rmd
@@
+# foehnix
+
+[![Build Status](https://api.travis-ci.org/retostauffer/Rfoehnix.svg?branch=master)](https://api.travis-ci.org/retostauffer/Rfoehnix.svg?branch=master)
+[![codecov](https://codecov.io/gh/retostauffer/Rfoehnix/branch/master/graph/badge.svg)](https://codecov.io/gh/retostauffer/Rfoehnix)
+[![Repository Status](https://www.repostatus.org/badges/latest/wip.svg)](https://www.repostatus.org/badges/latest/wip.svg)
+[![Python](https://img.shields.io/badge/Python-githubio-yellowgreen)](https://matthiasdusch.github.io/foehnix-python/_build/html/)
+
 ### Toolbox for Automated Foehn Classification based on Mixture Models

### 2019-08-15 R/datasets.R

@@ -1,14 +1,15 @@
 
-#' @usage
-#' # Combined data set Tyrol (A).
-#' # Valley station: Ellbögen; crest station: Sattelberg
-#' demodata()
-#' demodata(which = "tyrol")
-#'
-#' # Combined data set California (CA, USA).
-#' # Valley station: Viejas; crest station: Lucky Five Ranch
-#' demodata(which = "california")
-#'
+# @usage
+# # Combined data set Tyrol (A).
+# # Valley station: Ellbögen; crest station: Sattelberg
+# demodata()
+# demodata(which = "tyrol")
+#
+# # Combined data set California (CA, USA).
+# # Valley station: Viejas; crest station: Lucky Five Ranch
+# demodata(which = "california")
+#
+
 #' @details
 #' The Tyrolean data set consists of station
 #' Elbögen located in the Wipp valley, an area well known for south foehn
@@ -61,10 +62,11 @@
 #' @author Reto Stauffer, Deborah Detka
 NULL
 
-#' @usage
-#' # Hourly observations station Ellögen (Tyrol, Austria).
-#' # Serves as "valley" or "target" station for the "tyrol" data set.
-#' data(which = "ellboegen")
+# @usage
+# # Hourly observations station Ellögen (Tyrol, Austria).
+# # Serves as "valley" or "target" station for the "tyrol" data set.
+# data(which = "ellboegen")
+
 #' @source Station Ellögen operated by the Department of Atmospheric and Cryospheric
 #' Sciences (\url{http://acinn.uibk.ac.at}) of the University of Innsbruck.
 #' The data is available under the
@@ -73,10 +75,11 @@ NULL
 #' @rdname demodata
 NULL
 
-#' @usage
-#' # Hourly observations station Sattelberg (Tyrol, Austria).
-#' # Serves as "crest" station for the "tyrol" data set.
-#' data(which = "sattelberg")
+# @usage
+# # Hourly observations station Sattelberg (Tyrol, Austria).
+# # Serves as "crest" station for the "tyrol" data set.
+# data(which = "sattelberg")
+
 #' @source Station Sattelberg operated by the Department of Atmospheric and Cryospheric
 #' Sciences (\url{http://acinn.uibk.ac.at}) of the University of Innsbruck.
 #' The data is available under the
@@ -85,20 +88,22 @@ NULL
 #' @rdname demodata
 NULL
 
-#' @usage
-#' # Hourly observations station Viejas Casino and Resort (California, U.S.A.).
[…]

### 2019-08-15 R/demodata.R

@@ -76,7 +76,7 @@ demodata <- function(which = c("tyrol", "ellboegen", "sattelberg",
     stopifnot(identical(sort(names(stations)), c("crest", "valley")))
 
     # Load data sets
-    obj  <- data(list = stations, package = "foehnix")
+    obj  <- data(list = stations, package = "foehnix", envir = environment())
     data <- lapply(obj, function(x) return(eval(parse(text = x))))
 
     # Rename crest station variables

### 2019-08-15 R/foehnix.R

@@ -11,7 +11,7 @@
 # -------------------------------------------------------------------
 # - EDITORIAL:   2018-11-28, RS: Created file on thinkreto.
 # -------------------------------------------------------------------
-# - L@ST MODIFIED: 2019-08-13 15:16 on marvin
+# - L@ST MODIFIED: 2019-08-15 10:36 on marvin
 # -------------------------------------------------------------------
 
 
@@ -427,7 +427,7 @@ foehnix.reg.fit <- function(y, logitX, family, glmnet.control, switch = FALSE,
 #'        also input \code{left} and input \code{truncated}. Can be set to any finite
 #'        numeric value.
 #' @param truncated logical. If set to \code{TRUE} truncation is used instead of
-#'        \code{censored}. This only affects the \code{\link{foehnix}} model estimate
+#'        censoring. This only affects the \code{\link{foehnix}} model estimate
 #'        if input \code{left} and/or input \code{right} are specified.
 #' @param standardize logical flag, default is \code{TRUE}. Defines whether or
 #'        not the model matrix for the concomitant model should be standardized for
@@ -560,6 +560,7 @@ print.foehnix.control <- function(x, ...) str(x)
 #' This is the main method of the foehnix package to estimate
 #' two-component mixture models for automated foehn classification.
 #' 
+#' @param x a \code{foehnix} object.
 #' @param formula an object of class \code{formula} (or one that can be coerced
 #'        to that class): a symbolic description of the model to be fitted.  The
 #'        details of model specification are given under 'Details'.
@@ -581,6 +582,8 @@ print.foehnix.control <- function(x, ...) str(x)
 #' @param object a \code{foehnix} object (input to S3 methods)
 #' @param detailed boolean, default FALSE. If set to true, additional information
 #'        will be printed.
+#' @param eps threshold for posterior probabilities used in \code{summary}.
+#'        \code{numeric} value, default \code{1e-4}.
 #' @param ... forwarded to \code{\link{foehnix.control}}
 #'
 #' @return Returns an object of class \code{foehnix}.

### 2019-08-15 R/windrose.R

@@ -265,6 +265,8 @@ windrose.foehnix <- function(x, type = NULL, which = NULL, ddvar = "dd", ffvar =
 #'        if \code{x} is a multivariate object (see 'Details' section).
 #' @param ffvar string, custom name of the wind speed variable in \code{x}
 #'        if \code{x} is a multivariate object (see 'Details' section).
+#' @param breaks \code{NULL} or a sequence of \code{numeric} values to set
+#'        the breaks along \code{ff}.
 #'
 #' @details The \code{\link{windrose}} function can be used in different ways.
 #' The main purpose is to plot (conditional) wind roses for
@@ -497,6 +499,7 @@ windrose.default <- function(x, ff,
     # Breaks for classification
     dd.breaks <- seq(-interval / 2, 360, by=interval)
     ff.breaks <- if (!is.null(breaks)) breaks else pretty(ff)
+    stopifnot(is.numeric(ff.breaks))
 
     # Picking some colors
     cols <- rev(colorspace::heat_hcl(length(ff.breaks) - 1, h = hue, c. = c(60, 10),
@@ -566,12 +569,12 @@ windrose.default <- function(x, ff,
     # Adding polygons (density)
     if ( type == "density" ) {
         args <- as.list(match.call(expand.dots = TRUE))
-        args <- c(list(x = NA, y = NA, col = NA), args[c("border", "lty", "lwd")])
+        args <- c(list(x = NA, y = NA, col = NA),
+                  args[names(args) %in% c("border", "lty", "lwd")])
         for ( i in ncol(tab):1 ) { 
             tmp <- (-1) * ddff2uv(dd.breaks + interval / 2, c(tab[,i], tab[1,i]))
             args$x <- tmp$u; args$y = tmp$v; args$col = cols[i]
             do.call(polygon, args)
-            #polygon(tmp$u, tmp$v, col = cols[i], border = border)
         }
     # Plot type "histogram"
     } else {

### 2019-08-15 tests/testthat/test_windrose.R

@@ -194,3 +194,25 @@ test_that("Testing custom filters for windrose", {
     expect_silent(windrosefun(data, filter = list(hour = function(x) x == 12)))
 
 })
+
+test_that("Windrose customization", {
+
+    data <- demodata("sattelberg")
+
+    # Custom breaks
+    expect_silent(windrosefun(data, breaks = seq( 0, 35, by = 5)))
+    # Breaks not spannind the range of the data
+    expect_warning(windrosefun(data, breaks = seq(10, 35, by = 5)))
+    expect_warning(windrosefun(data, breaks = seq( 0, 20, by = 5)))
+    expect_error(windrosefun(data, breaks = "foo"))
+
+    # Custom borders
+    expect_silent(windrosefun(data, border = NA))
+    expect_silent(windrosefun(data, border = "red", lty = 3, lwd = 3))
+    expect_error(windrosefun(data, border = "red", lty = "foo", lwd = 3))
+    expect_error(windrosefun(data, lty = "foo"))
+    expect_error(windrosefun(data, lwd = "foo"))
+    # Well, this results in an error and an additional warning message
+    expect_warning(expect_error(windrosefun(data, lwd = "foo")))
+
+})

### 2019-08-15 tests/testthat.R
Rename test files; added Makefile; update vignette ellboegen

@@ -1,6 +1,15 @@
 # Run all unit tests
 
+
+# Test filter
+filter <- "[^(md5)]"
+
+#files <- list.files("testthat")
+#files <- gsub("(test-|\\.R)", "", files)
+#take  <- files[grep(filter, files)]
+#print(c(files = length(files), take = length(take)))
+
 # This is a wrapper around all the tests specified
 # in the tests/testthat directory. These tests are visible
 # to users and serve as small examples as well.
-testthat::test_check("foehnix")
+testthat::test_check("foehnix", filter = filter)

### 2019-08-15 tests/testthat/test-windrose.R

@@ -211,7 +211,6 @@ test_that("Windrose customization", {
     expect_silent(windrosefun(data, border = "red", lty = 3, lwd = 3))
     expect_error(windrosefun(data, border = "red", lty = "foo", lwd = 3))
     expect_error(windrosefun(data, lty = "foo"))
-    expect_error(windrosefun(data, lwd = "foo"))
     # Well, this results in an error and an additional warning message
     expect_warning(expect_error(windrosefun(data, lwd = "foo")))
 

### 2019-08-16 R/demodata.R
Update

@@ -1,20 +1,20 @@
 
 #' foehnix Demo Data Set
 #'
-#' The \code{\link{foehnix}} package comes with two sets of meteorological
-#' observations one for Tyrol, Austria, and one for Southern California, USA.
-#' Each region comes with observations from two stations, one valley station
-#' (or target station) and one station further upstream of the main wind direction
-#' (crest station) used to filter the data (see \code{\link{foehnix_filter}}).
+#' The \code{\link[foehnix]{foehnix}} package comes with two sets of meteorological
+#' observations: one for Tyrol, Austria, and one for Southern California, USA.
+#' Both data sets come with observations from two stations, one valley station
+#' (or target station) and one station further upstream of the main foehn wind direction
+#' (crest station) used to filter the data (see \code{\link[foehnix]{foehnix_filter}}).
 #' For Tyrol, observations for station Ellbögen (valley) and station Sattelberg
 #' (crest) are included, the Californian data set consists of the crest station
 #' 'Lucky Five Ranch' and the valley station 'Viejas Casino and Resort'.
 #' More details are provided in the 'Details' section.
 #'
-#' @param which \code{which = "tyrol"} returns the combined tyrolean data set (A) and
-#'      \code{which = "california"} the combined californian data set (USA), a time series
+#' @param which \code{which = "tyrol"} returns the combined Tyrolean data set (A) and
+#'      \code{which = "california"} the combined Californian data set (USA), a time series
 #'      object which contains measurements from both stations (valley/crest).
-#'      The function can also be used to get the demo data sets for specific sites
+#'      Else, the data for a specific site will be returned
 #'      (see 'Examples'/'Usage').
 #'
 #' @examples
@@ -45,9 +45,6 @@
 #' # Lucky Five Ranch
 #' x <- demodata("luckyfive")
 #' print(head(x))
-#'
-#' @seealso \code{\link{ellboegen}}, \code{\link{sattelberg}},
-#'          \code{\link{viejas}}, \code{\link{luckyfive}}
 #'
 #' @author Reto Stauffer
 #' @export

### 2019-08-16 R/families.R

@@ -1,14 +1,3 @@
-# -------------------------------------------------------------------
-# - NAME:        families.R
-# - AUTHOR:      Reto Stauffer
-# - DATE:        2018-12-13
-# -------------------------------------------------------------------
-# - DESCRIPTION:
-# -------------------------------------------------------------------
-# - EDITORIAL:   2018-12-13, RS: Created file on thinkreto.
-# -------------------------------------------------------------------
-# - L@ST MODIFIED: 2019-08-05 12:03 on marvin
-# -------------------------------------------------------------------
 
 
 #' foehnix Families for Two-Component Mixture Models
@@ -71,6 +60,7 @@
 #' @name foehnix.family
 #' @author Reto Stauffer
 #' @export
+#' @author Reto Stauffer
 print.foehnix.family <- function(x, ...) {
 
     cat(sprintf("foehnix family of class: %s\n", x$name))
@@ -86,6 +76,9 @@ print.foehnix.family <- function(x, ...) {
 }
 
 
+#' @usage
+#' # Check if truncation is set
+#' is.truncated(x, ...)
 #' @rdname foehnix.family
 #' @export
 is.truncated <- function(x, ...) UseMethod("is.truncated")
@@ -94,6 +87,9 @@ is.truncated <- function(x, ...) UseMethod("is.truncated")
 is.truncated.foehnix.family <- function(x, ...) "truncated" %in% names(x)
 
 
+#' @usage
+#' # Check if left censoring/truncation limit has been specified
+#' has.left(x, ...)
 #' @rdname foehnix.family
 #' @export
 
@@ -107,6 +103,9 @@ has.left.foehnix.family <- function(x, ...) {
 }
 
 
+#' @usage
+#' # Check if right censoring/truncation limit has been specified
+#' has.left(x, ...)
 #' @rdname foehnix.family
 #' @export
 has.right <- function(x, ...) UseMethod("has.right")
@@ -123,6 +122,10 @@ has.right.foehnix.family <- function(x, ...) {
 # -------------------------------------------------------------------
 # Logistic foehnix mixture model family.
 # -------------------------------------------------------------------
+
+#' @usage
+#' # Logistic mixture-model family
+#' foehnix_logistic()
 #' @rdname foehnix.family
 #' @import stats
 foehnix_logistic <- function() {
@@ -198,6 +201,9 @@ foehnix_logistic <- function() {
 # Censored logistic foehnix mixture model family.
[…]

### 2019-08-16 R/foehnix.R

@@ -1,29 +1,13 @@
-# -------------------------------------------------------------------
-# - NAME:        foehnix.R
-# - AUTHOR:      Reto Stauffer
-# - DATE:        2018-11-28
-# -------------------------------------------------------------------
-# - DESCRIPTION: Scripts to estimate Gaussian and logistic
-#                two-component mixture models with IWLS and
-#                weighted empirical moments.
-#                "simple" here referrs to non-gradient non-optimizer
-#                based model estimates.
-# -------------------------------------------------------------------
-# - EDITORIAL:   2018-11-28, RS: Created file on thinkreto.
-# -------------------------------------------------------------------
-# - L@ST MODIFIED: 2019-08-15 10:36 on marvin
-# -------------------------------------------------------------------
-
 
 
 #' Fitting foehnix Mixture Model Without Concomitant Model.
 #'
 #' Fitting method for two-component foehnix mixture model without
 #' concomitants.
-#' Typically not called directly, interfaced via \code{\link{foehnix}}.
+#' Typically not called directly, interfaced via \code{\link[foehnix]{foehnix}}.
 #' @param y numeric vector, covariate for the components of the
 #'        mixture model, dimension \code{N}.
-#' @param family object of class \code{\link{foehnix.family}}.
+#' @param family object of class \code{\link[foehnix]{foehnix.family}}.
 #' @param switch logical whether or not the two components should be switched.
 #'        By default (\code{switch = FALSE}) the component which shows
 #'        higher values of \code{y} is assumed to be the foehn cluster!
@@ -31,21 +15,22 @@
 #'        the clusters (by setting \code{switch = TRUE}).
 #' @param maxit positive integer, or vector of length 2 with positive
 #'        integer values. Maximum number of iterations of the EM algorithm.
-#'        Check manual of \code{\link{foehnix.control}} for more details.
+#'        Check manual of \code{\link[foehnix]{foehnix.control}} for more details.
 #' @param tol numeric, or vector of length 2 containing numeric values.
 #'        Tolerance for the EM algorithm.
-#'        Check manual of \code{\link{foehnix.control}} for more details.
+#'        Check manual of \code{\link[foehnix]{foehnix.control}} for more details.
[…]

### 2019-08-16 R/foehnix_filter.R

@@ -1,29 +1,16 @@
-# -------------------------------------------------------------------
-# - NAME:        foehnix_filter.R
-# - AUTHOR:      Reto Stauffer
-# - DATE:        2018-12-19
-# -------------------------------------------------------------------
-# - DESCRIPTION:
-# -------------------------------------------------------------------
-# - EDITORIAL:   2018-12-19, RS: Created file on thinkreto.
-# -------------------------------------------------------------------
-# - L@ST MODIFIED: 2019-08-07 16:31 on marvin
-# -------------------------------------------------------------------
 
 
-#' Evaluates Data Filter Rules for foehnix Mixture Moel Calls.
+#' Evaluates Data Filter Rules for foehnix Mixture Moel Calls
 #' 
-#' \code{\link{foehnix}} models allow to specify an optional
+#' \code{\link[foehnix]{foehnix}} models allow to specify an optional
 #' \code{foehnix_filter}. If a filter is given only a subset
-#' of the data set provided to \code{\link{foehnix}} is used
+#' of the data set provided to \code{\link[foehnix]{foehnix}} is used
 #' for the foehn classification.
-#' 
 #' A typical example is a wind direction filter such that
 #' only observations (times) are used where the observed
 #' wind direction was within a user defined wind sector
 #' corresponding to the wind direction during foehn events
 #' for a specific location.
-#' 
 #' However, the filter option allows to even implement complex
 #' filter rules if required. The 'Details' section contains
 #' further information and examples how this filter rules can
@@ -46,7 +33,7 @@
 #' 
 #' @details Foehn winds often (not always) show a very specific wind direction
 #' due to the canalization of the air flow trough the local topography. The
-#' \code{\link{foehnix_filter}} option allows to subset the data according to a
+#' \code{\link[foehnix]{foehnix_filter}} option allows to subset the data according to a
 #' user defined set of filters from simple filters to complex filters.
 #'
 #' These filters classify each observation (each row in \code{x}) as
@@ -70,7 +57,7 @@
 #'         as 'good' (within filter) if \code{x$dd >= 43 & x$dd <= 223}.
 #'     \item \code{filter = list(dd = c(330, 30)}: similar to the filter
 #'         rule above, allows to specify a wind sector going trough 0
[…]

### 2019-08-16 R/foehnix_functions.R

@@ -5,7 +5,7 @@ utils::globalVariables(c("time_mid", "yday_mid", "value"))
 #'
 #' Function to standardize the columns of a matrix, used to
 #' standardize the model matrix before estimating the regression
-#' coefficients of a generalized linear model (\code{\link{iwls_logit}}).
+#' coefficients of a generalized linear model (\code{\link[foehnix]{iwls_logit}}).
 #'
 #' @param x matrix of dimension \code{N x p}.
 #' @param ... additional arguments, ignored.
@@ -28,11 +28,17 @@ utils::globalVariables(c("time_mid", "yday_mid", "value"))
 #' print(head(S))
 #'
 #' is.standardized(X)
-#' is.standardized(X)
+#' is.standardized(S)
 #'
 #' # Get parameters used for standardization
 #' center(S)
 #' scale(S)
+#'
+#' # Destandardize
+#' D <- destandardize(S)
+#'
+#' # Check
+#' all.equal(D, X, check.attributes = FALSE)
 #'
 #' @author Reto Stauffer
 #' @export
@@ -57,11 +63,17 @@ standardize.matrix <- function(x, ...) {
 }
 
 
+#' @usage
+#' # Returns 'scaled:scale' used for standardization
+#' scale(x, ...)
 #' @rdname standardize
 #' @export
 scale.standardized <- function(x, ...) return(attr(x, "scaled:scale"))
 
 
+#' @usage
+#' # Returns 'scaled:center' used for standardization
+#' center(x, ...)
 #' @rdname standardize
 #' @export
 center <- function(x, ...) UseMethod("center")
@@ -70,13 +82,17 @@ center.standardized <- function(x, ...) return(attr(x, "scaled:center"))
 
 #' Destandardize a Standardized Matrix
 #'
-#' Reverse function of \code{\link[foehnix]{standardize}}.
+#' @details Reverse function of \code{\link[foehnix]{standardize}}.
 #'
 #' @param x standardized matrix of class \code{standardized}.
 #' @param ... forwarded, unused.
 #'
 #' @seealso \code{\link[foehnix]{standardize}}
 #' @author Reto Stauffer
+#'
+#' @examples
+#' # See example on R documentation for \code{?standardize}.
+#'
 #' @export
 destandardize <- function(x, ...) UseMethod("destandardize")
 
@@ -93,7 +109,7 @@ destandardize.standardized <- function(x, ...) {
 
[…]

### 2019-08-16 R/wind_functions.R

@@ -6,35 +6,35 @@
 # -------------------------------------------------------------------
 # - EDITORIAL:   2018-12-21, RS: Created file on thinkreto.
 # -------------------------------------------------------------------
-# - L@ST MODIFIED: 2019-04-16 15:47 on marvin
+# - L@ST MODIFIED: 2019-08-16 15:07 on marvin
 # -------------------------------------------------------------------
 
-#' Convert U and V Wind Components to Wind Speed and Direction
+#' Convert u and v Wind Components to Wind Speed and Wind Direction
 #'
-#' Takes U/V wind components as input and returns wind speed
-#' and meteorological wind direction.
+#' Takes u/v wind components (zonal and meridional wind components) as input
+#' and returns wind speed and meteorological wind direction.
 #'
-#' @param u \code{numeric} vector with U components or a 
+#' @param u \code{numeric} vector with u components or a 
 #'    \code{data.frame} or \code{zoo} object containing
 #'    a column named \code{u} and a column named \code{v}.
 #' @param v \code{NULL} (default) or \code{numeric} vector.
 #'    If input \code{u} is a single vector \code{v} has to be specified.
 #' @param rad \code{logical}, default is \code{FALSE}. Returns wind
-#'    direction in radiant rather than in degrees.
+#'    direction in radiant rather than degrees.
 #'
-#' @return Returns a data.frame or zoo object (depending on input
-#' object \code{u}) with two columns named \code{dd}
+#' @return Returns a \code{data.frame} or \code{zoo} object (depending on input
+#' \code{u}) with two columns named \code{dd}
 #' and \code{ff} containing wind speed (same physical unit as
 #' input \code{u}/\code{v}) and wind direction. Wind direction
 #' is either in meteorological degrees (0 from North, from 90 East,
 #' 180 from South, and 270 from West) or in mathematical radiant
 #' if input \code{rad = TRUE}.
 #'
-#' @author Jakob Messner, Reto Stauffer
+#' @author Reto Stauffer
 #'
 #' @details Note: if both, \code{u} and \code{v} are provided they
 #' do have to be of the same length OR one of them has to be of length 1.
-#' The one with length 1 will be repeated.
+#' The one with length 1 will be recycled.
 #'
 #' @seealso ddff2uv
 #' @examples
@@ -112,39 +112,40 @@ uv2ddff <- function(u, v = NULL, rad = FALSE){
    if (is.null(zoo_index)) return(res) else return(zoo(res, zoo_index))
 }
 
[…]

### 2019-08-16 R/windrose.R

@@ -23,7 +23,7 @@ utils::globalVariables(c("prob", "maxpp"))
 #' @param ffvar character, name of the column in the training data set which contains
 #'        the wind speed (or gust speed) data.
 #' @param breaks \code{NULL} or a numeric vector to define the wind speed (\code{ff}) breaks.
-#' @param mfcol integer, number of columns of subplots.
+#' @param ncol integer, number of columns of subplots.
 #' @param maxpp integer (\code{>0}), maximum plots per page. Not all plots fit on one
 #'        page the script loops trough.
 #'
@@ -110,7 +110,7 @@ windrose <- function(x, ...) UseMethod("windrose")
 #' @rdname windrose
 #' @export
 windrose.foehnix <- function(x, type = NULL, which = NULL, ddvar = "dd", ffvar = "ff",
-                             breaks = NULL, mfcol = 2L, maxpp = Inf, ...) {
+                             breaks = NULL, ncol = NULL, maxpp = Inf, ...) {
 
     # Just to be on the very safe side.
     stopifnot(inherits(x, "foehnix"))
@@ -126,10 +126,9 @@ windrose.foehnix <- function(x, type = NULL, which = NULL, ddvar = "dd", ffvar =
     if ( is.null(type) ) type <- allowed
     type <- match.arg(type, allowed, several.ok = TRUE)
 
-    # Input argument mfcol
-    if ( inherits(mfcol, "numeric") ) { mfcol <- as.integer(mfcol) }
-    else { stopifnot(inherits(mfcol, "integer")) }
-    stopifnot(mfcol > 0)
+    # Input argument ncol
+    if (inherits(ncol, "numeric")) ncol <- as.integer(ncol)
+    stopifnot(inherits(ncol, c("NULL", "integer")) | ncol <= 0)
 
     # Input argument maxpp
     if ( !maxpp == Inf ) {
@@ -184,18 +183,18 @@ windrose.foehnix <- function(x, type = NULL, which = NULL, ddvar = "dd", ffvar =
     # Else start plotting
     # Step 1: calculate number of pages/rows/columns based on
     # the number of requested plots (given which/type) and the
-    # user specification mfcol/maxpp.
+    # user specification ncol/maxpp.
     hold <- par(no.readonly = TRUE); on.exit(par(hold))
 
[…]

### 2019-08-16 tests/testthat/test-tsplot.R

@@ -25,36 +25,6 @@ test_that("Testing tsplot.foehnix function", {
     expect_is(mod0, "foehnix")
     expect_is(mod1, "foehnix")
 
-    # Testing tsplot.control
-    expect_silent(res <- tsplot.control())
-    expect_is(res, "tsplot.control")
-
-    expect_silent(res <- tsplot.control(foo = "baa"))
-    expect_silent(res <- tsplot.control(t = "T2m"))
-
-    # Renaming covariates
-    expect_identical(res$t$name, "T2m")
-    expect_silent(res <- tsplot.control(t = "T2m", rh = "relhum"))
-    expect_identical(res$t$name,  "T2m")
-    expect_identical(res$rh$name, "relhum")
-    expect_silent(res <- tsplot.control(t = list(name = "T2m"), rh = list(name = "relhum")))
-    expect_identical(res$t$name,  "T2m")
-    expect_identical(res$rh$name, "relhum")
-
-    # Adding fresh color
-    expect_silent(res <- tsplot.control(t = list(col = "red")))
-    expect_identical(res$t$col, "#FF0000")
-    expect_silent(res <- tsplot.control(t = list(col = 2)))
-    expect_identical(res$t$col, "#FF0000")
-
-    # Adding a new label
-    expect_silent(res <- tsplot.control(t = list(ylab = "foo")))
-    expect_identical(res$t$ylab, "foo")
-
-    # Adding all in one go
-    args <- list(dd = list(name = "winddir", col = 2, ylab = "my wind direction"),
-                 ff = list(name = "windspd", col = "#001122", ylab = "my wind speed"))
-
     # Plotting helper function for the tests
     tsplotfun <- function(x, ...) {
         pdf(file = tempfile())
@@ -76,19 +46,7 @@ test_that("Testing tsplot.foehnix function", {
     expect_error(tsplotfun(mod0, end = 11111))
     expect_error(tsplotfun(mod0, ndays = "foo"))
     expect_error(tsplotfun(mod0, ndays = -10))
-    expect_error(tsplotfun(mod0, control = list(foo = "bar")))
-    expect_error(tsplotfun(mod0, control = 1))
     expect_error(tsplotfun(data))
-
-    # Renaming ALL covariates. Should result in an error as none of the
-    # subplots can be plotted anymore.
-    modX <- mod0; modX$prob$prob <- NA
[…]

### 2019-08-16 R/families.R
Update, fixed docs

@@ -513,7 +513,7 @@ foehnix_gaussian <- function() {
 
 #' @usage
 #' # Censored Gaussian mixture-model family
-#' foehnix_cgaussian()
+#' foehnix_cgaussian(left = -Inf, right = Inf)
 #' @rdname foehnix.family
 #' @import stats
 #' @useDynLib foehnix, .registration = TRUE 
@@ -628,7 +628,7 @@ foehnix_cgaussian <- function(left = -Inf, right = Inf) {
 
 #' @usage
 #' # Truncated Gaussian mixture-model family
-#' foehnix_tgaussian()
+#' foehnix_tgaussian(left = -Inf, right = Inf)
 #' @rdname foehnix.family
 #' @import stats
 #' @useDynLib foehnix, .registration = TRUE 

### 2019-08-16 R/foehnix.R

@@ -905,7 +905,7 @@ foehnix <- function(formula, data, switch = FALSE, filter = NULL,
 #' @param object a \code{\link[foehnix]{foehnix}} mixture model object.
 #' @param newdata if \code{NULL} (default) the prediction of the
 #'        underlying training data set will be returned
-#'        (see also \code{\link[foehnix]{fitted}}). Else \code{newdata}
+#'        (see also \code{\link[foehnix]{fitted.foehnix}}). Else \code{newdata}
 #'        has to be a \code{zoo} object providing the required
 #'        variables which have been used for model fitting
 #'        and filtering (see \code{\link[foehnix]{foehnix}}).

### 2019-08-16 index.Rmd
Update, fixed docs
Add logo

@@
-# foehnix
+# foehnix <a href='https://retostauffer.github.io/Rfoehnix/'><img src='https://github.com/retostauffer/Rfoehnix/raw/master/foehnix-logo.png' align="right" height="139" /></a>
 
@@
 ```{r, echo = FALSE, fig = TRUE, results = 'hide', fig.width = 12, fig.height = 4}
-windrose(mod, type = "histogram", mfcol = 3)
+windrose(mod, type = "histogram", ncol = 3)
 ```

### 2019-09-10 R/windrose.R
Update

@@ -26,6 +26,8 @@ utils::globalVariables(c("prob", "maxpp"))
 #' @param ncol integer, number of columns of subplots.
 #' @param maxpp integer (\code{>0}), maximum plots per page. Not all plots fit on one
 #'        page the script loops trough.
+#' @param main \code{NULL} (default) or character. If \code{NULL} the function will
+#'        add default figure titles.
 #'
 #' @details Allows to draw windrose plots from \code{\link{foehnix}} mixture model
 #' object or a set of observations. If input \code{x} to \code{\link{windrose}} is
@@ -110,7 +112,7 @@ windrose <- function(x, ...) UseMethod("windrose")
 #' @rdname windrose
 #' @export
 windrose.foehnix <- function(x, type = NULL, which = NULL, ddvar = "dd", ffvar = "ff",
-                             breaks = NULL, ncol = NULL, maxpp = Inf, ...) {
+                             breaks = NULL, ncol = NULL, maxpp = Inf, ..., main = NULL) {
 
     # Just to be on the very safe side.
     stopifnot(inherits(x, "foehnix"))
@@ -201,28 +203,28 @@ windrose.foehnix <- function(x, type = NULL, which = NULL, ddvar = "dd", ffvar =
     if ( "density_unconditional" %in% which )
         tmp[["density_unconditional"]] <- with(data,
               windrose(dd, ff, hue = c(10, 130), breaks = breaks,
-                       main = expression("Unconditional"), ...))
+                       main = if (!is.null(main)) main else expression("Unconditional"), ...))
     if ( "density_nofoehn" %in% which )
         tmp[["density_nofoehn"]] <- with(subset(data, prob < 0.5),
               windrose(dd, ff, hue = c(100, 180), breaks = breaks,
-                       main = expression(paste("No Foehn [",pi < 0.5,"]")), ...))
+                       main = if (!is.null(main)) main else expression(paste("No Foehn [",pi < 0.5,"]")), ...))
     if ( "density_foehn" %in% which )
         tmp[["density_foehn"]] <- with(subset(data, prob >= 0.5),
               windrose(dd, ff, hue = c(-20, 30), breaks = breaks,
-                       main = expression(paste("Foehn [",pi >= 0.5,"]")), ...))
+                       main = if (!is.null(main)) main else expression(paste("Foehn [",pi >= 0.5,"]")), ...))
     # Plot type = "histogram"
     if ( "histogram_unconditional" %in% which )
[…]

### 2019-09-10 tests/testthat/test-results.R

@@ -5,36 +5,39 @@ library("testthat")
 
 test_that("Testing results for models, Tyrolean data set", {
 
-    # Loading test data
-    expect_silent(data_ell   <- demodata("ellboegen"))
-    expect_silent(data_tyrol <- demodata("tyrol"))
+    # MD5 only works on linux (windows calculates the hash different)
+    if (.Platform$OS.type == "unix") {
+        # Loading test data
+        expect_silent(data_ell   <- demodata("ellboegen"))
+        expect_silent(data_tyrol <- demodata("tyrol"))
 
-    # Specify our wind filter rule(s)
-    expect_silent(filter_ell   <- list(dd = c(43, 223)))
-    expect_silent(filter_tyrol <- list(dd = c(43, 223), crest_dd = c(90, 270)))
+        # Specify our wind filter rule(s)
+        expect_silent(filter_ell   <- list(dd = c(43, 223)))
+        expect_silent(filter_tyrol <- list(dd = c(43, 223), crest_dd = c(90, 270)))
 
-    # Temporary file to store the results
-    tmpfile <- tempfile()
+        # Temporary file to store the results
+        tmpfile <- tempfile()
 
-    expect_silent(ell1 <- foehnix(ff ~ 1, data = data_ell, filter = filter_ell, verbose = FALSE))
-    expect_silent(write.csv(ell1, file = tmpfile, info = FALSE))
-    #print(tools::md5sum(tmpfile))
-    expect_true(tools::md5sum(tmpfile) == "02d5c0c4ab6afafb27e178977e89fe1b")
+        expect_silent(ell1 <- foehnix(ff ~ 1, data = data_ell, filter = filter_ell, verbose = FALSE))
+        expect_silent(write.csv(ell1, file = tmpfile, info = FALSE))
+        #print(tools::md5sum(tmpfile))
+        expect_true(tools::md5sum(tmpfile) == "02d5c0c4ab6afafb27e178977e89fe1b")
 
-    expect_silent(ell2 <- foehnix(ff ~ rh, data = data_ell, filter = filter_ell, verbose = FALSE))
-    expect_silent(write.csv(ell2, file = tmpfile, info = FALSE))
-    #print(tools::md5sum(tmpfile))
-    expect_true(tools::md5sum(tmpfile) == "1807cfb60f0b28438ab0fe80621a8ca1")
[…]

### 2019-09-20 R/foehnix_functions.R
Change format write.csv; new md5 sums (R/python)

@@ -267,13 +267,13 @@ write.csv.foehnix <- function(x, file, info = TRUE, format = NULL, ...) {
     # Prepare the data set
     if (is.null(format)) {
         datetime <- data.frame(timestamp = as.numeric(index(x$prob)))
-        fmt_data <- "%10d; %6.4f; %4d"
+        fmt_data <- "%10d; %6.3f; %4d"
         fmt_head <- "%10s; %6s; %4s"
     } else {
         datetime <- data.frame(datetime = strftime(index(x$prob), format), 
                                stringsAsFactors = FALSE)
         nmax <- max(nchar(datetime$datetime), na.rm = TRUE)
-        fmt_data <- paste0("%", nmax, "s; %6.4f; %4d")
+        fmt_data <- paste0("%", nmax, "s; %6.3f; %4d")
         fmt_head <- paste0("%", nmax, "s; %6s; %4s")
         if (anyDuplicated(datetime$datetime))
             warning(sprintf("Custom format \"%s\" results in non-unique datetime information.",

### 2019-09-20 tests/testthat/test-results.R

@@ -21,22 +21,22 @@ test_that("Testing results for models, Tyrolean data set", {
         expect_silent(ell1 <- foehnix(ff ~ 1, data = data_ell, filter = filter_ell, verbose = FALSE))
         expect_silent(write.csv(ell1, file = tmpfile, info = FALSE))
         #print(tools::md5sum(tmpfile))
-        expect_true(tools::md5sum(tmpfile) == "02d5c0c4ab6afafb27e178977e89fe1b")
+        expect_true(tools::md5sum(tmpfile) == "b604d81ce93f47d6479acf57c52f53a2")
 
         expect_silent(ell2 <- foehnix(ff ~ rh, data = data_ell, filter = filter_ell, verbose = FALSE))
         expect_silent(write.csv(ell2, file = tmpfile, info = FALSE))
         #print(tools::md5sum(tmpfile))
-        expect_true(tools::md5sum(tmpfile) == "1807cfb60f0b28438ab0fe80621a8ca1")
+        expect_true(tools::md5sum(tmpfile) == "c2fadc758ddbcb1973872f4a8c4f3db5")
 
         expect_silent(tyr1 <- foehnix(ff ~ 1, data = data_tyrol, filter = filter_tyrol, verbose = FALSE))
         expect_silent(write.csv(tyr1, file = tmpfile, info = FALSE))
         #print(tools::md5sum(tmpfile))
-        expect_true(tools::md5sum(tmpfile) == "0825a50c768fca512590f1ce49d0f37d")
+        expect_true(tools::md5sum(tmpfile) == "43562856bc775e05f8f17d3deb27447d")
 
-        expect_silent(tyr2 <- foehnix(ff ~ 1, data = data_tyrol, filter = filter_tyrol, verbose = FALSE))
+        expect_silent(tyr2 <- foehnix(ff ~ rh + diff_t, data = data_tyrol, filter = filter_tyrol, verbose = FALSE))
         expect_silent(write.csv(tyr2, file = tmpfile, info = FALSE))
         #print(tools::md5sum(tmpfile))
-        expect_true(tools::md5sum(tmpfile) == "0825a50c768fca512590f1ce49d0f37d")
+        expect_true(tools::md5sum(tmpfile) == "f1cb14d57e0a8697fd8d0e3c6c0360ff")
     }
 
 })
